### 2014-02-07 thesis/_region_.tex
A lot of things...

\message{ !name(thesis.tex)}\documentclass{book}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{wrapfig}
\usepackage{amssymb}
\usepackage{amsfonts, amsmath}
\usepackage{mathrsfs}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\usepackage{bookmark}

\input{../physics_common.tex}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

\title{Mathematical Aspects of the Capital Market}
\author{Xie Xiaolei}
\date{\today}
\begin{document}

\message{ !name(thesis.tex) !offset(1382) }
\cite{NIST:DLMF}:

\message{ !name(thesis.tex) !offset(1837) }

\end{document}