### 2014-04-30 thesis/_region_.tex
1st version that has been handed to Sven. Gaussian Wishart Matrices are normized by 1/T.

\message{ !name(thesis.tex)}\documentclass{report}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{amssymb}
\usepackage{amsfonts, amsmath}
\usepackage{mathrsfs}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\usepackage{bookmark}

\input{../physics_common.tex}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

% \title{Stocks' Cross-Correlations}
% \subtitle{}

\author{Xie Xiaolei}
\date{\today}
\begin{document}

\message{ !name(thesis.tex) !offset(1283) }
 matrix
elements but becomes clumsy with regard to eigenvalues and
eigenvectors. Therefore we look at the following return series
instead:

\message{ !name(thesis.tex) !offset(1795) }

\end{document}