### 2014-04-30 thesis/price_models.tex

\chapter{Return Models}\label{chp:PriceModels}
In this chapter we review some of the discrete-time return models and
fit them to intraday returns. The intention is to assess the
applicability of these models and to understand their statistical
properties. The implications of these properties on the returns'
correlation and covariance matrix are studied in chapters
\ref{chp:Gaussian} and \ref{chp:CrossCorrelationFat}.

\S\ref{sec:Garch_model} reviews the GARCH model using the Nordea
15-minute returns as an example. \S\ref{sec:SLV_model} examines the
stochastic log-volatility model and discusses the possibility of
describing the residuals of the log-volotility series using the
Johnson Su distribution. The same Nordea return series is fitted as an
example. \S\ref{sec:XieCalc} derives the unconditional distribution
function of the stochastic log-volotility model and compares the
analytic results to the 30-minute returns of a number of Swedish
stocks.

\section{Gaussian Distribution}
The justification of modeling return series as independent,
identically distributed Gaussian variates comes from imagining the
price process $S(t)$ as a Brownian motion whose increment $\sigma
dw_t$ at each time step is independent and scales as $\sigma
\sqrt{dt}$. Then, by adding a drift term $\mu dt$ that represents some
deterministic trend in the price process, one can express the
price $S(t)$ as a stochastic differential equation:
\begin{eqnarray*}
  dS &=& S\mu dt + S\sigma dw_t \\
  d(\ln S) &=& (\mu - \frac{1}{2} \sigma^2)dt + \sigma dw_t \\
\end{eqnarray*}
It follows from this equation that $\ln S(t) - \ln S_0$, where $S_0$
is the price at time 0, has Gaussian distribution with mean $(\mu -
\sigma^2/2)t$ and variance $\sigma^2 t$. Therefore, in a discrete-time
model, where the length of each time step $\Delta t$ is fixed, the
return $r_t = \ln S(t + \Delta t) - \ln S(t)$ is assumed to be
Gaussian distributed and have mean and variance that are functions of
$\Delta t$.

The picture depicted above is of course overly simplified, and as we discussed in
\S\ref{sec:StylizedFacts}, the distribution of returns is not really
Gaussian. Nevertheless, the assumption of Gaussian distributed returns
underlies such important theories as Black and Scholes theory of
option pricing. Thus, albeit hugely inaccurate, the Gaussian
distribution as a returns model shall not be forgotten. In the next
few sections, we discuss some more realistic returns models.

\section{GARCH models} \label{sec:Garch_model}
``GARCH'' is the acronym for ``Generalized Autoregressive Conditional
Heteroscedasticity''. A GARCH(p, q) model is defined by the following
equation system:
\begin{equation}
  \label{eq:garch_def}
  \begin{aligned}
    r_t &= \mu + \epsilon_t \\
    \sigma_t^2 &= \alpha_0 + \sum_{i=1}^q \alpha_i \epsilon_{t-i}^2 +
    \sum_{i=1}^p \beta_i \sigma_{t-i}^2
  \end{aligned}
\end{equation}
where $\epsilon_t$ is a random variable of a prescribed distribution,
the most common ones being Gaussian, Student's t, Asymmetric Student's
t, etc; $\sigma_t^2$ is the variance of $\epsilon_t$
\cite{Bollerslev86}. The autocorrelation function of $\epsilon_t^2$ is
given by (c.f. \cite{Bollerslev87})
\begin{eqnarray*}
\rho_n &=& \sum_{i=1} ^{p \vee q} (\alpha_i + \beta_i) \rho_{n-i}
\text{ for $n > p$}
\end{eqnarray*}
where $\alpha_i$ with $i > q$ and $\beta_i$ with $i > p$ are taken as
zeros. $p \vee q$ denotes the maximum of p and q. From these
equations, it is clear that the partial autocorrelation function cuts
off at $\max(p, q)$.

To assess the applicability of GARCH models at the face of intra-day
returns, we study the {\it Nordea Bank 15min} returns sampled during
the period 2012/01/16 - 2012/04/20. Firstly we plot the autocorrelation
function (ACF) of $\epsilon_t^2$ from this sample, as shown
in figure \ref{fig:nordea_15min_acf}.

At the absence of GARCH effects, the ACF is expected to have an
asymptotic Gaussian distribution with mean 0 and variance 1/T
\cite{Bollerslev86, Bollerslev87}, which is clearly not the case in
figure \ref{fig:nordea_15min_acf}: The first 5 autocorrelations have
comparable sizes and do not fall off as in a Gaussian
scheme. Moreover, figure \ref{fig:nordea_15min_vlt_acf} shows 
even more clearly that the conditional variances of the series are
correlated. These observations suggest a GARCH(p, q) model can be
appropriate.

\begin{figure}[htb!]
  \centering
  \subfigure[]{
    \includegraphics[scale=0.4, clip=true, trim=95 236 118
    200]{../pics/nordea_15min_acf.pdf}
    \label{fig:nordea_15min_acf}
  }
  \subfigure[]{
    \includegraphics[scale=0.4, clip=true, trim=95 236 118
    200]{../pics/nordea_15min_vlt_acf.pdf}
    \label{fig:nordea_15min_vlt_acf}
  }
  \caption{\small \it Auto-correlations among Nordea Bank 15min
    returns. \ref{fig:nordea_15min_acf}: Auto-correlations (ACF) of the
    squared returns; \ref{fig:nordea_15min_vlt_acf}: Auto-correlations
    (ACF) of the conditional variances.}
\end{figure}

According to Andersen and Bollerslev et al \cite{Andersen03},
$\epsilon_t$ can be decomposed as $\epsilon_t = \sigma_t z_t$, where
$z_t \sim N(0, 1)$ and $\sigma_t$ can be estimated using realized
volatility. To find out whether this is true for the aforementioned
series we look at the probability plot of $\epsilon_t/\sigma_t$
(figure \ref{fig:nordea_bank_15min_z_prob}), which is consistent with
a standard Gaussian distribution except for a few outliers at the two
ends. Hence the fat tails of the probability density function can be
entirely accounted for by GARCH effects, implying a Gaussian
conditional distribution is sufficient for $z_t$.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=80 258 104
    220]{../pics/nordea_bank_15min_z_prob.pdf}
  \caption{\small \it{Probability plot of $z_t = \epsilon_t /
      \sigma_t$. $\epsilon_t$ are derived from Nordea Bank 15min
      returns while $\sigma_t$ are realized volatilities calculated
      using 30s returns within each 15min interval. Horizontal axis:
      $z_t$}; Vertical axis: cummulative probability function (CDF) of
    $z_t$, arranged on such a scale that the CDF of the standard
    Gaussian is a straight line.}
  \label{fig:nordea_bank_15min_z_prob}
\end{figure}

To determine the orders p and q of the GARCH model, we again check
the partial autocorrelation function of $\epsilon_t^2$. As just
mentioned, the first 5 partial autocorrelations are significant. Thus
we can tentatively estimate a GARCH(1, 5) model. Then by applying the
Lagrange multiplier test and accordingly removing insignificant
parameters, we obtain a GARCH(1, 3) model
\ref{tab:nordea_15min_garch}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    parameter & value & std. error & t statistic \\
    \hline
     $\alpha_0$ &   1.05668e-06 &   3.56871e-07 &    2.96096\\
     \hline
     $\beta_1$  &     0.687212  &   0.0299314   &    22.9596\\
     \hline
     $\alpha_1$ &      0.128614 &   0.0268144   &    4.79643\\
     \hline
     $\alpha_3$ &    0.0568022  &   0.0301503   &    1.88397 \\
     \hline
  \end{tabular}
  \caption{\small \it GARCH(1, 3) model of Nordea Bank 15min returns}
  \label{tab:nordea_15min_garch}
\end{table}

\section{ARIMA Log-volotility Models}\label{sec:SLV_model}
With the availability of transaction data, estimating conditional
volatility using returns sampled at a higher frequency (realized
volatility) gives superior acuracy and reliability. However, at which
frequency the time series should be sampled in order to give an
unbiased and consistent estimate of the volatility is not a trivial
question. Naively one would argue by the doctrine of statistics that
the often the series is sampled, the better the estimate. This is also
well supported in theory, since the quadratic variation which is the
theoretial variance of the time series (see \cite{Andersen03}, theorem
1 and corollary 1), is approximated as (see \cite{Protter05}, ch.II,
\S6, theorem 22):
\[
\var(p_{t+h} - p_t) = [r, r]_{t+h} - [r, r]_t = \lim_{n \to \infty}
\sum_{T^i \in \pi_n} (p^{T_{i+1}} - p^{T_i})^2
\]
where $p_t$ is the log price process and $r_t = p_t - p_0$. $\pi_n$
is a refining partition of the interval $[t, t+h]$.

However, due to the noise introduced by market micro-structure, there
is actually an optimal sampling frequency relative to $h$. While the
method to determine this optimal frequency is a subject of debate (see
for example \cite{Sahalia05}), it is not difficult to find a fairly
satisfactory frequency in practice.

Andersen and Bollerslev et al proved in \cite{Andersen03} that, with
the conditional volatility well approximated by the realized
volatility, $r_{t+h, h} = p_{t+h} - p_t$ is governed by a Gaussian
distribution (see theorem 2 of \cite{Andersen03}):
\[
r_{t+h, h} | \sigma(\mu_{t+s}, \sigma_{t+s})_{s \in [0, h]} \sim
N(\int_{0}^h \mu_{t+s} ds, \int_{0}^h \sigma_{t+s}^2 ds)
\]

Thus one can try a few frequencies and compare the distribution of
$r_{t+h, h}/\hat{\sigma}_{t+h, h}$ with the standard Gaussian. If the
two match, $\hat{\sigma}_{t+h, h}$ is a good estimate of $\sigma_{t+h,
  h}$, the volatility of $r_{t+h, h}$.

For the Nordea Bank 15-minute returns, $h = 15\text{min}$, and a
reasonally good sampling frequency is 1/30sec, as confirmed by the
probability plot \ref{fig:nordea_bank_15min_z_prob} as well as by the
autocorrelation functions of $r_{t+h, h} / \sigma_{t+h, h}$ and of
$(r_{t+h, h} / \sigma_{t+h, h})^2$ which are shown in
\ref{fig:nordea_15min_quotient_acf} and
\ref{fig:nordea_15min_quotient_squared_acf} respectively.
\begin{figure}[htb!]
  \centering
  \subfigure[ACF of $r_{t+h, h} / \sigma_{t+h, h}$]{
    \includegraphics[scale=0.4, clip=true, trim=95 236 118
    200]{../pics/nordea_15min_quotient_acf.pdf}
    \label{fig:nordea_15min_quotient_acf}
  }
  \subfigure[ACF of $(r_{t+h, h} / \sigma_{t+h, h})^2$]{
    \includegraphics[scale=0.4, clip=true, trim=95 236 118
    200]{../pics/nordea_15min_quotient_squared_acf.pdf}
    \label{fig:nordea_15min_quotient_squared_acf}
  }
  \caption{\small \it Nordea 15min $r_t/\sigma_t$ and
    $(r_t/\sigma_t)^2$ ACF.}
\end{figure}

We see in figure \ref{fig:nordea_15min_quotient_acf} and
\ref{fig:nordea_15min_quotient_squared_acf} there is essentially no
autocorrelation in the series $r_{t+h, h} / \sigma_{t+h, h}$ and
$(r_{t+h, h} / \sigma_{t+h, h})^2$, supporting the assumption that the
series is i.i.d Gaussian distributed. As a result, the problem of
modeling $r_{t+h, h}$ boils down to modeling $\sigma_{t+h,
  h}$. Andersen and Bollerslev et al reported that, for the exchange
rates between Deutch mark, yen and dollar, $\ln \sigma_{t+h, h}$ is
gaussian distributed \cite{Andersen03}. This is, however, not the case
for our series in question. In fact, in our case, $\ln \sigma_{t+h,
  h}$ is right skewed (skewness 0.3342) and leptokurtic (kurtosis
6.1006). See figure \ref{fig:nordea_15min_logvol_prob}
\begin{figure}[htb!]
  \centering
  %\vspace{-15mm}
  \includegraphics[scale=0.5, clip=true, trim=80 223 107
  4]{../pics/nordea_15min_logvol_prob.pdf}
  \caption{\small \it Probability plot of Nordea 15min
    $\ln\sigma_{t+h, h}$ unconditional distribution}
  \label{fig:nordea_15min_logvol_prob}
\end{figure}

Moreover, the series of $\ln\sigma_{t+h, h}$ shows long-lasting and
periodic autocorrelations with an apparent period of 33 (see figure
\ref{fig:nordea_15min_logvol_acf}). This suggests the series may be
described by an integrated seasonal ARMA\footnote{ARMA ---
  Autoregressive Moving Average} model. But before we set out
to specify and estimate this model of the logvolatility
$\ln\sigma_{t+h, h}$, it is useful to first write down the model of
the returns $r_t$:
\begin{eqnarray}
   r_{t+h, h} &=& \ln p_{t+h} - \ln p_t \nonumber \\
   r_{t+h, h} &=& \mu + \sigma_{t+h, h} b_t \label{eq:SLV_spec}
\end{eqnarray}
where $p_t$ is the price of the asset at time $t$ and $b_t \sim N(0,
1)$. For simplicity, in the rest of this chapter we shall just write
$t$ for the subscript ``t+h, h'', since the time interval $h$ is fixed
for each time series and is a known constant. In the case of the
Nordea Bank 15-minute returns, h = 15 minutes.

As has just been said, $\ln \sigma_t$ for the series in question has
long-lasting auto-correlations (often referred to as ``long
memory''). This suggests simplifying the series by differencing
\cite{BoxJenkins94}:
\[
w_t = (1-B)(1-B^s)\ln\sigma_t
\]
where $B$ is the back-shift operator and $s=33$ is the seasonality.

The autocorrelation function of the differenced process $w_t$, as shown in
figure \ref{fig:nordea_15min_w_acf}, clearly points to a seasonal
moving-average model: There are only 4 non-zero autocorrelations in
the plot, located at lags 1, 32, 33, 34, respectively; furthermore,
the two at 32 and 34 are approximately equal. Thus we can write down
the model as
\begin{eqnarray}
  w_t &=& (1 - \theta B)(1 - \Theta B^s) y_t \label{eq:nordea_w}
\end{eqnarray}
where $\theta$ and $\Theta$ are parameters to be determined and $y_t$
is a noise process with constant variance $\sigma_y^2$ and mean
0. $y_t$ is often refered to as the residuals.
\begin{figure}[htb!]
  \centering
  \subfigure[ACF of $\ln\sigma_{t+h,h}$]{
    \includegraphics[scale=0.4, clip=true, trim=95 230 112
    235]{../pics/nordea_15min_logvol_acf.pdf}
    \label{fig:nordea_15min_logvol_acf}
  }
  \subfigure[ACF of $w_t$]{
    \includegraphics[scale=0.4, clip=true, trim=95 230 112
    235]{../pics/nordea_15min_w_acf.pdf}
    \label{fig:nordea_15min_w_acf}
  }
  \caption{\small \it Nordea 15min $\ln\sigma_{t+h, h}$ and $w_t$
    autocorrelations}
\end{figure}

The above seasonal moving average model has the following
autocovariance structure \cite{BoxJenkins94}:
\begin{eqnarray*}
  \gamma_0 &=& \sigma_y^2 (1 + \theta^2)(1 + \Theta^2) \\
  \gamma_1 &=& -\sigma_y^2\theta(1 + \Theta^2) \\
  \gamma_s &=& -\sigma_y^2\Theta(1 + \theta^2) \\
  \gamma_{s+1} &=& \gamma_{s-1}\;=\;\sigma_y^2\theta\Theta
\end{eqnarray*}
These equations together with the measured autocorrelations make
possible an initial estimate of the parameters $\theta$ and $\Theta$:
\begin{eqnarray*}
  {\rho_{s+1}/\rho_s} &=& {\gamma_{s+1}/\gamma_s} \;=\; -{\theta \over
    1 + \theta^2} \\
  {\rho_{s+1}/\rho_1} &=& {\gamma_{s+1}/\gamma_1} \;=\; -{\Theta \over
    1 + \Theta^2} \\
\end{eqnarray*}
Substituting in the measured values shown in table
\ref{tab:nordea_15min_w_acf},
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    $\rho_1$ & $\rho_{s-1}$ & $\rho_s$ & $\rho_{s+1}$ \\
    \hline
    -0.4703 &  0.2053 & -0.4564 &  0.2212 \\
    \hline
  \end{tabular}
  \caption{\small \it Nordea 15min $w_t$ autocorrelations}
  \label{tab:nordea_15min_w_acf}
\end{table}
we get
\begin{eqnarray*}
  \theta &=& 0.6890 \\
  \Theta &=& 0.6378
\end{eqnarray*}
Among the two roots of each of the 2nd order equations in the above,
we have chosen the one in the range $(-1, 1)$ so as to ensure
invertibility of the model \cite{BoxJenkins94}.

With an estimate of $\theta$ and $\Theta$, one can then infer the
noise process i.e. the residuals $y_t$:
\begin{equation}
  \label{eq:infer_y}
  y_t = w_t + \theta y_{t-1} + \Theta y_{t-s} - \theta \Theta y_{t-s-1}
\end{equation}
where we substitute $y_t\;(t \leq 0)$ with their unconditional
expectation 0.
% By inverting the seasonal moving average model of $w_t$ we can then
% infer the noise process $y_t$ \footnotemark, i.e. the residuals:
% \begin{eqnarray*}
%   y_t &=& {1 \over 1 - \theta B}{1 \over 1 - \Theta B^s} w_t \\
%   &=& \sum_{i=0}^{\infty}\sum_{j=0}^{\infty} \theta^i \Theta^j B^{i+j}
%   w_t \\
%   &=& w_t + \sum_{k=1}^{\infty} \sum_{j=0}^{\bot(k/s)} \theta^{k-sj} \Theta^j
%   w_{t-k}
% \end{eqnarray*}
% where $\bot(\cdot)$ denotes the floor function, i.e. the largest
% integer smaller than its argument.
% \footnotetext{
%   The autoregressive process resulting from the inversion is inifinite
%   in extent. Hence one must choose a cut-off value for k. For the
%   present case of Nordea 15min log-volatility, I choose the value
%   $k=sj$ for which $|\Theta|^j < 0.02$. This value of k, call it
%   $k_{\top}$, is around 550. To avoid data reduction of this amount,
%   we add to the beginning of $\{w_t\}_{t=1}^n$ $k_{\top}$ zeros
%   corresponding to t=0, -1, ..., -$k_{\top}$+1. Recall $\E(w_t) = \E(y_t
%   - \theta y_{t-1} - \Theta y_{t-s} + \theta \Theta y_{t-s-1}) = 0$.
% }

In order to forecst the $w_t$ process, and hence the return process
itself, we must also know the distribution of $y_t$. Moreover, to
properly estimate the parameters of the model in the sense of maximum
likelihood, we are also in need of the distribution of $y_t$.

If $y_t$ is indeed i.i.d random variables as we assumed, we only need
to work out its unconditional distribution. This is shown in figure
\ref{fig:nordea_15min_y_qq} and \ref{fig:nordea_15min_y_acf}.
\begin{figure}[htb!]
  \centering
  \subfigure[QQ-plot of $y_t$]{
    \includegraphics[scale=0.4, clip=true, trim=95 230 112
    210]{../pics/nordea_15min_y_qq.pdf}
    \label{fig:nordea_15min_y_qq}
  }
  \subfigure[ACF of $y_t$]{
    \includegraphics[scale=0.4, clip=true, trim=95 230 112
    210]{../pics/nordea_15min_y_acf.pdf}
    \label{fig:nordea_15min_y_acf}
  }
  \caption{\small \it Nordea 15min $y_t$ QQ-plot and autocorrelations}
\end{figure}
In addition, the measured moments of $y_t$ are shown in table
\ref{tab:nordea_15min_y_moments}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    mean & variance & skewness & kurtosis \\
    \hline
    0.0012 & 0.0935 & 0.2988 & 6.8691 \\
    \hline
  \end{tabular}
  \caption{\small \it Nordea 15min $y_t$ moments}
  \label{tab:nordea_15min_y_moments}
\end{table}

If we ignore the weak autocorrelations in this crude estimate of
$y_t$ and assume $y_t$ to be i.i.d random variables governed
by a skewed and leptokertic distribution, the parameters of this
distribution together with $\theta$ and $\Theta$ can be determined via
maximum likelihood estimate (MLE).

It turns out that $y_t$ can be well described by a Johnson Su
distribution \cite{Shang2004}:
\[
  y_t = \xi + \lambda\sinh{z_t - \gamma \over \delta}
\]
where $\gamma, \delta, \lambda, \xi$ are parameters to be determined
and $z_t \sim N(0, 1)$. The goodness of fitting is demonstrated in
figure \ref{fig:nordea_15min_y_js_fit} by the
empirical cummulative distribution function in comparison to the
theoretical one.
\begin{figure}[htb!]
  %\vspace{-18mm}
  \centering
    \includegraphics[scale=0.5, clip=true, trim=92 229 116
    133]{../pics/nordea_15min_y_js_fit.pdf}
    \caption{\small \it Nordea 15min $y_t$ fitted to a Johnson Su
      distribution. Horizontal: values of $y_t$, denoted x; Vertical:
      $\ln\left(P(y_t < x)\right)$.}
    \label{fig:nordea_15min_y_js_fit}
\end{figure}

The first 4 moments of the Johnson Su distribution are expressible
in closed form in $\gamma, \delta, \lambda, \xi$ \cite{Shang2004}:
\begin{eqnarray*}
  w &=& \exp{1 \over \delta^2} \\
  \Omega &=& {\gamma \over \delta} \\
  \text{E}(y) &=& -w^{1/2} \lambda \sinh\Omega + \xi\\
  \text{std}(y) &=& \lambda \left[{1 \over 2}(w-1)(w\cosh 2\Omega +
    1)\right]^{1/2} \\
  \text{skewness}(y) &=& {
    \sqrt{(1/2)w(w-1)} [w(w+2)\sinh 3\Omega + 3\sinh\Omega]
    \over
    (w\cosh 2\Omega + 1)^{3/2}} \\
  \text{kurtosis}(y) &=& {
    w^2(w^4 + 2w^3 + 3w^2 - 3)\cosh 4\Omega + 4w^2 (w+2) \cosh 2\Omega
    + 3(2w+1) \over
    2(w\cosh 2\Omega + 1)^2 }
\end{eqnarray*}
By matching the theoretical expressions of the moments with their
measured values, and taking help from published tables
\cite{Johnson1965}, one can solve for the parameters $\gamma, \delta,
\lambda, \xi$.

Under the assumption of i.i.d Johnson Su distributed residuals, the
log-likelihood function of the parameters $\gamma, \delta$
conditional on the sample $w_t$ can be written as
\[
L(\theta, \Theta) = -{1 \over 2}\sum_{t=1}^n z_t^2 + n \ln{\delta
  \over \lambda \sqrt{2\pi}} - {1 \over 2}\sum_{t=1}^n \ln\left[
  1 + \left({y_t - \xi \over \lambda}\right)^2
\right]
\]
where $y_t$ are inferred from $w_t$ using eq.\ref{eq:infer_y}
and $z_t$ from $y_t$ using
\[
z_t = \delta \sinh^{-1}{y - \xi \over \lambda} + \gamma
\]
Note that $\gamma, \delta, \lambda, \xi$ are not really free
parameters but rather are implied by $\theta$ and $\Theta$: Once the
latter have been chosen and the corresponding $y_t$ inferred, the
former are determined by the moments of $y_t$.

The eventual MLE is done in Matlab with the ``active set''
algorithm. The initial as well as the final estimation results are
listed in table \ref{tab:nordea_15min_js_param}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    & $\gamma$ & $\delta$ & $\lambda$ & $\xi$ & $\theta$ & $\Theta$ \\
    \hline
    initial estimate & 0.1476 & 1.5121 & 0.3633 & 0.0454 & 0.6890 &
    0.6378 \\
    \hline
    MLE estimate & 0.1319 & 1.5266 & 0.3735 & 0.0410 & 0.6639 & 0.6025
    \\
    \hline
  \end{tabular}
  \caption{\small \it Nordea 15min estimation results}
  \label{tab:nordea_15min_js_param}
\end{table}

To verify that the log-volatility can in general be described by
ARIMA(p, d, q) models, we apply the above procedure to a number of
Swedish stocks whose data have been collected by the author with
home-developed software. In \S\ref{sec:nordea} and \S\ref{sec:volvo}
we model the log-volatilities of Nordea Bank 15-minute returns and
Volvo B 30-minute returns. Both return series span the period
2013/10/10 - 2014/04/04. They are computed using minute-by-minute
average prices, and their variances, i.e. volatilities squared, are
computed as the sum of squared higher-frequency returns, which are in
turn computed using second-by-second prices. Whenever a needed average
price cannot be computed due to the lack of price observations in the
relevent time interval, the value of the previous average price is
assumed.

After modeling, we perform 1-step-ahead forecast using the estimated
models, and compare the predicted log-volatilities with the actual
data.

\subsection{Nordea Bank Log-Volatility}\label{sec:nordea}
In this section we model the log-volatility of {\it Nordea Bank}
15-minute returns during the period 2013/10/10 - 2014/04/04, which
contains 3879 data points. The squared volatilities, i.e. the
variances, are computed as the sum of squared 30-second returns. Among
the 3879 data points, we use the first 80\% for model estimation and
the last 20\%, i.e. 776 data points for forecast verification.

This Nordea series has a auto-correlation structure similar to its
predecessor that has been studied in previous sections. We show it in
figure \ref{fig:nordea2_w_acf}. The reader may want to compare it with
figure \ref{fig:nordea_15min_w_acf}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=95 230 112
  235]{../pics/nordea2_w_acf.pdf}
  \caption{\small \it Auto-correlation function of $w_t$}
  \label{fig:nordea2_w_acf}
\end{figure}
The significant auto-correlations and their respective lags are listed
in table \ref{tab:nordea2_w_acf}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    lag & 1 & 32 & 33 & 34 \\
    \hline
    ACF &   -0.4805 & 0.2861 &  -0.5034 & 0.2093 \\
    \hline
  \end{tabular}
  \caption{\small \it Significant auto-correlations of $w_t$}
  \label{tab:nordea2_w_acf}
\end{table}

Apparently this series also has the seasonality $s = 33$, but judging
from the relatively big differences between auto-correlations at lags
32 and 34, we don't expect the moving average process to be in the
multiplicative form of equation \ref{eq:nordea_w}. Instead, we model
$\ln\sigma_t$ has
\begin{eqnarray}
  (1 - B)(1 - B^s) \ln\sigma_t &=& (1 - \theta_1 B - \theta_s B^s -
  \theta_{s+1} B^{s+1} ) y_t \label{eq:nordea2_v}
\end{eqnarray}
The parameters are etimated to be
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    parameter & $\theta_1$ & $\theta_{33}$ & $\theta_{34}$ & $\var(y_t)$ \\
    \hline
    value & 0.8339 & 0.8448 & -0.7067 & 0.10071 \\
    \hline
  \end{tabular}
  \caption{\small \it: Parameter values of $\ln \sigma_t$}
\end{table}

The residuals $y_t$ do not have a Gaussian distribution, as confirmed
by the probability plot shown on the left of figure
\ref{fig:nordea2_y}. The non-normality, however, cannot be accounted
by the sine hypobolic transformation of Johnson Su, because of the
negative skewness of -1.2669 found in table
\ref{tab:nordea2_y_mmt}. How the non-normality can be accomodated is
quite another discussion and we do not pursue it here.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=20 220 30
  115]{../pics/nordea2_y.pdf}
  \caption{\small \it The residuals' distribution and auto-correlations. Left:
    cummulative distribution function (CDF) of the residuals $y_t$; Right:
    auto-correlation function of $y_t$.}
  \label{fig:nordea2_y}
\end{figure}

\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Moment & mean & variance & skewness & kurtosis \\
    \hline
    value & -0.0022 & 0.1007 & -1.2669 & 15.5965 \\
    \hline
  \end{tabular}
  \caption{\small \it: Moments of the residuals $y_t$}
  \label{tab:nordea2_y_mmt}
\end{table}

The auto-correlations of the residuals $y_t$ are insignificant, as
shown on the right of figure \ref{fig:nordea2_y}. Though the Ljung-Box Q
test rejects the null hypothesis of no-autocorrelation in the
residuals, the weak auto-correlations, if any at all,  are not worth
further complicating the model.

Forecast of $\ln\sigma_t$ is done by reverting equation
\ref{eq:nordea2_v} and write
\begin{eqnarray*}
  \ln\sigma_t &=& \ln\sigma_{t-1} + \ln\sigma_{t-s} -
  \ln\sigma_{t-s-1} + (1 - \theta_1 B - \theta_s B^s + \theta_{s+1}
  B^{s+1} ) y_t \\
  \E(\ln\sigma_t | \mathcal{F}_{t-1}) &=& \ln\sigma_{t-1} + \ln\sigma_{t-s} -
  \ln\sigma_{t-s-1} - \theta_1 y_{t-1} - \theta_s y_{t-s} +
  \theta_{s+1} y_{t-s-1}
\end{eqnarray*}

Thus the forecast, namely $\E(\ln\sigma_t | \mathcal{F}_{t-1})$, is
found in terms of previous values in the same series and of previous
residuals.

To compare the performance, we also estimate a GARCH model with
Gaussian conditional distribution for the Nordea Bank 15-minute
return. The procedure is standard and we only list the results here
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$ & $\alpha_5$ & $\beta_1$ \\
    \hline
    Value & $7.3686 \times 10^{-7}$ & 0.154131 & 0.0360 & 0.491832 \\
    \hline
  \end{tabular}
  \caption{\small \it Parameters of GARCH(1, 5) model of Nordea
    Bank 15-minute returns}
  \label{tab:nordea2_garch}
\end{table}

Forecasting using both models, we get two series of one-step-ahead
forecast, $\ln \sigma^{\text{ARIMA}}_t$ and $\ln \sigma^G_t$,
respectively. We denote the empirical log-volatilities as $\ln
\hat{\sigma}_t$. Then the accuracy of the two models can be compared
by comparing $\ln \sigma^{\text{ARIMA}}_t - \ln \hat{\sigma}_t$ and
$\ln \sigma^G_t - \ln \hat{\sigma}_t$. As a reference, we also include
in the comparison the naive approach of taking the sample mean of
$\ln \sigma_t$ as the forecast.

Table \ref{tab:slv_garch_comparison} shows the standard deviation of
$\ln \sigma^F_t - \ln \hat{\sigma}_t$ for the 3 kinds of
forecast. Here we use $\ln \sigma^F_t$ to denote the forecast of $\ln
\sigma_t$. Clear the ARIMA model has the smallest standard deviation,
the sample mean has the largest, and the GARCH model sits in between.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    & ARIMA & GARCH & Sample mean \\
    \hline
    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ &
    0.2876  &  0.3096 &   0.3349 \\
    \hline
  \end{tabular}
  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
    \ln\hat{\sigma}_t$}
  \label{tab:slv_garch_comparison}
\end{table}
Even more informative about the quality of forecast is the
distribution of $\ln \sigma^F_t - \ln \hat{\sigma}_t$, which we show
in figure \ref{fig:slv_garch_comparison}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=0 292 0
  255]{../pics/slv_garch_comparison.pdf}
  \caption{\small \it Comparison of the forecast from ARIMA, GARCH,
    and sample mean. Blue: ARIMA forecast; Green: GARCH forecast; Red:
    Sample mean forecast.}
  \label{fig:slv_garch_comparison}
\end{figure}
The left plot in the figure shows the situation of under-estimates,
i.e. $\ln \sigma^F_t - \ln \hat{\sigma}_t < 0$. We see that, to the
far left, the blue line is below the green, which is in turn below the
red, meaning the ARIMA forecast has the lowest probability of
under-estimate. The GARCH forecast comes second and the sample
mean forecast under-estimates most.

The same ranking of forecast quality is also seen on the right plot of
figure \ref{fig:slv_garch_comparison}, which shows the complementary
distribution of $\ln \sigma^F_t - \ln \hat{\sigma}_t$ for postive
values, i.e. over-estimates: The ARIMA forecast has the lowest
probability of over-estimate; GARCH comes second and the sample mean
is most likely to over-estimate.

As another measure of the quality of forecast, the fraction of
``good'' forecasts that differ no more than a given percentage from
the measured value is worth a few words. Table
\ref{tab:nordea_good_percentage} compares this fraction of good
forecasts from the 3 alternatives with respect to different
requirements of goodness.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    ARIMA & GARCH & Sample Mean \\
    \hline
    1\% & 0.1920 & 0.2010 & 0.1843 \\
    \hline
    5\% & 0.7887 & 0.7436 & 0.7229 \\
    \hline
    10\% & 0.9665 & 0.9601 & 0.9369 \\
    \hline
  \end{tabular}
  \caption{\small \it Fraction of ``good'' forecasts as defined by
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
  \label{tab:nordea_good_percentage}
\end{table}
Clearly, although ARIMA and GARCH rank differently with respect to
different requirements of goodness, they both out-perform the sample
mean forecast. So the efforts of building models and estimating them
have paid off.
\subsection{Volvo B Log-Volatility}\label{sec:volvo}
In this section we model the log-volatility of {\it Volvo B} 30-minute
returns during the period 2013/10/10 - 2014/04/04. This series
contains 1884 log-volatilities computed using 2-minute returns in each
30-minute interval. Among them we use the first 1507 for model
estimation and the last 377 for forecast and model verification.

The left plot of figure \ref{fig:volvo_inno_and_lv_acf} shows the
distribution of $(r_t - \E(r_t))/\sigma_t$. We observe in the figure a
nice Gaussian variate, so we can be sure that the sum of squared
2-minute returns makes a good approximation to the variance of
30-minute returns in this particular case.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=0 256 0
    151]{../pics/volvo_inno_and_lv_acf.pdf}
    \label{fig:volvo_inno_and_lv_acf}
  \caption{\small \it Left: probability plot of $(r_t -
    \E(r_t))/\sigma_t$; Right: auto-correlations of $\ln \sigma_t$}
\end{figure}

Guided by the auto-correlations of $\ln\sigma_t$ shown in the right
plot of figure \ref{fig:volvo_inno_and_lv_acf} we find the following
model:
\begin{eqnarray}
  && (1 - B)(1 - B^s) \ln \sigma_t \nonumber \\
  &=& (1 - \theta_1 B - \theta_2B^2 -
  \theta_3B^3 - \theta_4B^4) (1 - \Theta B^s)
  y_t \label{eq:volvo_lv_model}
\end{eqnarray}
where $s = 16$ is the seasonality and is apparent from the
auto-correlations of $\ln\sigma_t$. Fitting this model to the
actual log-volatilities yields the parameter values listed in table
\ref{tab:volvo_params}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Parameter & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ &
    $\Theta$ & residual variance \\
    \hline
    Value & 0.7305 & 0.0575 & 0.0574 & 0.0346 & 0.8324 & 0.1340\\
    \hline
  \end{tabular}
  \caption{\small \it Volvo B log-volatility parameters}
  \label{tab:volvo_params}
\end{table}
Forecasting using the estimated model parameters gives the forecast
series $\ln \sigma^{\text{ARIMA}}_t$. To access the quality of the
forecast, we also estimate a GARCH model using the returns. The result
is a GARCH(1, 1) model:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$  & $\beta_1$ \\
    \hline
    Value & $3.125 \times 10^{-7}$ & 0.05 & 0.90 \\
    \hline
  \end{tabular}
  \caption{\small \it GARCH(1, 1) model of Volvo B 30-minute returns}
  \label{tab:volvo_garch}
\end{table}

The forecasts from ARIMA, GARCH, and the sample mean are compared
using the difference $\ln \sigma^F_t - \ln\hat{\sigma}_t$, where $\ln
\sigma^F_t$ stands for the forecast. The distributions of this
difference is plotted in figure \ref{fig:volvo_slv_garch_cmp}, and
the standard deviation of the distributions are listed in table
\ref{tab:volvo_slv_garch_cmp}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    & ARIMA & GARCH & Sample mean \\
    \hline
    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.3261 &
    0.4250 & 0.3708 \\
    \hline
  \end{tabular}
  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
    \ln\hat{\sigma}_t$}
  \label{tab:volvo_slv_garch_cmp}
\end{table}

\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=10 282 0
    243]{../pics/volvo_slv_garch_cmp.pdf}
    \label{fig:volvo_slv_garch_cmp}
  \caption{\small \it Comparison of the forecast from ARIMA, GARCH,
    and sample mean. Blue: ARIMA forecast; Green: GARCH forecast; Red:
    Sample mean forecast.}
\end{figure}
Figure \ref{fig:volvo_slv_garch_cmp} shows, as in the previous case of
Nordea Bank 15-minute returns, the ARIMA model performs the best, GARCH
the second, and the sample mean the worst. However, when it comes to
over-estimates, the sample mean appears to be the best estimator,
while ARIMA excels over GARCH. But a check of $\E(\ln\sigma_t)$ over
the sample for estimation (1507 data points) and over the sample for
comparison (377 data points) reveals that the first sample has mean
-6.3127 while the second has -6.2072. This increment explains the low
probability of over-estimation when using the first sample mean as
forecast.

Table \ref{tab:volvo_good_percentage} compares the fraction of
``good'' estimates as measured by ${|\ln\sigma^F_t -
  \ln\hat{\sigma}_t| \over |\ln \hat{\sigma}_t|}$ being less than 1\%,
5\% and 10\%.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    ARIMA & GARCH & Sample Mean \\
    \hline
    1\% & 0.2175 & 0.1220 & 0.1406 \\
    \hline
    5\% & 0.7162 & 0.6154 & 0.6631 \\
    \hline
    10\% & 0.9231 & 0.8806 & 0.8992 \\
    \hline
  \end{tabular}
  \caption{\small \it Fraction of ``good'' forecasts as defined by
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
  \label{tab:volvo_good_percentage}
\end{table}
We see from the table that the ARIMA model consistently excels over
the other two alternatives. Once again, it confirms the validity and
good performance of an ARIMA log-volatility model.

% \subsection{Ericsson B Log-Volatility}\label{sec:ericsson}
% The probability density function of the residuals $y_t$ that result
% from the MLE estimate is shown in figure
% \ref{fig:nordea_15min_y1_hist} together with its counterpart from the
% initial estimate.
% \subsection{45min Returns}
\section{Unconditional Distribution Functions of ARIMA Log-Volatility
  Models}\label{sec:XieCalc}
In this section we study the unconditional distribution function of
the Stochastic Log-Volatility model specified as equation
\ref{eq:SLV_spec}. As is discussed in \S\ref{sec:SLV_model},
$\ln\sigma_t$ can be described by an ARMA or ARIMA \footnote{ARIMA ---
  Integrated Autoregressive Moving Average} model, possibly with
seasonal components. Here we note that all these models can be
re-written as a moving average model, which is infinite in extent if
autoregressive components are present:
\begin{eqnarray*}
  \ln \sigma_t &=& y_t + \sum_{n=1}^\infty c_n y_{t-n} + \text{Const.}
\end{eqnarray*}
Since the $y_t$ are independent and identically distributed,
\begin{eqnarray*}
  y_t + \sum_{n=1}^\infty c_n y_{t-n}  
\end{eqnarray*}
has Gaussian distribution by the central limit theorem, on condition
that $y_t$ for all $t$ have finite second moment --- this is what we
assume in the rest of this section. It follows from the above equation
that the unconditional distribution of $\ln \sigma_t$ is the same as the
distribution of $\bar{v} + v$ where $v \sim N(0, \sigma)$ and
$\bar{v}$, $\sigma$ are constants. Now we can state that the
unconditional distribution of the returns $r_t$
\begin{eqnarray*}
  r_t &=& \mu + \sigma_t b_t\\
  &=& \mu + \exp\left(
    y_t + \sum_{n=1}^\infty c_n y_{t-n} + \text{Const.}
  \right)
\end{eqnarray*}
 is the same as
\begin{equation}  \label{eq:UnconditionalPdf}
  \begin{aligned}
    r &= \mu + e^{\bar{v} + v} b \\
  \end{aligned}
\end{equation}
where $b \sim N(0, 1)$. For convenience, let $r' = e^v b$

In \S\ref{sec:SLV_Symmetric} we first study the model in the
relatively simple case when $v$ and $b$ are uncorrelated and $\mu
= 0$. If this simplified version proves inadequate, one may resort to
the general model studied in \S\ref{sec:SLV_Asymmetric}.

\subsection{The Simplified model}\label{sec:SLV_Symmetric}
In the following we derive the unconditional PDF of $r'$, denoted
$f_{r'}(x)$. Then the PDF of $r$ is $e^{-\bar{v}}f_{r'}(e^{-\bar{v}}x)$.
\begin{eqnarray*}
  P(r' < x) &=& P(b < xe^{-v}) \\
  f_{r'}(x) &=& f_b(xe^{-v}) e^{-v}
\end{eqnarray*}
Averaging over all $v$, we get
\begin{equation}\label{eq:UncondPDFSymmetric}
  \begin{aligned}
    f_{r'}(x) =& \int_{-\infty}^{\infty} dv (2\pi\sigma^2)^{-1/2}
    e^{-v^2/2\sigma^2}(2\pi)^{-1/2} \exp(-x^2e^{-2v}/2) e^{-v} \\
    =& {1 \over 2\pi\sigma} \int_{-\infty}^{\infty} dv
    \exp\left(
      -{1 \over 2\sigma^2} v^2 - v -{1 \over 2} x^2 e^{-2v}
    \right)
  \end{aligned}
  \end{equation}
The last part of the integrand, $e^{-x^2 e^{-2v} / 2}$, is plotted in
figure \ref{fig:DoubleExp}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=85 252 100
  231]{../pics/DoubleExp.pdf}
  \caption{\small \it Plot of $\exp(-{1 \over 2} x^2 e^{-2v})$}
  \label{fig:DoubleExp}
\end{figure}
Therefore we make the following approximation:
\[
\exp\left(-{1 \over 2} x^2 e^{-2v}\right) \approx \left\{
  \begin{array}{lr}
    0 & \text{if } v < \ln|x| -{1 \over 2} \ln(2\ln 2) \\
    1 & \text{otherwise}
  \end{array}
\right.
\]
Here we note that $\exp\left(-{1 \over 2} x^2 e^{-2v}\right) = 1/2$ at
$v = \ln|x| -{1 \over 2} \ln(2\ln 2)$.

With this approximation we have
\begin{eqnarray*}
  f_{r'}(x) &=& {1\over C}{1 \over 2\pi\sigma} \int_{\ln(|x|/\sqrt{\ln
      4})}^{\infty} dv
  \exp\left(-{1 \over 2\sigma^2} v^2 - v\right) \\
  &=& {1\over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}} \text{erfc} \left(
    {1 \over \sqrt{2}\sigma} \ln{|x| \over \sqrt{\ln 4}} + {\sigma
      \over \sqrt{2}}
  \right)
\end{eqnarray*}
where $1/C$ has been added for the purpose of normalization.

At large $x$, we may use the asymptotic expansion of $\mathrm{erfc}$
to write
\begin{eqnarray*}
  Cf_{r'}(x) &=& {e^{\sigma^2 / 2} \over \sqrt{8\pi}} \mathrm{erfc}(\xi) \\
  &=& {e^{\sigma^2 / 2} \over \sqrt{8 \pi}}
  \frac{e^{-\xi^2}}{\xi\sqrt{\pi}}\left[
    1 +
    \sum_{n=1}^N (-1)^n \frac{(2n-1)!!}{(2\xi^2)^n} \right] +
  O(\xi^{-2N-1} e^{-\xi^2})
\end{eqnarray*}
where $\xi = {1 \over \sqrt{2}\sigma} \ln{|x| \over \sqrt{\ln 4}} +
{\sigma \over \sqrt{2}}$. The slowest-decaying term is
\[
f_{r,0}(x) = {e^{\sigma^2 / 2} \over \sqrt{8\pi}}
\frac{e^{-\xi^2}}{\xi\sqrt{\pi}}
\]
Let $\zeta = {|x| \over \sqrt{\ln 4}}$. With a bit manipulation one
obtains
\begin{equation*}
  f_{r,0}(x) = {1 \over \pi \sqrt{8}}{1 \over
    \left(\ln\zeta/\sigma\sqrt{2} +
      \sigma/\sqrt{2}\right)\zeta\zeta^{\ln \zeta / 2\sigma^2}
  }
\end{equation*}
From the last equation one can see that, at any neighborhood of large
$x$, $f_{r'}(x)$ may be approximated by $C/|x|^\alpha$, i.e. a power law.

In the following we work out the normalization constant C.
\begin{eqnarray*}
  \int_{-\infty}^{\infty} f_{r'}(x) dx &=& {1 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{-\infty}^{\infty} \text{erfc} \left({1 \over
      \sqrt{2}\sigma} \ln{|x| \over \sqrt{\ln 4}} + {\sigma \over
      \sqrt{2}} \right) dx\\
  &=& {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{0}^{\infty} \text{erfc} \left({1 \over
      \sqrt{2}\sigma} \ln{x \over \sqrt{\ln 4}} + {\sigma \over
      \sqrt{2}} \right) dx\\
\end{eqnarray*}
Let
\begin{eqnarray*}
  a &=& {1 \over \sigma \sqrt 2} \\
  b &=& -{1 \over 2 \sigma \sqrt 2}\ln\ln 4 + {\sigma \over \sqrt 2} \\
  y &=& a\ln|x|+b \\
\end{eqnarray*}
Then
\begin{eqnarray*}
  \int_{-\infty}^{\infty} f_{r'}(x) dx &=& {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{-\infty}^{\infty} dy {e^{(y-b)/a} \over a}
  \text{erfc}(y) \\
  &=& {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \left.e^{(y-b)/a}
    \text{erfc}(y)\right|_{y=-\infty}^{\infty}
  + {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{-\infty}^{\infty} dy {2 \over \sqrt \pi}
  e^{(y-b)/a} e^{-y^2} \\
  &=& {2 \over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}}  2 e^{\ln\ln 4/2
    - \sigma^2/2}\\
  C &=& \sqrt{2 \ln 4\over \pi} \\
  &\approx& 0.9394
\end{eqnarray*}
Thus we can write in summary:
\begin{eqnarray*}
  f_r(x) &=& e^{-\bar{v}} f_{r'}(e^{-\bar{v}} x) \\
    &=& {e^{-\bar{v}} \over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}}
    \text{erfc} \left({1 \over \sqrt{2}\sigma} \ln{|e^{-\bar{v}}x| \over \sqrt{\ln
          4}} + {\sigma \over \sqrt{2}}
    \right) \\
\end{eqnarray*}
Using the same technique for integration as for normalization, the
cummulative distribution function of $r$ is found to be $F(x)$,
which is the following:
\begin{enumerate}
\item if $x < 0$
  \begin{eqnarray*}
    F(x) &=& {\sqrt{\ln 4} \over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}}
    \left[
      {e^{-\bar{v}}x \over \sqrt{\ln 4}}\text{erfc}\left(
        {1 \over \sigma \sqrt 2} \ln{-e^{-\bar{v}}x \over \sqrt{\ln
            4}} + {\sigma \over \sqrt 2}
      \right) \right.\\
      && \left. + e^{-\sigma^2 / 2} \text{erfc}\left(
        {1 \over \sigma \sqrt 2} \ln{-e^{-\bar{v}}x \over \sqrt{\ln 4}}
      \right)
    \right]
  \end{eqnarray*}
\item if $x \geq 0$
  \begin{eqnarray*}
    F(x) &=& \frac{1}{2} + {\sqrt{\ln 4} \over C}{e^{\sigma^2 / 2} \over
      \sqrt{8\pi}} \left[
      {e^{-\bar{v}}x \over \sqrt{\ln 4}} \text{erfc}\left(
        {1 \over \sigma \sqrt 2} \ln{e^{-\bar{v}}x \over \sqrt{\ln
            4}} + {\sigma \over \sqrt 2}
      \right) \right. \\
      && \left. + e^{-\sigma^2/2} \text{erfc}\left(
        -{1 \over \sigma \sqrt 2} \ln{e^{-\bar{v}}x \over \sqrt{\ln 4}}
      \right)
    \right]
  \end{eqnarray*}
\end{enumerate}

To verify the validity of the model, we fit the above probability
density function to the de-meaned 30min returns of Volvo B
\footnote{By ``de-meaned returns'' we mean the quantity $r_t - 
  \mean{r_t}$, where $r_t$ are the measured returns and $\mean{r_t}$
  is the sample mean. The data set covers the transaction records of
  Volvo B on the OMX market (Stockholm) between 2013-10-10 and
  2014-03-12. The returns are computed using 1-minute mean prices.}
by means of maximum likelihood estimate using the MATLAB function
``mle''. Then for the parameters $\sigma$ and $\bar{v}$ we get
\begin{eqnarray*}
  \sigma &=& 0.6355 \\
  \bar{v} &=& -6.0308
\end{eqnarray*}
Then we plot $P(r' > x)$ of the model against its empirical
counterpart on a log-log scale, as shown in figure
\ref{fig:volvo_30min_ret}.
\begin{figure}[htb!]
  % \vspace{-15mm}
  \begin{center}
    \includegraphics[scale=0.4, clip=true, trim=98 231 116
    126]{../pics/volvo_30min_ret.pdf}
  \end{center}
  %\vspace{-5mm}
  \caption{\small \it{$P(r' > x)$ for $x > 0$. Blue: empirical
      probabilities. Red: probabilities predicted by the model.}}
  \label{fig:volvo_30min_ret}
\end{figure}

In general figure \ref{fig:volvo_30min_ret} shows a good fit, but a
closer look reveals that deviations are significant in the regions $r
\in (0, \sigma_r)$ and $r \in (2\sigma_r, 3\sigma_r)$, where
$\sigma_r$ stands for the empirical standard deviation of the
de-meaned returns. These observations are
shown in greater details in figure \ref{fig:volvo_30min_ret2}.
\begin{figure}[htb!]
  \centering
  \subfigure[$\ln(P(r > x))$ with $x \in (0, \sigma_r)$]{
    \includegraphics[scale=0.4, clip=true, trim=93 224 115
    125]{../pics/volvo_30min_ret_0-sigma.pdf}
  }
  \subfigure[$\ln(P(r > x))$ with $x \in (2\sigma_r, 3\sigma_r)$]{
    \includegraphics[scale=0.4, clip=true, trim=93 224 115
    125]{../pics/volvo_30min_ret_2sigma-3sigma.pdf}
  }
  \caption{\small \it Surviving probabilities. Blue: empirical values of $\ln(P(r
    > x))$. Red: Predicted values of $\ln(P(r > x))$.}
  \label{fig:volvo_30min_ret2}
\end{figure}

Moreover, the inefficiency of the model is also manifest in the
skewness of the data. For the Volvo 30min returns, the data has a
skewness of 0.2419, but our model is strictly symmetric, since
$x$ appears in equation \ref{eq:UncondPDFSymmetric} only as $x^2$.
Hence the above model needs to be improved to accommodate the non-zero
skewness as well as to account for the discrepancies shown in figure
\ref{fig:volvo_30min_ret2}. This is the subject of the next section.

\subsection{The General model}\label{sec:SLV_Asymmetric}
It has long been hypothesized in the liturature that skewness is the
result of price-volatility correlation (for example
\cite{Potters2003}). Therefore, the most apparent modification is to
allow $v$ and $b$ to be correlated. For convenience we decompose
$v \sim N(0, \sigma)$ as $v = \sigma a$ and assume
\begin{eqnarray*}
  \begin{pmatrix}
    a \\
    b
  \end{pmatrix} \sim N(0, \Sigma)
\end{eqnarray*}
with a covariance matrix
\begin{eqnarray*}
    \Sigma &=&
  \begin{pmatrix}
    1 & \psi \\
    \psi & 1
  \end{pmatrix}
\end{eqnarray*}
where $|\psi| < 1$. Then we rewrite equation \ref{eq:UnconditionalPdf}
as
\begin{equation}
  \label{eq:r_t}
  \begin{aligned}
    r &= \mu + e^{\bar{v}} r' \\
    r' &= e^{\sigma a} b \\
  \end{aligned}
\end{equation}
Then
\begin{eqnarray*}
  && P(r' < x)\\
  &=& P(b < e^{-a\sigma}x) \\
  &=& {1 \over 2\pi |\det(\Sigma)|^{1/2}}
  \int_{-\infty}^{\infty} da \int_{-\infty}^{e^{-a\sigma}x} db
  \exp\left[
    -{1 \over 2} (a, b) \Sigma^{-1}
    \begin{pmatrix}
      a \\
      b
    \end{pmatrix}
  \right] \\
\end{eqnarray*}
After some manipulations we get
\begin{equation}\label{eq:UncondCDFAsymmetric}
  \begin{aligned}
    & P(r' < x) \\  
    &= {1 \over 2\pi \sqrt{1 - \psi^2}} \int_{-\infty}^{\infty} da
    e^{-a^2/2} \int_{-\infty}^{e^{-a\sigma}x} db
    \exp\left[
      - {(b - a\psi)^2 \over 2(1 - \psi^2)}
    \right] \\
    &= {1 \over 2\sqrt{2\pi}} \int_{-\infty}^{\infty} e^{-a^2/2}
    \text{erfc}{a\psi - e^{-a\sigma} x \over \sqrt{2(1-\psi^2)}} da
  \end{aligned}
\end{equation}
Differentiating with respect to $x$ yields
\begin{eqnarray}
  f_{r'}(x) &=& {1 \over 2\pi \sqrt{1 - \psi^2}} \times \nonumber \\
  && \int_{-\infty}^\infty da \exp
  \left[- {
      a^2 + 2\sigma(1 - \psi^2)a - 2a\psi e^{-a\sigma} x + e^{-2a\sigma} x^2
      \over
      2(1 - \psi^2)
  }
  \right] \label{eq:UncondPDFAsymmetric} \\
  f_r(x) &=&  \td{r'}{r} f_{r'}[e^{-\bar{v}} (x - \mu)] \nonumber \\
  &=& e^{-\bar{v}} f_{r'}[e^{-\bar{v}}
  (x-\mu)] \label{eq:UncondPDFAsymmetric1}
\end{eqnarray}
Unlike \ref{eq:UncondPDFSymmetric} where a relatively simple
approximation can be found and thus leads to an analytic result of the
integral, no such approximation has been found by the
author for the integral \ref{eq:UncondPDFAsymmetric}. However, the
moment generating functions (MGF) of PDF \ref{eq:UncondPDFAsymmetric}
and \ref{eq:UncondPDFAsymmetric1} are easy enough to find and give the
moments about the origin in closed form. Then by matching the analytic
expressions of the moments with their statistical values from the
sample, the parameters $\sigma$, $\psi$ and $\bar{v}$ corresponding to the
sample can be obtained.

The MGF of \ref{eq:UncondPDFAsymmetric1} can be found as follows:
\begin{eqnarray*}
  && M_{r}(t) \\
  &=& \E(e^{tr}) \\
  &=& {1 \over 2\pi \sqrt{1 - \psi^2}} \int_{-\infty}^{\infty} da \exp \left(
    -{a^2 \over 2} - \sigma a
  \right) \int_{-\infty}^{\infty} dx \exp\left[
    t(e^{\bar{v}}x + \mu) - {
      (a\psi - e^{-a \sigma} x)^2 \over 2 (1 - \psi^2)
    }
  \right] \\
  &=& {1 \over \sqrt{2 \pi}} \int_{-\infty}^{\infty} da \exp\left[
    -{a^2 \over 2} + {1 \over 2} e^{2 a \sigma + 2\bar{v}} (1 - \psi^2) t^2 +
    a \psi e^{a \sigma + \bar{v}} t + \mu t
  \right]
\end{eqnarray*}
With the MGF, the first 4 moments of $r$ can be computed:
\begin{equation}
  \label{eq:AsymmetricMoments}
  \begin{aligned}
    \E(r) &= \mu +e^{\bar{v}+\frac{\sigma ^2}{2}} \sigma  \psi\\
    \E(r^2) &= \mu ^2+2 e^{\bar{v}+\frac{\sigma ^2}{2}} \mu  \sigma  \psi
    +e^{2 \left(\bar{v}+\sigma ^2\right)} \left(1+4 \sigma ^2 \psi^2\right)\\
    \E(r^3) &= \mu ^3+3 e^{\bar{v}+\frac{\sigma ^2}{2}} \mu ^2 \sigma \psi
    +9 e^{3 \bar{v}+\frac{9 \sigma ^2}{2}} \sigma \psi \left(1+3 \sigma ^2
      \psi ^2\right)
    +3e^{2 \left(\bar{v}+\sigma ^2\right)} \mu  \left(1+4 \sigma ^2 \psi
      ^2\right) \\
    \E(r^4) &= \mu ^4+4 e^{\bar{v}+\frac{\sigma ^2}{2}} \mu ^3 \sigma  \psi
    +36 e^{3 \bar{v}+\frac{9 \sigma ^2}{2}} \mu  \sigma  \psi  \left(1+3
      \sigma ^2 \psi ^2\right)\\ 
    & +6 e^{2 \left(\bar{v}+\sigma ^2\right)} \mu ^2 \left(1+4 \sigma ^2 \psi
      ^2\right)
    +e^{4 \bar{v}+8 \sigma ^2} \left(3+96 \sigma ^2 \psi ^2+256 \sigma ^4
      \psi ^4\right)
  \end{aligned}
\end{equation}
These equations are rather complicated and directly solving them to
obtain the parameters $\mu$, $\sigma$, $\psi$ and $\bar{v}$ is
infeasible. However, in practice, $\E(r)$ is often very small --- so, to
get a rough estimate of the parameters, we may set $\mu = 0$ in the
above equations and, with a bit of manipulation, find the following
equation for $\psi\sigma$:
\begin{equation}
  \label{eq:moment1}
  {\E(r^3) \E(r^4)^{-3/8} \over \E(r^2)^{3/4}} =
  {9\sigma \psi (1 + 3\sigma^2\psi^2) \over (1 +
    4\sigma^2\psi^2)^{3/4}}
  {1
    \over
    (3 + 96\sigma^2\psi^2 + 256\sigma^4 \psi^4)^{3/8}
  }
\end{equation}
This equation can be solved numerically for a given sample to yield an
estimate for $\sigma\psi$, which in turn can be substituted in
equations \ref{eq:AsymmetricMoments}, where $\mu$ has been set to 0,
to give estimates for all the 4 parameters. These estimates then serve
as initial values in a numerical solution to
\ref{eq:AsymmetricMoments} where $\mu$ is kept as a free
variable. This full solution can now be used as the initial estimate
in an MLE procedure.

It is particularly important in our situation to have a good initial
estimate, because, as has been shown earlier, the PDF and CDF cannot
be obtained in closed forms and consequently have to be evaluated by
numerical integration. This is a rather costly procedure especially in
the context of MLE. So the computation of the moments under the
assumption of $\mu = 0$ is worthwhile.

Following the aforementioned procedure, i.e. computing the initial
estimate by matching the moments and then refining the estimate by
MLE, we obtain the parameter values for a number of return series,
including the Volvo B 30-minute returns described in
\S\ref{sec:SLV_Symmetric}. In the following, we list the parameter
values and show the comparison between the empirical distribution
functions and their counterparts from the model.

Clearly these figures show a fairly good match of the empirical
distribution functions to their model counterparts. So we are once
again assured of the validity of ARIMA log-volatility models. However,
we also see that the skewness of the returns of Volvo B and Ericsson B
have rather different values when computed from the sample and from
the model. This could be the consequence of the limited sample size or
deficiencies in the estimation procedure described above. We leave
these issues to future studies.

%  This indicates the parameter $\psi$, which is the
% correlation between the log-volatility and the return, may be more
% complicated than the simple constant that has been assumed so far. We
% leave this to later studies.

\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    & $\psi$ & $\sigma$ & $\bar{v}$ & $\mu$ \\
    \hline
    Volvo B & -1.5747e-02 & 4.6929e-01 & -6.2304e+00 &
    -5.3040e-05 \\
    Nordea Bank & 1.7234e-02 & 4.2780e-01 & -6.3758e+00 & 7.5230e-05
    \\
    Ericsson B & 1.8486e-02 & 4.4905e-01 & -6.2834e+00 & -9.2817e-05 \\
    \hline
  \end{tabular}
  \caption{\small \it Parameter Values of Selected Assets'
    returns. The time series are 30-minute returns and run from
    2013/10/10 to 2014/03/12.}
  \label{tab:assets_params}
\end{table}

\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    & mean & std & skewness & kurtosis \\
    \hline
    \multirow{2}{*}{Volvo B} & -8.6766e-05 & 2.4500e-03 &
    -2.7217e-02 & 7.3394e+00 \\
    & -6.9281e-05 & 2.4538e-03 & -7.2677e-02 & 7.2472e+00 \\
    \hline
    \multirow{2}{*}{Nordea Bank} & 7.4821e-05 & 2.0500e-03 &
    6.9433e-02 & 6.7427e+00 \\
    & 8.8983e-05 & 2.0443e-03 & 6.7125e-02 & 6.2446e+00 \\
    \hline
    \multirow{2}{*}{Ericsson B} & -8.2606e-05 & 2.2939e-03 &
    2.6586e-01 & 7.6716e+00 \\
    & -7.5675e-05 & 2.2844e-03 & 7.8573e-02 & 6.7299e+00 \\
    \hline
  \end{tabular}
  \caption{\small \it Moments of Selected Assets' returns. For each return
    series, the 1st row contains the sample moments, while the 2nd
    constains those computed with MLE parameters. The time series are
    30-minute returns and run from 2013/10/10 to 2014/03/12.}
  \label{tab:assets_moments}
\end{table}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=50 233 63
  136]{../pics/Volvo_B_30m_returns.pdf}
  \caption{\small \it Volvo B 30min returns' unconditional distribution fit to
    the Stochastic Volatility model. The time series runs from
    2013/10/10 to 2014/03/12. Left: $P(r < x)$ with $x < 0$; Right:
    $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
  \label{fig:Volvo_B_30m_returns}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=65 241 72
  143]{../pics/Nordea_30m_returns.pdf}
  \caption{\small \it Nordea Bank 30min returns' unconditional
    distribution fit to the Stochastic Volatility model. The time
    series runs from 2013/10/10 to 2014/03/12. Left: $P(r < x)$ with
    $x < 0$; Right: $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
  \label{fig:Nordea_30m_returns}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=21 241 28
  146]{../pics/Ericsson_30m_returns.pdf}
  \caption{\small \it Ericsson B 30min returns' unconditional
    distribution fit to the Stochastic Volatility model. The time
    series runs from 2013/10/10 to 2014/03/12. Left: $P(r < x)$ with
    $x < 0$; Right: $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
  \label{fig:Ericsson_30m_returns}
\end{figure}

\subsection{Relation to Conditional Distribution Functions}
In the last section we have shown that the unconditional distribution of
the returns are skewed, which implies the variate in the
log-volatility $a$ is correlated to the variate in the return
$b$.

In the context of conditional distributions and forecast, this
correlation translates to the correlation between the residual of the
log-volatility, denoted $y_t$ in \S\ref{sec:nordea} and
\S\ref{sec:volvo}, and $b_t$. Now let us consider the forecast
function of an ARIMA model:
\begin{eqnarray*}
  \ln \sigma_t &=& y_t + \sum_{i=1}^P \phi_i \ln \sigma_{t-i} -
  \sum_{i=1}^Q \theta_i y_{t-i}
\end{eqnarray*}
where $\phi_i \leq 1$ if the model involves integration, and $\phi_i <
1$ otherwise. Comparing this equation with equation \ref{eq:r_t}, one
immediately realizes the PDF of $r_t$
\begin{eqnarray*}
  r_t &=& \mu + \sigma_t b_t
\end{eqnarray*}
is given by equation \ref{eq:UncondPDFAsymmetric1} and its moments
given by the equations \ref{eq:AsymmetricMoments} if one makes the
following substitutions:
\begin{eqnarray*}
  \sigma &\to& \text{std}(y_t) \\
  \bar{v} &\to& \sum_{i=1}^P \phi_i \ln \sigma_{t-i} - \sum_{i=1}^Q
  \theta_i y_{t-i} \\
  \psi &\to& \text{corr}(y_t, b_t)
\end{eqnarray*}

For the Nordea series considered in \S\ref{sec:nordea},
$\text{corr}(y_t, b_t)$ is found to be $3.56 \times 10^{-2}$; while
for the Volvo series considered in \S\ref{sec:volvo},
$\text{corr}(y_t, b_t)$ is found to be $-7.3 \times 10^{-3}$. These
values are comparable to those in table \ref{tab:assets_params}, where
$\psi$ for the Nordea series is $1.7234 \times 10^{-2}$ and $-1.5747
\times 10^{-2}$ for the Volvo series. The similarity in these values
provides some evidence about the validity of the model.

It is certainly impossible to compare a predicted conditional
distribution with observations, but conditional distributions are of
great interest in the context of risk management and derivative
pricing, so we point out how they may be calculated for ARIMA
log-volatility models.