### 2014-05-05 thesis/_region_.tex
This is the version to be used for job searching.

\message{ !name(NordeaVolvo.tex)}
\message{ !name(NordeaVolvo.tex) !offset(9) }
$r_t / \sigma_t$ is approximately Gaussian, confirming the choice of
50-second returns for the computation of the realized volatility.

\message{ !name(NordeaVolvo.tex) !offset(-6) }