### 2014-06-11 thesis/cross_correlation_gaussian.tex
Move matlab files to a separate folder.

\chapter{Covariance Matrix of Gaussian Returns}
\label{chp:Gaussian}
In this chapter we present some results about the covariance
matrix of Gaussian returns. The assumption of Gaussian returns is of
course an over-simplification of reallity, but nevertheless lends some
insight into the problem of elements' and eigenvalues' distributions of
a covariance matrix.
Our primary interest is in the influence that auto-correlations in
returns exert on the covariance matrix. Section \ref{sec:GCC-analytical} and
section \ref{sec:GCC-numerical} discuss how the distributions of the matrix
elements and the eigenvalues are affected by such autocorrelations,
respectively.

\section{Distribution of the Matrix Elements}
\label{sec:GCC-analytical}
In this section we study the distribution of the elements of a
covariance matrix $C = \mtx{RR'}/T$, where
\begin{eqnarray*}
  \mtx R &=&
  \begin{pmatrix}
    r_{11} & r_{12} & \cdots & r_{1T} \\
    r_{21} & r_{22} & \cdots & r_{2T} \\
    \vdots & \vdots & \ddots & \vdots \\
    r_{N1} & r_{N22} & \cdots & r_{NT} \\
  \end{pmatrix}
\end{eqnarray*}
In words, an element $r_{it}$ of $\mtx R$ is the return of asset $i$
at time $t$, with $i=1,\cdots, N$ and $t = 1, \cdots, T$. If each
column of $\mtx R$ follows a zero-mean Gaussian distribution, i.e.
\begin{eqnarray*}
  \begin{pmatrix}
    r_{1t} \\
    \vdots \\
    r_{Nt}
  \end{pmatrix} \sim N(0, \mtx \Sigma)
\end{eqnarray*}
for all $t = 1, \cdots, T$, and none of the return series is
auto-correlated, i.e. $\text{corr}(r_{it}, r_{i,t'}) = 0$ for all $i =
1, \cdots, N$ and $t \neq t'$, then $\mtx{RR'}$ is a Wishart matrix whose
probability density function is well known \cite{Anderson2003}.

When auto-correlations are indeed present in the returns, $\mtx{RR'}$
no longer follows the Wishart distribution. However, the joint distribution
function of its elements can be expressed in terms of the Wishart \gls{pdf}
and the auto-correlations. We show this in appendix
\ref{app:pdf_gaussian1}. Also, in appendix
\ref{chp:gaussian_elements_dist} we derive an approximate expression
for the asymptotic distribution of these matrix elements, assuming
$r_{it}$ is an \gls{ar}(1) process \footnotemark, i.e. $r_{it} =
\phi r_{i, t-1} + a_{it}$, and
\footnotetext{
  It is straight forward to derive the auto-correlation function
  $\varrho_k$ of an \gls{ar}(1) process with autoregressive
  coefficient $\phi$:
  \begin{eqnarray*}
    \varrho_k &=& \text{corr}(r_{it}, r_{i, t-k}) \\
    &=& \phi \varrho_{k-1} \\
    &=& \phi^k
  \end{eqnarray*}
  Let $\tau$ denote the time lag at which $\varrho_{\tau} = 1/2$, then
  it follows $\tau = -\ln 2/\ln \phi$.
}
\begin{eqnarray*}
  \begin{pmatrix}
    a_{1t} \\
    \vdots \\
    a_{Nt}
  \end{pmatrix} \sim N(0, \mtx \Sigma)
\end{eqnarray*}
where
\begin{eqnarray*}
  \mtx \Sigma &=& \sigma^2
  \begin{pmatrix}
    1 & \rho & \cdots & \rho \\
    \rho & 1 & \cdots & \rho \\
    \vdots & \vdots & \ddots & \vdots \\
    \rho & \rho & \cdots & 1
  \end{pmatrix}
\end{eqnarray*}
where $-1 < \rho < 1$ is a constant parameter describing the correlation
between $a_{it}$ and $a_{jt}$ --- so, by construction, we assume such
correlations are constant across all asset pairs and over all
time.

The elements $C_{ij}$ ($i \neq j$) of the covariance matrix, $C
= \mtx{RR'}/T$, are found to be normally distributed with mean
\begin{eqnarray}
  \mu'_X &=& {\sigma^2 \over \sqrt{2\pi} (1 - \phi^2)(1 -
    \rho^2)^{1/4}} \left[ P^{-3/2}_{-1/2}(-\rho) -
    P^{-3/2}_{-1/2}(\rho)
  \right] \label{eq:gaussian_mean2}
\end{eqnarray}
and variance
\begin{eqnarray}
  \sigma'^2_X &=& {1 \over (1 - \phi^2)^2}\left[
    \sum_{t=1}^T \sum_{k=1}^{t-1} 2\left(
      \phi^k \over T
    \right)^2 \sigma^6 + \sum_{t=1}^T
    {\sigma^4 (1 - \rho^2)^2 \over T^2} v^2(\rho)
  \right] \nonumber \\
  &=& {2 \sigma^6 \over T (1 - \phi^2)^2} \left[
    {\phi^2 \over 1 - \phi^2} -
    {\phi^2 (1 - \phi^{2T}) \over
      T(1 - \phi^2)}
  \right] + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
    T (1 - \phi^2)^2} \nonumber \\
  &\approx& {2 \sigma^6 \phi^2 \over T (1 - \phi^2)^3}
  + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
    T (1 - \phi^2)^2} \label{eq:gaussian_variance2}
\end{eqnarray}
where $P^\mu_\nu(\cdot)$ is Ferrer's function of the first
kind. See appendix \ref{chp:gaussian_elements_dist}. For the
definition of Ferrer's function, see equation
\ref{eq:Ferrers_1st}.

Equation \ref{eq:gaussian_mean2} tells that, if
two return series $i$ and $j$ are not correlated, i.e. $\rho = 0$,
auto-correlation in the returns does not introduce a bias into the
estimation of their covariance, i.e. $\mu'_X$, since the difference
between the two Ferrer's functions evaluate to 0 in equation
\ref{eq:gaussian_mean2}; if, however, the return series are indeed
correlated, auto-correlation in the returns rescales the
covariance through a multiplicative factor $1/(1 - \phi^2)$.

In addition, equation \ref{eq:gaussian_variance2} tells that
auto-correlation in the returns always makes the covariance
estimation more noisy --- auto-correlation not only rescales the
variance of the no-autocorrelation estimation by $1/(1 - \phi^2)^2$
but even adds an extra term ${2 \sigma^6 \phi^2 \over T (1 -
  \phi^2)^3}$.

For the diagonal elements of the covariance matrix $C$ we have
\begin{eqnarray}
  \E(C_{ii}) &=& {1 \over T}\left[
    \sum_{k=0}^{t-1} \phi^{2k} \sigma^2
  \right] \nonumber \\
  &=& {\sigma^2 \over (1 - \phi^2) T} \left[
    T - {\phi^2 (1 - \phi^{2T}) \over 1 - \phi^2}
  \right] \nonumber \\
  & \approx & {\sigma^2 \over 1 - \phi^2} \left[
    1 - {\phi^2 \over T (1 - \phi^2)}
  \right] \label{eq:gaussian_cii_mean2}
\end{eqnarray}
and
\begin{eqnarray}
  \var(C_{ii}) &=& \sum_{t=1}^T \left[
    \sum_{k=0}^{t-1} {\phi^{4k} \sigma^4 \over T^2} 2 +
    \sum_{k,l=0}^{t-1} {\phi^{2(k+l)} \over T^2} \sigma^6
  \right] \nonumber \\
  &=& \sum_{t=1}^T \left[
    {2 \sigma^4 \over T^2} {1 - \phi^{4t} \over 1 - \phi^4} +
    {\sigma^6 \over T^2} \left(
      {1 - \phi^{2t} \over 1- \phi^2}
    \right)^2 \right] \nonumber \\
  &=& {2 \sigma^4 \over T (1 - \phi^4)} -
  {2 \sigma^4 \phi^4 (1 - \phi^{4T}) \over T^2(1 - \phi^4)^2} +
  \nonumber \\
  && {\sigma^6 \over T (1 -\phi^2)^2} -
  {2 \sigma^6 \phi^2 (1 - \phi^{2T}) \over T^2 (1 - \phi^2)^3} +
  {\sigma^6 \phi^4 (1 - \phi^{4T}) \over T^2 (1 - \phi^2)^2 (1 -
    \phi^4)} \nonumber \\
  &\approx& {2 \sigma^4 \over T (1 - \phi^4)} + {\sigma^6 \over T (1
    -\phi^2)^2} \label{eq:gaussian_cii_variance2}
\end{eqnarray}

From equation \ref{eq:gaussian_cii_mean2} we see that auto-correlation
in the returns increases the variance of the return series; and from
equation \ref{eq:gaussian_cii_variance2} we see that the variance of
that variance estimation is also increased by
auto-correlations. Moreover, we note that $\var(C_{ii})$ scales with T
approximately as $1/T$, similar to the behavior of
$\var(C_{ij})$. This is to be compared with the case of \gls{garch}
returns discussed in chapter \ref{chp:CrossCorrelationFat}.

\section{Distribution of the Eigenvalues}
\label{sec:GCC-numerical}
For a Wishart matrix $\mtx{RR'}$, theoretical results are available
for the eigenvalue distribution, and we summarize them in appendix
\ref{sec:wishart_eigen_dist}. In short, the joint probability density
function of the eigenvalues is given by equation
\ref{eq:wishart_eigen_pdf} when neither auto-correlation nor
cross-correlation is present in $\mtx R$. Moreover, the mariginal
distribution of the largest eigenvalue is approximated by a gamma
distribution \cite{Chiani2012}.

However, as detailed in the derivation leading to equation
\ref{eq:cross-corr-matrix-PDF}, the distribution of $\mtx{RR'}$ is not
Wishart when the columns of $\mtx{R}$ are correlated
(auto-correlation). Deriving the eigenvalue distribution analytically
in this case is beyond the scope of this thesis. Instead, we resort to
numerical methods.

As before we consider the \gls{ar}(1) process:
\begin{equation*}
  \vec{r}_t = \phi \vec{r}_{t-1} + \vec{a}_{t}
\end{equation*}
where $\vec{a}_t \sim N(0, \mtx{I})$, i.e. the elements of $\vec{a}_t$ are
independent Gaussian random variable with zero mean and unit
variance. Now we investigate how the eigenvalue distribution depends on
the auto-correlation strength parameter $\phi$. Figure
\ref{fig:GaussianMarkovSpectrumPDF} shows the results of the
simulation.
\begin{figure}[htb!]
  \begin{center}
    \includegraphics[scale=0.4, clip=true, trim=90 228 115
    226]{../pics/GaussianMarkovSpectrumPDF.pdf}
  \end{center}
  \vspace{-10mm}
  \caption{\small \it
      Eigenvalue distribution with correlation time $\tau$ ranging
      from 0 to 3. The 1st blue line, which is shown as stairs, is the
      theoretical eigenvalue distribution according to the
      Marcenko-Pastur law (see eq.\ref{eq:MP_pdf}). In the simulation
      we have chosen $q = N/T = 50/1000 = 0.05$ and the standard
      deviation of the returns $\sigma=1$. For each value of $\tau$ we
      generate 2000 instances of $N \times T$ random matrix $R$, and
      compute C as $C=RR'/T$. Hence each curve in the figure is
      constructed from 2000 sets of eigenvalues. The correspondance
      between the correlation time $\tau$ and the auto-regressive
      coefficient $\phi$ is $\tau = -\ln 2 / \ln \phi$.
    }
  \label{fig:GaussianMarkovSpectrumPDF}
\end{figure}
It is clear from the figure that the maximum eigenvalue moves
consistently to the right as the value of $\phi$ increases, and, as
shown in figure \ref{fig:Gaussian_mineig}, the minimum eigenvalue
also increases with $\phi$.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=100 226 116
  133]{../pics/Gaussian_mineig.pdf}
  \caption{\small \it The minimum eigenvalue versus
    auto-correlation strength $\phi$. For each value of $\phi$ 2000
    random matrices are generated and their eigenvalues are
    calculated. The minimum eigenvalue of each random matrix is noted
    and the mean of the 2000 such minimum eigenvalues are plotted
    against the chosen value of $\phi$. 20 values of $\phi$ are
    included in the plot, ranging from 0 to 0.95 with step size 0.05.}
  \label{fig:Gaussian_mineig}
\end{figure}

Chiani showed that the marginal distribution of the maximum eigenvalue
($\lambda_1$) is approximately gamma when neither cross-correlation
nor auto-correlation is present\cite{Chiani2012}. So we compare
in figure \ref{fig:GaussianMarkov005MaxEigCDF_loglog} the empirical
cumulative distribution function (CDF) of the maximum eigenvalue 
with the CDF of a gamma distribution. The cases where
auto-correlations are present ($\phi > 0$) have also been included.
\begin{figure}
  \begin{center}
    \includegraphics[scale=0.5, clip=true, trim=0 238 0
    197]{../pics/GaussianMarkov005MaxEigCDF_loglog.pdf}
  \end{center}
  \caption{\small \it Cummulative distribution function (CDF) of the
    maximum eigenvalue ($\lambda_1$) from numerical simulations (blue
    lines) are compared to the fitted gamma distribution (red
    lines). Each pair of CDFs correspond to a fixed autocorrelation
    strength ($\phi$). The parameters $k$ and $\theta$ of the gamma
    distribution are fit to data by matching the 2nd and the 3rd
    moments of the gamma distribution to the corresponding moments of
    the empirical distribution. Then the parameter $\alpha$ in
    equation \ref{eq:TracyWidom-Gamma} is chosen to be $\alpha =
    k\theta - \E\left({\lambda_1 - \mu_{NT} \over
        \sigma_{NT}}\right)$. The curves are plotted on log-log
    scale.}
  \label{fig:GaussianMarkov005MaxEigCDF_loglog}
  %\vspace{-10mm}
\end{figure}
It is seen in the figure that gamma distributions with different
parameters fit fairly well in all cases. So we conclude that a
gamma distribution not only approximates the maximum eigenvalue
distribution at the absence of autocorrelations but does so even at
the {\it presence} of autocorrelations.

Since the maximum eigenvalue distribution is approximated by a gamma
distribution characterized by parameters $k$, $\theta$, and $\alpha$
\footnotemark, the influence of the autocorrelations can be
characterized by the dependence of $k$, $\theta$, and  $\alpha$ on
$\phi$. While these dependences are rather intricate, good support can
be found in the data for the following approximate relation:
\footnotetext{
  The mean, variance and skewness of the gamma distribution are given
  by
  \begin{eqnarray*}
    \text{mean} &=& k\theta \\
    \text{variance} &=& k\theta^2 \\
    \text{skewness} &=& 2/\sqrt{k}
  \end{eqnarray*}
}
\begin{eqnarray}
  k\theta &=& a \tan^2{\pi \phi \over 2} + b\tan{\pi \phi \over 2} +
  c \label{eq:k_theta-phi}
\end{eqnarray}
Here we note that $k\theta$ is the mean of the gamma distribution.
To verify this relation, we first fit a 2nd order polynomial and
obtain the coefficients $a$, $b$, $c$; then for each data point
$k_n\theta_n$ we solve the quadratic equation
\begin{eqnarray}
  a \tan^2{\pi \phi'_n \over 2} + b\tan{\pi \phi'_n \over 2} + c -
  k_n\theta_n &=& 0\label{eq:k_theta-phi_2}
\end{eqnarray}
for $\tan{\pi \phi'_n \over 2}$. If relation \ref{eq:k_theta-phi} is a
good approximation, a close match between $\tan{\pi \phi'_n \over 2}$
and $\tan{\pi \phi_n \over 2}$ is expected. From figure 
\ref{fig:phi_and_roots} one can see this is indeed the case.

\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=37 217 39
    170]{../pics/phi_and_roots.pdf}
  \caption{\small \it Upper plot: $\tan{\pi \phi'_n \over 2}$ against
    $\tan{\pi \phi_n \over 2}$. The fitted line has equation $y_n =
    0.995 \tan{\pi \phi_n \over 2} + 0.0146$. Lower plot: Residuals of
    the linear fit, i.e. $\tan{\pi \phi'_n \over 2} - y_n$. 20 values
    of $\phi$ are included in the plot, ranging from 0 to 0.95 with
    step size 0.05.}
  \label{fig:phi_and_roots}
\end{figure}

% Figure \ref{fig:GaussianMarkovMaxEig_k-phi} shows how the mean of the
% fitted gamma distribution, i.e. $k\theta$, varies as the
% autocorrelation strengthens ($\phi$ increases). 
% \begin{figure}[htb!]
%   %\vspace{-10mm}
%   \centering
%     \includegraphics[scale=0.5, clip=true, trim=100 223 112
%     141]{../pics/GaussianMarkov05MaxEig_k-phi.pdf}
%     \caption{\small \it The parameter $k$ against autocorrelation strength
%       $\phi$. Blue crosses: Empirical values of $k$. Red: best fitting
%       line in terms of {\it Least Square Errors}.}
%   \label{fig:GaussianMarkovMaxEig_k-phi}
% \end{figure}
% From the equation of the fitting line we can directly read out
% \begin{eqnarray*}
%   k &=& a\phi + b \\
%   &=& a\left(1 \over 2\right)^{1/\tau} + b
% \end{eqnarray*}
% where $a = -26$ and $b = 47$.

% For the parameter $\theta$, its behavior is more conveniently
% described in terms of the correlation time $\tau$. Figure
% \ref{fig:GaussianMarkovMaxEig_theta-tau} plots the values of $\theta$
% against those of $\tau$ together with a fitting quadratic
% function. Higher order polynomials provide a slightly better
% fit, but coefficients of the 3rd order and above are less than 1/1000
% times the coefficients of the 2nd and the 1st order. Therefore a 2nd
% order polynomial has been chosen.
% \begin{figure}[htb!]
%   \vspace{-15mm}
%   \centering
%   \includegraphics[scale=0.5, clip=true, trim=99 230 114
%   139]{../pics/GaussianMarkov05MaxEig_theta-tau.pdf} 
%   \caption{\small \it Vertical axis: $\theta$; Horizontal axis:
%     correlation time $\tau$. Blue: empirical values of $\theta$. Cyan:
%     Fitting quadratic function.}
%   \label{fig:GaussianMarkovMaxEig_theta-tau}
% \end{figure}
% The equation of this polynomial is
% \begin{eqnarray*}
%   \theta &=& A\tau^2 + B\tau + C
% \end{eqnarray*}
% where $A = 2.1\times 10^{-4}$, $B = -1.1\times 10^{-4}$, $C =
% 3.2\times 10^{-4}$.

% The parameter $\alpha$ shifts the gamma distribution $\mathscr{G}(k,
% \theta)$ to match the mean of ${(\lambda_1 -
%   \mu_{NT})/\sigma_{NT}}$. The behavior of $\alpha$ with respect to
% changing autocorrelation is shown in figure
% \ref{fig:GaussianMarkovMaxEig_alpha-tau}.
% \begin{figure}[htb!]
%   \vspace{-15mm}
%   \centering
%   \includegraphics[scale=0.5, clip=true, trim=104 229 114
%   139]{../pics/GaussianMarkovMaxEig_alpha-tau.pdf}
%   \caption{\small \it The mean-shift parameter $\alpha$ against
%     correlation time $\tau$. Blue: empirical values of $\alpha$. Cyan:
%     quadratic fit.}
%   \label{fig:GaussianMarkovMaxEig_alpha-tau}
% \end{figure}
% The equation of $\alpha$ is then inferred from the fitting line:
% \begin{equation*}
%   \alpha = D\tau^2 + E\tau + F
% \end{equation*}
% where $D = -3.4\times 10^{-3}$, $E = -4.6\times 10^{-2}$ and $F = 69$.

% So combining the results of $\alpha$, $k$ and $\theta$ we can express
% the moments of the Tracy-Widom variate ${(\lambda_1 -
%   \mu_{NT})/\sigma_{NT}}$:
% \begin{equation}\label{eq:lambda1_MeanVariance}
%   \begin{aligned}
%     \E({\lambda_1 - \mu_{NT} \over \sigma_{NT}}) &= k\theta -
%     \alpha \\
%     &= \left[-a\left(1 \over 2\right)^{1/\tau} + b\right](A\tau^2 +
%     B\tau - C) - (D\tau^2 + E\tau + F) \\
%     \var({\lambda_1 - \mu_{NT} \over \sigma_{NT}}) &=
%     k\theta^2 \\
%     &= \left[-a\left(1 \over 2\right)^{1/\tau} + b\right](A\tau^2 +
%     B\tau - C)^2 \\
%   \end{aligned}
% \end{equation}
% Figure \ref{fig:GaussianMarkovMaxEig_tw_moments} shows the empirical
% moments of ${(\lambda_1 - \mu_{NT})/\sigma_{NT}}$ together with the
% corresponding values computed using the above formulas.
% \begin{figure}[htb!]
%   \vspace{-15mm}
%   \centering
%   \includegraphics[scale=0.46, clip=true, trim=48 243 6
%   134]{../pics/GaussianMarkovMaxEig_tw_moments.pdf}
%   \caption{\small \it empirical moments of ${(\lambda_1 -
%       \mu_{NT})/\sigma_{NT}}$ against their theoretical
%     counterparts. Left: Empirical/theoretical mean against $\tau$;
%     Right: Empirical/theoretical variance against $\tau$. Blue
%     crosses: empirical values; Red line: fitting curves. Horizontal
%     axis: correlation time $\tau$.}
% \label{fig:GaussianMarkovMaxEig_tw_moments}
% \end{figure}
% The good fitness shown in figure
% \ref{fig:GaussianMarkovMaxEig_tw_moments} allows us to conclude 
% that, within the range of the correlation time that we have studied,
% namely $\tau \in [0, 13.51]$, the $k$ parameter is a linear function
% of $\phi = 2^{-1/\tau}$ while $\theta$ and $\alpha$ are quadratic
% functions of $\tau$. The moments of the transformed maximum
% eigenvalue, namely ${(\lambda_1 - \mu_{NT})/\sigma_{NT}}$, are thus
% expressed as functions of the correlation time $\tau$ via the
% parameters $k$, $\theta$ and $\alpha$.

% Figure \ref{fig:GaussianMarkovMaxEigPDF_original} shows the PDF of the
% maximum eigenvalue ($\lambda_1$) for a range of values of the
% correlation time $\tau$. One can clearly see that the mean of the
% distribution moves to the right and the width of the distribution
% increases as $\tau$ takes on larger and larger values. However, one
% must not forget that the behavior of the mean and the width is random
% in nature. Eq. \ref{eq:lambda1_MeanVariance} describes such behavior in
% the asymptotic limit, i.e. if an infinite number of random matrices
% are generated and their respective maximum eigenvalues computed for
% each and every value of correlation time $\tau$.
% \begin{figure}[htb!]
%   \begin{center}
%     \includegraphics[scale=0.5, clip=true, trim=98 228 112
%     221]{../pics/GaussianMarkovMaxEigPDF_original.pdf}
%     \caption{\small \it Probability density function of the maximum eigenvalue
%       ($\lambda_1$).}
%   \end{center}
%   \label{fig:GaussianMarkovMaxEigPDF_original}
% \end{figure}