### 2014-06-11 thesis/thesis.tex

\documentclass{report}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{amssymb}
% \usepackage{amsfonts, amsmath}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\usepackage[acronym]{glossaries}
\usepackage{bookmark}

\input{../physics_common.tex}
\input{./acronyms.tex}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

% \title{Stocks' Cross-Correlations}
% \subtitle{}
\makeglossaries
\author{Xie Xiaolei}
\date{\today}
\begin{document}

\begin{titlepage}
\begin{center}

% Upper part of the page. The '~' is needed because \\
% only works if a paragraph has started.
\includegraphics[width=0.5\textwidth]{../pics/lund_uni-logo_s}~\\[1cm]

% \textsc{\LARGE Lund University}\\[1.5cm]

\textsc{\Large Master's Thesis Project}\\[0.5cm]

% Title
%\HRule \\[0.4cm]
{ \huge \bfseries Return Models and Covariance Matrices
  \\[0.4cm] }

%\HRule \\[1.5cm]

% Author and supervisor
\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
\textsc{Xie} Xiaolei
\end{flushleft}
\end{minipage}
\begin{minipage}{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
Prof. Sven \textsc{\AA berg}
\end{flushright}
\end{minipage}

\vfill

% Bottom of the page
{\large \today}

\end{center}
\end{titlepage}
% \maketitle
\begin{abstract}
Return models and covariance matrices of return series have been
studied. In particular, \gls{garch} and \gls{sv} models
are compared with respect to their forecasting accuracy when applied
to intraday return series. \gls{sv} models are found to be considerably more
accurate and more consistent in accuracy in forecasting.

Covariance matrices formed from Gaussian and \gls{garch} return series, and
in particular, return series auto-correlated as an AR(1) process, have
been studied. In the case of Gaussian returns, the largest eigenvalue
is found to approximately follow a gamma distribution also when the
returns are auto-correlated. Expressions relating the mean and the
variance of the asymptotic Gaussian distribution of the matrix
elements are derived. In the case of \gls{garch} returns, both the largest
and the smallest eigenvalues of the covariance matrix are seen to
increase with increasing auto-correlation. The matrix elements are
found to follow L\'evy distributions with different L\'evy indices for
the diagonal and the non-diagonal elements.

Localization of eigenvectors of covariance matrices of
returns from \gls{garch} processes has been investigated. It is
found that the localization is reduced as the auto-correlation is
increased. Quantitatively, the number of localized eigenvectors
decreases approximately as a quadratic function with the
auto-correlation strength, i.e. the autoregressive coefficient of the
AR(1) process.
\end{abstract}

\section*{Acknowledgements}
First of all, I thank my supervisor, Prof. Sven \AA berg, for the many
hours of mentoring, discussion, encouragement and guidance. Without
them, this report will not exist. Then I wish to give my gratitude to
the people in the division of mathematical physics. The casual
discussions that I had with them truely inspired me, and the friendly
and lively atmosphere that they create has helped bring peace and clarity
to my mind.

Last but not least, I want to thank my wife, Wang Xiaodan, for her
unconditional support and understanding.

\tableofcontents

% \chapter*{List of Acronyms}

\printglossaries

\chapter{Introduction}
The mathematics of the financial market has always been a topic that
arouses interest and imagination, and with no doubt, has been
studied from many aspects and in many different ways.

Central to all these studies are the concepts of probability
distributions and correlations. The values of stocks, futures,
options, etc. are stochastic in nature and are governed by the laws of
stochastic processes, which are expressed in terms of probability
distributions and correlations.

Meanwhile, the observables of the market are prices, volumes,
turn-over (the amount of money paid in a trade), names of the brokers,
and time of the trades. These quantities don't make much sense
by themselves but do reveal the probabilistic dynamics of the market
when put together and turned into statistics.

The dynamics of an asset is affected by its own history as well as by
the histories and current values of other assets in the market. The
influence from the asset's own history is termed autocorrelations,
i.e. correlations in time, while the influence from other assets are
termed cross correlations.

\Gls{sv} models built on historical data describe the time evolution
of the aforementioned observables in terms of probability
distributions, autocorrelations and cross correlations. They may have
some predictive power and help determine a fair price of a given
asset.

Quite often, instead of an asset's price, the relative price change,
i.e. the return, is studied. Analytically solvable models often assume
Gaussian return distribution, although data suggest fatter tails. In
numerical return models, realistic reproduction of historical data may
be achieved.

% In the domain of discrete-time returns' models,
% \gls{garch} and stochastic volatility models (SV) are the most popular, so in
% this thesis project we study a total of 7 intraday return series 
% using \gls{garch} and SV models and compare their performance in volatility
% forecast.
In addition to the model of a single asset, the correlation between
a group of assets is also of great interest. For example, in principle
component analysis, one wishes to identify a number of factors that
``drive'' the price evolution of a group of assets in the sense that
each of the assets' returns can be expressed as a linear combination
of the factors' returns. In this scenario, the eigenvalues of the
covariance matrix are the variances of the factors' returns and their
corresponding orthogonalized eigenvectors give the composition of the
factors, i.e. the coefficients with which the factors are contructed
as a linear combination of the assets.

Therefore, in this thesis we also study the elements and the
eigenvalues distribution, as well as the eigenvectors composition of
the covariance matrix. When the matrix is constructed from returns
with simple Gaussian distribution, the matrix is termed a Wishart matrix
and has been studied extensively in the literature. If the returns
have L\'evy distributions, the matrix is termed Wishart-L\'evy and has
been studied to some extent, particularly regarding its eigenvalue
distribution \cite{politi2010}.

However, it is understood that real stock/index returns are much more
complicated than a straight-forward Gaussian or L\'evy distribution can
describe --- instead, one needs structured models. For this reason, we
are particularly interested in a covariance matrix obtained for
realistic return models. The so called \gls{garch}(1,1) model is a realistic
return model proven to have regularly varying tails
\cite{mikosch2000}. So we study properties of eigenvalues and
eigenvectors of such covariance matrices. Moreover, we also study how
auto-correlations in the returns influence the aforementioned
properties. Such auto-correlations, known as second-order
auto-correlations decay exponentially but may still leave footprints
in the covariance matrix.

This report is organized as follows: Chapter \ref{chp:PriceModels}
reviews some of the most influential return model. Parameters of the
models are fitted to a few intraday return series and the predictive
power of the models is compared. A calculation of the
unconditional distribution functions of \gls{sv} models,
especially in the case where the residual of the log-volatility
and the innovation of the return are correlated normal variates, is
also presented. In chapter \ref{chp:Gaussian} we investigate
distributions of eigenvalues of the Wishart matrix, and study the
influence of auto-correlated returns. In chapter
\ref{chp:CrossCorrelationFat} distributions of elements and
eigenvalues of the covariance matrix of identically specified
\gls{garch}(1,1) series are studied. Finally, chapter \ref{chp:summary}
summarizes the results. Supplementary materials are provided in the
appendices.

\input{price_models}
\input{cross_correlation_gaussian}

\chapter{Covariance Matrix of GARCH(1,1) Returns}
\label{chp:CrossCorrelationFat}
\gls{garch} models and particularly \gls{garch}(1,1) models are widely used to
model financial return series of various time scales, ranging from
daily and monthly returns that have been studied extensively in the
literature to intraday returns that we have selectively investigated
in chapter \ref{chp:PriceModels}. One particularly nice feature of
\gls{garch}(1,1) models is that they have regularly varying tails (power-law
tails) \cite{Mikosch2009, mikosch2000} even when the innovations
(denoted $z_t$ in the following text) are normally distributed ---
something not shared by other classes of models (e.g. not by
\gls{sv} models, as shown in section \ref{sec:XieCalc}) but well
documented for realistic returns data by empirical studies
\cite{Mantegna2000, Potters2003, Guhr2007}. However, what this tail
behavior implies for the covariance matrix is much less understood,
especially when the return series of the covariance matrix are
auto-correlated.

So in this chapter we consider the covariance matrix of N
identically specified, possibly auto-correlated \gls{garch}(1, 1)
processes:
\begin{eqnarray}
  r_{it} &=& \phi r_{i, t-1} + \epsilon_{it} \nonumber \\
  \epsilon_{it} &=& \sigma_{it} z_{it} \label{eq:garch_spec}
\end{eqnarray}
where $i=1,2,...,N$; $t=1,2,...,T$; $z_{it}$ is independent,
identically distributed, and
\begin{eqnarray*}
  \sigma_{it}^2 &=& \alpha_0 + \alpha_1 z_{i, t-1}^2 + \beta_1
  \sigma_{i,t-1}^2
\end{eqnarray*}
Mikosch and Starica showed in \cite{mikosch2000} that a
\gls{garch}(1,1) process satisfying
\begin{eqnarray*}
  \alpha_0 &>& 0 \\
  \E\ln(\alpha_1 Z^2 + \beta_1) &<& 0 \\
\end{eqnarray*}
and
\begin{eqnarray*}
  \E[(\alpha_1 Z^2 + \beta_1)^{p/2}] &\geq& 1 \\
  \E|Z|^p \ln|Z| &\leq& \infty
\end{eqnarray*}
for some $p > 0$, is stationary and has regularly varying tails. The
tail exponent $\alpha$ is determined by:
\begin{equation}\label{eq:garch_alpha}
  \E[(\alpha_1 Z^2 + \beta_1)^{\alpha/2}] = 1
\end{equation}
Here $Z$ is a random variable that has the same distribution as
$z_{it}$. In our simulations described hereafter, $z_{it}$ and $Z$
have standard Gaussian distribution. In this section and the next, we
first study situations where no auto-correlations are present among
the returns, i.e. $\phi = 0$; then in section \ref{sec:garch_nonzero_phi} we
look at how auto-correlations change the picture.

With regularly varying tails, the eigenvalue distribution of a
covariance matrix built from \gls{garch}(1,1) returns is expected to
differ from the Marcenko-Pastur law discussed in section
\ref{sec:wishart_eigen_dist}. In figure \ref{fig:garch_ev_cii} we
simulate N=50 independent \gls{garch}(1,1) 
returns series, each with identical parameters, namely $\alpha_0 =
2.3\times 10^{-6}$, $\alpha_1 = 0.15$, $\beta_1 = 0.84$, $\phi = 0$
and T=$8\times10^4$ time steps, then we build the covariance
matrix as
\begin{eqnarray*}
  \mtx C &=& {1 \over T^{2/\alpha}} \mtx{RR'}
\end{eqnarray*}
where $\mtx R$ is an $N\times T$ matrix, whose elements $r_{it}$ are
specified by equation \ref{eq:garch_spec} with $\phi = 0$. The
normalization factor $1 \over T^{2/\alpha}$ has been chosen such that the
eigenvalue distribution is independent of $T$ in the limit $T \to
\infty$ \cite{politi2010, Cizeau1994}.

The \gls{pdf} of the eigenvalue distribution of C is plotted in figure
\ref{fig:garch_ev_cii_linear} and \gls{cdf} of the distribution is
plotted on log-log scale in \ref{fig:garch_ev_cii_loglog}.
\begin{figure}[htb!]
  \centering
  \subfigure[]{
    \includegraphics[scale=0.40, clip=true, trim=104 271 111
    216]{../pics/garch_ev_cii.pdf}
    \label{fig:garch_ev_cii_linear}
  }
  \subfigure[]{
    \includegraphics[scale=0.40, clip=true, trim=104 274 104
    226]{../pics/garch_ev_cii_loglog.pdf}
    \label{fig:garch_ev_cii_loglog}
  }
  \caption{\small \it \ref{fig:garch_ev_cii_linear}: Eigenvalues and
    Diagonal elements' distribution of a covariance matrix
    built from independent GARCH return series. Blue: PDF of
    eigenvalues; Green: PDF of diagonal elements; Red: PDF of a
    $\alpha$-stable distribution fitted to the diagonal
    elements. \ref{fig:garch_ev_cii_loglog}: CDF of the same
    quantities in 10-based log-log scale.}
  \label{fig:garch_ev_cii}
\end{figure}
Also plotted in the same figure is the distribution of the diagonal
elements of $\mtx C$. It is clear from the figure that the two PDFs
coincide, implying $\mtx C$ is diagonal. This is further confirmed by
figure \ref{fig:garch_cij} which shows the distribution of the
non-diagonal elements of C. One can see the non-diagonal elements are
distributed symmetrically around 0 with a very small width in
comparison to the distribution of the diagonal elements --- in fact, 1
order of magnitude smaller ($2.14\times10^{-5}$ v.s. $7.31\times
10^{-4}$). Hence $\mtx C$ is very close to a diagonal matrix.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=95 253 91
    223]{../pics/garch_cij.pdf}
  \caption{\small \it Distribution of the non-diagonal elements of
    the covariance matrix. Blue: PDF of the non-diagonal
    elements; Green: $\alpha$-stable distribution fitted to the
    non-diagonal elements' PDF.}
  \label{fig:garch_cij}
\end{figure}

Figure \ref{fig:garch_ev_cii} also shows the two curves are well fitted
by an $\alpha$-stable distribution. An estimate of the L\'evy index
$\alpha$ of the stable distribution is also obtained via fitting,
$\alpha \approx 1.38$, as shown in figure \ref{fig:garch_ev_cii}. This
is really an expected result for the diagonal elements. Using
$\alpha_1 = 0.15$, $\beta_1 = 0.84$, which are the values used for
simulating the \gls{garch} returns, one can obtain, by solving equation
\ref{eq:garch_alpha}, $\alpha=2.96$. Then according to Mikosch and Starica
\cite{mikosch2000}
\begin{eqnarray*}
  P(|r_t| > x) &\sim& {\E(z^\alpha) c_0 \over x^\alpha} \\
  P(|r_t|^2 > x) &\sim& {\E(z^\alpha) c_0 \over x^{\alpha/2}} \\
\end{eqnarray*}
for some constant $c_0$. Now that $P(|r_t|^2 > x)$ has power-law tail
behavior with power $\alpha/2 < 2$, one can deduce
\begin{equation}
  \label{eq:stable_CLT}
  \begin{aligned}
    \sum_{t=1}^T r_t^2 &\xrightarrow{d} S(\alpha/2,
    1, \gamma, \mu) \text{ as $T \to \infty$}
  \end{aligned}
\end{equation}
where $\xrightarrow{d}$ denotes convergence in distribution, and
$S(\alpha/2, 1, \gamma, \mu)$ denotes an $\alpha$-stable distribution
with parameters $(\alpha/2, 1, \gamma, \mu)$. Here $\alpha/2$ is the
L\'evy index, 1 is the asymmetry, $\gamma$ is the scaling parameter
and $\mu$ is the mean value of the distribution. Asymmetry being 1
means a random variable so distributed only takes positive values
\cite{Bilik2008, Embrechts1997}.

The mean $\mu$ in equation \ref{eq:stable_CLT} is given by
\begin{eqnarray*}
  \mu &=& T \E(|r_t|^2) \\
  &=& T{\alpha_0 \over 1 - \alpha_1 - \beta_1}
\end{eqnarray*}
where we have used the result $\E(|r_t|^2) = \alpha_0 / (1 - \alpha_1
- \beta_1)$ \cite{Bollerslev86}. The scaling parameter $\gamma$ in
\ref{eq:stable_CLT} is determined by the limit \cite{Bilik2008}
\begin{eqnarray*}
  \lim_{T\to\infty} {T \E(z^{\alpha/2}) c_0 \over \gamma^{\alpha/2}}
  &=& C_{\alpha/2}
\end{eqnarray*}
where
\begin{eqnarray*}
  C_{\alpha/2} &=& \left( \int_0^\infty {\sin x \over x^{\alpha/2}} dx
  \right)^{-1} \\
  &\approx& {1 \over \sqrt{2 \pi}}
\end{eqnarray*}
Therefore
\begin{eqnarray*}
  \gamma^{\alpha/2} &=& \sqrt{2\pi} T \E\left(
    |z|^{\alpha/2}
  \right) c_0 \\
  \gamma &=& (2\pi)^{1/\alpha} T^{2/\alpha} \left(\E
    |z|^{\alpha/2}
  \right)^{2/\alpha} c_0^{2/\alpha}
\end{eqnarray*}
Here we note that an $\alpha$-stable distribution $S(\alpha, \beta,
\gamma, \mu)$ has characteristic function \cite{Guhr2007}
\begin{eqnarray*}
  \varphi(k; \alpha, \beta, \gamma, \mu) &=& \exp\left[
    i\mu k - \gamma^\alpha |k|^\alpha \left(
      1 - i \beta {k \over |k|} \tan{\pi \alpha \over 2}
    \right) \right] \text{ for $\alpha \neq 1$}
\end{eqnarray*}
from which we see $\varphi(ak; \alpha, \beta, \gamma, \mu) = \varphi(k;
\alpha, \beta, a\gamma, a\mu)$, implying that, if $x \sim S(\alpha,
\beta, \gamma, \mu)$, then $ax \sim S(\alpha, \beta, a\gamma, a\mu)$.

Now that
\begin{eqnarray*}
  \sum_{t=1}^T r_t^2 &\xrightarrow{d} S(\alpha/2,
  1, \gamma, \mu) \text{ as $T \to \infty$}  
\end{eqnarray*}
we have
\begin{eqnarray}
  C_{ii} &=& {1 \over T^{2/\alpha}}\sum_{t=1}^T r_{it}^2 \\
  &\xrightarrow{d}&
  S(\alpha/2, 1, \gamma_D, \mu_D) \text{ as $T \to \infty$}
  \label{eq:garch_cii_dist}
\end{eqnarray}
where
\begin{eqnarray}\label{eq:garch_wishart_cij_params}
  \gamma_D &=& (2\pi)^{1/\alpha} \E\left(|z|^{\alpha/2}
  \right)^{2/\alpha} c_0^{2/\alpha} \\
  \mu_D &=& {\alpha_0 \over 1 - \alpha_1 - \beta_1} T^{1 - 2/\alpha}
  \nonumber
\end{eqnarray}
So the diagonal elements of the covariance matrix converge to
an $\alpha$-stable distribution with L\'evy index $\alpha/2 \approx
1.48$. This is comparable to the index value 1.38 obtained by
fitting. Considering the slow convergene of regularly varying tails,
this is a reasonably good match.

Now we look at the distribution of the non-diagonal elements. Figure
\ref{fig:garch_cij} shows that an $\alpha$-stable distribution fits
rather well, and additionally, figure \ref{fig:garch_nondiag_probplot}
shows that the distribution of these non-diagonal elements has fat
tails, supporting an $\alpha$-stable distribution.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=81 229 114
    121]{../pics/garch_cij_prob.pdf}
    \caption{\small \it Probability plot of the non-diagonal
      elements. Blue: The accumulative probability function
      (CDF) of the non-diagonal elements. Black, dashed: CDF of the
      Gaussian distribution that has the same mean and variance as the
      sample. The graph is arranged on such a scale that the Gaussian
      CDF is a straight line.
    }
  \label{fig:garch_nondiag_probplot}
\end{figure}

The parameters of the fitted $\alpha$-stable distribution $S(\alpha',
\beta', \gamma', \mu')$ are obtained in the procedure of fitting. The
results have been shown in the legend of figure
\ref{fig:garch_cij}. In table \ref{tab:garch_wishart_cij_params} we
list them with a higher precision.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    $\alpha'$ & $\beta'$ & $\gamma'$ & $\mu'$ \\
    \hline
    1.9453 & 0.0018 & $2.1381 \times 10^{-5}$ & $2.0637\times 10^{-9}$ \\
    \hline
  \end{tabular}
  \caption{\small \it Parameters of the non-diagonal elements' distribution}
  \label{tab:garch_wishart_cij_params}
\end{table}
Since $r_{it}$ and $r_{jt}$ are independent of each other, $\beta'$
and $\mu'$ are expected to be 0 --- but with a finite T, some deviation
from 0 is not surprising.

Now that the parameters' values have been obtained, the way they scale
with T, i.e. the length of the return series, can be deduced. Consider
\begin{equation}\label{eq:garch_cij1}
  T^{2/\alpha} C_{ij} = \sum_{t=1}^T r_{it} r_{jt} \xrightarrow{d} S(\alpha',
  0, \gamma', 0)
\end{equation}
where the width $\gamma'$ is determined by \cite{Bilik2008},
\begin{eqnarray*}
  \lim_{T\to\infty} {T C \over \gamma'^{\alpha'}}
  &=& C_{\alpha'} \\
  C_{\alpha'} &=& \left( \int_0^\infty {\sin x \over x^{\alpha'}} dx
  \right)^{-1} \\
\end{eqnarray*}
Here the constant C is such that
\begin{equation*}
  P(|r_{it}r_{jt}| > x) \sim {C \over x^{\alpha'}} \text{ as $x \to \infty$}
\end{equation*}
So we have
\begin{eqnarray*}
  \gamma' &=& {\left( CT \over C_{\alpha'}\right)^{1/\alpha'}}
\end{eqnarray*}
Divide throughout equation \ref{eq:garch_cij1} by $T^{2/\alpha}$ then
gives
\begin{equation}\label{eq:garch_cij2}
  C_{ij} = {1 \over T^{2/\alpha}}\sum_{t=1}^T r_{it} r_{jt}
  \xrightarrow{d}
  S(\alpha', 0, \gamma_N, 0)
\end{equation}
where
\begin{eqnarray*}
  \gamma_N &=& {\left(C \over C_{\alpha'}\right)^{1/\alpha'}}
  T^{1/\alpha'-2/\alpha} \\
\end{eqnarray*}
So we see that distribution of the non-diagonal elements has a width
that scales with T as $T^{1/\alpha' - 2/\alpha} \approx T^{-1/6}$,
while the width of the diagonal elements' distribution, as shown in
equation \ref{eq:garch_wishart_cij_params}, does not scale with
T. This is to be compared with the Wishart case where the return
series have Gaussian distribution, and hence the asymptotic
distributions of both the diagonal and the non-diagonal elements of the
covariance matrix are Gaussian with a variance that scales as $1/T$
(c.f. \ref{sec:GCC-analytical}).

% We also notice that, given our particular choice of $\alpha_1$ and
% $\beta_1$, the covariance matrix's convergence to a diagonal matrix
% is slow, since the non-diagonal elements decay to zero only

% In the Wishart case, the diagonal and the non-diagonal elements
% scale in the same way with T and thus have comparable sizes at large
% T --- so the matrix is not diagonal and the eigenvalue distribution has
% the Marcenko-Pastur law. In the GARCH case, the diagonal elements do
% not scale with T while the non-diagonal elements scale as $T^{-1/6}$,
% thus at large T, the non-diagonal elements are significantly smaller
% than the diagonal ones --- so a diagonal matrix arises.

\section{Implications of Finite Number of Observations}
In the last section we have discussed the limiting situation where the
return series of the covariance matrix have an infinite number of
observations ($T \to \infty$), and in the simulation studies we have
generated a large number of observations for each return series,
namely $T = 8 \times 10^4$. However, in practice, one often does not
have such a large number of observations available. Therefore, it is
useful to investigate situations where $T$ only has a modest size.

Figure \ref{fig:GarchCiiEig1} shows the distributions of the diagonal
elements as well as the eigenvalues when the number of return series
(N) is fixed at 250 and the number of observations in each series (T)
is gradually increased from 3000 to 6000.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.7, clip=true, trim=81 248 75
    195]{../pics/GarchCiiEig1.pdf}
  \caption{\small \it The Diagonal elements' and the eigenvalues'
    distributions with modest T. Blue: eigenvalues' CDF;
    Green: Diagonal elements' CDF; Red: CDF of the $\alpha$-stable
    distribution fitted to the diagonal elements. All the curves are
    drawn on 10-based log-log scale.
  }
  \label{fig:GarchCiiEig1}
\end{figure}
It is seen from the figure that, compared to the earlier case where $T
= 8 \times 10^4$, the eigenvalue distribution and the diagonal
elements' distribution do not coincide as well but differ rather
siginificantly for small values. For large values of the diagonal
elements and the eigenvalues, the two do coincide and comply with the
fitted $\alpha$-stable distribution. Moreover, the difference between
the eigenvalues' distribution and the diagonal elements' distribution,
as well as that between the diagonal elements' distribution and the
fitting $\alpha$-stable distribution, is also seen to diminish as T
increases.

The convergence of the eigenvalues ($\lambda$) towards the diagonal
elements $C_{ii}$ as $\lambda \to \infty$ is really an anticipated
result. For convenience, we order the diagonal elements so that
\begin{eqnarray*}
  C_{11} < C_{22} < \cdots < C_{NN}
\end{eqnarray*}
where, as before, $N$ is the dimension of the covariance matrix
$C$. Then, as $x \to \infty$,
\begin{eqnarray*}
  P(C_{ii} > x) &\sim& {c \over x^{\alpha/2}} \\
  f(C_{ii}) &\sim& {c \over x^{\alpha/2 + 1}} \\
\end{eqnarray*}
where $f(\cdot)$ denotes the \gls{pdf} of the diagonal elements and $c$ is
some constant. Thus, at the limit $C_{ii} \to \infty$, the distance
between two adjacent diagonal elements $C_{ii}$ and $C_{i+1, i+1}$ can
be expressed as
\begin{eqnarray*}
{1 \over N (C_{i+1, i+1} - C_{ii})} &=& f(C_{ii})\\
C_{i+1, i+1} - C_{ii} &\sim& {C_{ii}^{\alpha/2 + 1} \over c N}
\end{eqnarray*}

Thus as $C_{ii} \to \infty$, $C_{i+1, i+1} - C_{ii} \to \infty$ while
$C_{ij} \to 0$ ($i \neq j$). Now that the spacing between adjacent
diagonal elements become wider and the non-diagonal elements become
smaller, the eigenstates considered as a mixture of the basis states,
become more and more localized to a prominent basis state. This
localization can be measured by the size of the component in each
eigenvector that has the largest absolute value($|c|_\M$), provided
that the eigenvectors have been normalized.

Figure \ref{fig:garch_eigenvec_Cmax} shows how $|c|_\M$ changes
in response to increasing $\lambda$. It is seen that as T increases,
localization of the eigenstates proceeds from those with very
large eigenvalues towards those with relatively smaller
eigenvalues. At the same time, the minimum of $|c|_\M$ increases
and so does the minimum of the eigenvalues. The last point here is
further illustrated in figure \ref{fig:garch_eigenvec_Cmax_dist} where
the \gls{pdf} of $|c|_\M$ is plotted. We see in this figure that an
increased value of T leads to advancement of $\min(|c|_\M)$ to larger
values as well as to an increased proportion of large $|c|_\M$. The
mean of $|c|_\M$ has apparently been increased too.
\begin{figure}[htb!]
  \centering
  \subfigure[]{
  \includegraphics[scale=0.29, clip=true, trim=14 199 31
  155]{../pics/garch_eigenvec_Cmax.pdf}
    \label{fig:garch_eigenvec_Cmax}
  }
  \subfigure[]{
  \includegraphics[scale=0.41, clip=true, trim=86 257 112
  136]{../pics/garch_eigenvec_Cmax_dist.pdf}
  \label{fig:garch_eigenvec_Cmax_dist}
  }
  \caption{\small \it \ref{fig:garch_eigenvec_Cmax}: Horizontal:
    eigenvalues ($\log \lambda$); Vertical: $\log |c|_\M$ of the
    corresponding eigenvector. From left to right and 1st to 2nd row,
    T has values 800, 1000, 1200,
    80000. \ref{fig:garch_eigenvec_Cmax_dist}: PDF of the largest
    component ($|c|_\M$) of each eigenvector. In both plots the number
    of returns (N) is 50.}
%  \label{fig:garch_eigenvec_Cmax_both}
%   \includegraphics[scale=0.4, clip=true, trim=14 199 31
%   155]{../pics/garch_eigenvec_Cmax.pdf}
%   \caption{\small \it largest component ($|c|_\M$) corresponding to
%     the eigenvalue. N is fixed to 50.}
%   \label{fig:garch_eigenvec_Cmax}
% \end{figure}

% \begin{figure}[htb!]
%   \centering
%   \includegraphics[scale=0.4, clip=true, trim=86 257 112
%   136]{../pics/garch_eigenvec_Cmax_dist.pdf}
%   \caption{\small \it PDF of the largest component ($|c|_\M$) of
%     each eigenvector. N is fixed to 50.}
%   \label{fig:garch_eigenvec_Cmax_dist}
\end{figure}

Another informative quantity that measures the localization is the
``Inverse Participation Ratio'' (IPR). For a given normalized eigenvector
$\vec{c}_i = (c_{1, i}, c_{2, i}, ..., c_{N, i})$, the \gls{ipr} is defined
as \cite{Aberg2013}
\begin{eqnarray}
  \text{IPR}(\vec{c}_i) &=& \sum_{k=1}^N c_{k,i}^4 \label{eq:IPR_def}
\end{eqnarray}
Figure \ref{fig:garch_eigenvec_PR} shows ${1 \over
  N \cdot \text{IPR}(\vec{c}_i)}$ in correspondence to the
eigenvalues. This quantity is sometimes termed the
normalized \gls{pr} and measures the proportion of
basis vectors that contribute considerably to the eigenvector in
question. From this figure we see that, for all values of T, if an
eigenvalue is larger than $10^{-1.5} \approx 0.03$, its corresponding
participation ratio is less than $10^{-1.6} = 2.5\%$, meaning less
than $50 \times 0.025 = 1.26$ basis vectors contribute --- each of the
corresponding eigenvectors is localized to a single basis vector and
hence the distribution of such large eigenvalues is the same as the
diagonal elements' distribution.

Figure \ref{fig:garch_eigenvec_PR_dist} shows the \gls{pdf} of the
normalized \gls{pr}. We see that as T increases, the distribution of
\gls{pr} is compressed towards 0, suggesting increased localization of
the eigenvectors.
\begin{figure}[htb!]
  \centering
  \subfigure[]{
    \label{fig:garch_eigenvec_PR}
    \includegraphics[scale=0.34, clip=true, trim=49 208 70
    162]{../pics/garch_eigenvec_PR.pdf}
  }
  \subfigure[]{
    \includegraphics[scale=0.4, clip=true, trim=97 259 113
    226]{../pics/garch_eigenvec_PR_dist.pdf}
  \label{fig:garch_eigenvec_PR_dist}
  }
  \caption{\small \it \ref{fig:garch_eigenvec_PR}: Normalized
    participation ratio (PR) versus eigenvalue
    ($\lambda$). Horizontal: $\log \lambda$; Vertical: $\log
    \text{PR}$ of the corresponding
    eigenvector. From left to right and 1st to 2nd row, T has values
    800, 1000, 1200, 80000. \ref{fig:garch_eigenvec_PR_dist}: PDF of
    the normalized PR. In both plots, the number of returns (N) is
    50.}
\end{figure}
In conclusion, localization of the eigenvectors, which implies 
coincident eigenvalue and diagonal elements' distributions, begins with
those associated to large eigenvalues. Increased observation points
lead to increased localization and hence increased coincident sections
of the eigenvalue and diagonal elements' distributions. However, this
increment with T is slow, because the diagonal elements mean $\mu_D$
increases only as a fractional power of $T$, namely $T^{1 -
  2/\alpha}$, and the non-diagonal elements' variance decreases only
as a fractional power too, namely $T^{1/\alpha' - 2/\alpha}$. These
have been detailed in equations \ref{eq:garch_wishart_cij_params} and
\ref{eq:garch_cij2}.

\section{Influence of auto-correlations}
\label{sec:garch_nonzero_phi}
In the previous two sections we have studied situations where
$\phi=0$ in the specification \ref{eq:garch_spec}, i.e. no
auto-correlation is in the returns. In this section we investigate how
auto-correlations change the picture.

% When auto-correlations are present among the returns, the
% covariance matrix is expected to change. How exactly it changes
% is the subject of this section. Here we consider a model specified
% as follows:
% \begin{eqnarray*}
%   r_{it} &=& \phi r_{i, t-1} + \epsilon_{it} \\
%   \epsilon_{it} &=& \sigma_{it} z_{it} \\
%   \sigma_{it}^2 &=& \alpha_0 + \alpha_1 \epsilon_{i, t-1}^2 + \beta_1
%   \sigma_{i, t-1}^2 \\
%   z_{it} &\sim& N(0, 1)
% \end{eqnarray*}
% where $r_{it}$ is the element of the R matrix at the $i$-th row and the
% $t$-th column. In the simulations described below R has N=50 rows and
% $T = 8 \times 10^4$ columns. The parameters $\alpha_1$ and $\beta_1$
% are the same as in the previous case of zero auto-correlations, namely
% $\alpha_1 = 0.15$ and $\beta_1 = 0.84$. The covariance matrix C
% is built from R using
% \begin{eqnarray*}
%   C &=& {1 \over T^{2/\alpha}} RR'
% \end{eqnarray*}
% where $\alpha$ is 2.96 as before.

Figure \ref{fig:GarchEigDiag1} shows the eigenvalue as well as the
diagonal elements' distribution when $\phi = 0.95$, i.e. $\tau =
13.51$. The values of N and T are 50 and $8\times 10^4$ as before. The
\gls{garch}(1,1) parameters are also unchanged, namely $\alpha_0 = 2.3\times
10^{-6}$, $\alpha_1 = 0.15$, and $\beta_1 = 0.84$. Figure
\ref{fig:GarchNondiag1} shows the non-diagonal elements' distribution
in the same setup. Included in these plots are $50 \times 2000 =
1\times 10^5$ eigenvalues and diagonal elements, as well as ${50
  \choose 2} \times 2000 = 2,450,000$ non-diagonal elements. These
data come from 2000 simulated matrices.
\begin{figure}[htb!]
  \centering
  \subfigure[]{
    \includegraphics[scale=0.42, clip=true, trim=112 272 103
    217]{../pics/GarchEigDiag1.pdf}
    \label{fig:GarchEigDiag1}
  }
  \subfigure[]{
    \includegraphics[scale=0.42, clip=true, trim=105 272 101
    213]{../pics/GarchNondiag1.pdf}
    \label{fig:GarchNondiag1}
  }
  \caption{\small \it \ref{fig:GarchEigDiag1}: Eigenvalues' and
    diagonal elements' distribution when $\phi = 0.95$, i.e. $\tau$ =
    13.51; \ref{fig:GarchNondiag1}: Non-diagonal elements'
    distribution in the same situation.}
\end{figure}

From figure \ref{fig:GarchEigDiag1} we see that, as auto-correlations
become significant, the distribution of the eigenvalues no longer
coincides with the diagonal elements' distribution --- instead it
becomes wider and fatter on the tails. We also notice that the widths
of both the diagonal and the non-diagonal elements' PDF's have
increased.  In figure \ref{fig:garch_ev_cii} we see that, when no
auto-correlation is present, the \gls{pdf} of the diagonal elements has
width ($\gamma$) $7.31 \times 10^{-4}$, while in figure
\ref{fig:GarchEigDiag1} we see that the width has become $7.91 \times
10^{-3}$ as $\tau$ becomes 13.51. Similarly the non-diagonal elements'
\gls{pdf} has width $2.14 \times 10^{-5}$ when $\tau=0$, as shown in figure
\ref{fig:garch_cij}, and this width becomes $9.60 \times 10^{-4}$ when
$\tau = 13.51$, as shown in figure \ref{fig:GarchNondiag1}.

Figure \ref{fig:GarchSpectrumAutocorrelated} shows the eigenvalues'
distribution corresponding to a range of $\phi$ values. The number of
eigenvalues in each curve is the same as in figure
\ref{fig:GarchEigDiag1}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=85 258 103
  229]{../pics/GarchSpectrumAutocorrelated.pdf}
  \caption{\small \it Eigenvalue distribution of covariance
    matrix built from auto-correlated GARCH processes. $\phi$ = 0,
    0.5, 0.8, 0.955, 0.97 correspond to correlation time $\tau$ =
    0, 1.00, 3.11, 15.05, 22.76.}
  \label{fig:GarchSpectrumAutocorrelated}
\end{figure}
From this figure one can see that, as auto-correlation strengthens,
\begin{itemize}
\item the \gls{pdf} of the eigenvalue distribution flattens and widens;
\item the minimum as well as the maximum eigenvalues increase.
\end{itemize}

To find out more about this series of deformation, we first look at
how the largest component and the normalized participation ratio of
the eigenvectors change as $\phi$ takes on larger values. Figure
\ref{fig:garch_eigenvec_Cmax_corr} shows the largest eigenvector
component $|c|_\M$ in correspondence to the eigenvalue. Apparently, as
auto-correlation strengthens, the eigenvectors' composition
fractures, leading to a reduced degree of localization and even
reduced certainty of localization --- for a fixed eigenvalue, $|c|_\M$
now varies in a larger range than it does with smaller $\phi$.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=0 197 0
  154]{../pics/garch_eigenvec_Cmax_corr.pdf}
  \caption{\small \it Eigenvectors' largest component versus
    eigenvalue, for 4 auto-correlation strengths $\phi$ = 0, 0.6, 0.9,
    and 0.99. The number of returns (N) is 50.}
  \label{fig:garch_eigenvec_Cmax_corr}
\end{figure}
The same story of reduced localization is also evident from the plot
of the normalized \gls{pr}, shown in figure
\ref{fig:garch_eigenvec_PR_dist_corr}, and from the \gls{pdf} of $|c|_\M$
shown in figure \ref{fig:garch_eigenvec_Cmax_dist_corr}. It is seen
in \ref{fig:garch_eigenvec_PR_dist_corr} that the peak at the left 
of the plot, representing the group of localized eigenvectors, falls
with increased auto-correlation, and essentially disappears when
$\phi$ reaches the extreme value 0.99. Figure
\ref{fig:garch_eigenvec_Cmax_dist_corr} shows the proportion of large
$|c|_\M$ values is severely reduced and the mean of $|c|_\M$ is pushed
to smaller values by increased auto-correlation.
\begin{figure}[htb!]
  \centering
  \subfigure[]{
    \includegraphics[scale=0.4, clip=true, trim=95 260 111
    232]{../pics/garch_eigenvec_PR_dist_corr.pdf}
    \label{fig:garch_eigenvec_PR_dist_corr}
  }
  \subfigure[]{
    \includegraphics[scale=0.4, clip=true, trim=95 260 111
    232]{../pics/garch_eigenvec_Cmax_dist_corr.pdf}
    \label{fig:garch_eigenvec_Cmax_dist_corr}
  }
  \caption{\small \it \ref{fig:garch_eigenvec_PR_dist_corr}: PDF of
    the normalized participation ratio
    (PR). \ref{fig:garch_eigenvec_Cmax_dist_corr}: PDF of the largest
    component of the eigenvectors. The number of returns (N) is
    50. $\phi$ values of 0, 0.6000, 0.9000, 0.9900 correspond to
    correlation time $\tau$ = 0, 1.3569, 6.5788, 68.9676}
\end{figure}

It is also useful to look at how the fraction of localized eigenvectors
changes with the auto-correlation. For definiteness, we classify
an eigenvector as being localized when (1) its number of participating
basis vectors is less than 2, or (2) the largest of its components'
absolute values is larger than 0.9.

Figure \ref{fig:localization_ratio} shows how the ratio of localized
eigenvectors depends on the auto-correlation strength $\phi$. In
either way of classification, the ratio falls with $\phi$ in
accordance with a power law, the power exponent lying a bit below 2.
This is further confirmed in plot \ref{fig:localization_ratio2}, where
the ratios are plotted versus $\phi$ on log-log scale.
\begin{figure}[htb!]
  \centering
  \subfigure[]{
    \includegraphics[scale=0.38, clip=true, trim=93 229 115
    134]{../pics/localization_ratio.pdf}
    \label{fig:localization_ratio}
  }
  \subfigure[]{
    \includegraphics[scale=0.38, clip=true, trim=87 227 115
    133]{../pics/localization_ratio2.pdf}
    \label{fig:localization_ratio2}
  }
  \caption{\small \it \ref{fig:localization_ratio}: Ratio of localized
    eigenvectors versus the auto-correlation strength $\phi$. Black
    ``+'': ratio of localized eigenvectors as measured by the number
    of participating basis vectors being lower than 2; Black ``x'':
    ratio of localized eigenvectors as measured by the largest
    component being larger than 0.9. Blue curve: quadradic function
    fitted to ``+''. Green curve: quadratic function fitted to
    ``x''. There are 27 data points in the plot, corresponding to 27
    $\phi$ values: 0 to 0.8 with step size 0.05, and 0.9 to 0.99 with
    step size 0.01. The corresponding values of the correlation time
    $\tau$ range from 0 to 69. \ref{fig:localization_ratio2}: $\ln
    (f_\M - f)$ is plotted versus $\ln \phi$, where $f$ stands for the
    ratio of localized eigenvectors. The fitted curves are linear.}
\end{figure}

\input{summary}

\chapter{Self-reflection}
First of all, during the thesis project I learned a lot about models
of time series in finance, such as \gls{garch}, and \gls{sv} models,
as well as traditional time series analysis, for example, \gls{arima}
models. In addition, a study of continuous-time models gave me an
opportunity to increase my knowledge of stochastic differential
equations, probability theory, and statistics. I am a beginner, but
nonetheless find it comfortable to read books and articles on these
subjects after the thesis project.

Moreover, I have introduced myself to the random matrix theory,
multivariate analysis, and extreme value theory. These became relevant
when I worked on the covariance matrix of returns described by
Gaussian or \gls{garch} models. I am still a long way from being able
to do serious proofs or derivations in these areas, but I know what to
read and can understand the basics.

In addition, I have become better at Matlab programming and dealing
with databases, and have grown more confident in numerical analysis.

\appendix
\input{appendix}

\bibliographystyle{unsrt}
\bibliography{econophysics}
\end{document}