### 2014-09-29 thesis/appendix.tex

% \section{Fundamental Concepts \& Notations}
% \label{sec:FundamentalConcepts}
% This section is a list of a few concepts that may be unfamiliar to the
% reader and that we will often refer to later in the thesis:
% \begin{itemize}
% \item Return. Given a fixed time interval $[t - \Delta t, t]$, for example, a
%   day, a week, a month, etc, and a particular asset, for example, a
%   share in company ABC, the return of this asset over the time
%   interval is defined as
%   \begin{eqnarray*}
%     r_{t, \Delta t} &=& \ln S_{t} - \ln S_{t - \Delta t} \\
%     &=& \ln \left(1 + {S_{t} - S_{t - \Delta t} \over S_{t - \Delta t}}\right) \\
%     &\approx& {S_{t} - S_{t - \Delta t} \over S_{t - \Delta t}}
%   \end{eqnarray*}
%   where $S_t$ is the price of the asset at time $t$. $\Delta t$ is
%   sometimes called the time-lag of the return. Quite often, where
%   confusion is not possible, we will just write $r_t$ to mean $r_{t,
%     \Delta t}$, the time lag either does not matter or is clear from
%   the context.

% \item Autocorrelation. By autocorrelation, denoted $\rho_k$ here, we
%   mean the correlation between two temporally separated observations
%   of the same time series:
%   \begin{eqnarray*}
%     \rho_k &=& {
%       \E\left[(a_t -\E(a_t))(a_{t-k} - \E(a_{t-k})\right]
%       \over
%       \sqrt{\text{var}(a_t)}\sqrt{\text{var}(a_{t-k})}
%     }
%   \end{eqnarray*}
%   Here $\E(x)$ stands for the expectation value of $x$, and
%   $\text{var}(x)$ stands for the variance of $x$. $k$ is called the
%   time-lag and is the temporal seperation of the two observations
%   measured by the number of observations in between. For example, the
%   time-lag between the 1st and 3rd observation is 2.

% \item Covariance Matrix. Correlations between the returns of a
%   group of assets are described by the covariance matrix. When
%   the asset returns are described by a stable distribution law with
%   L\'evy index $\alpha$, the empirical covariance matrix is
%   constructed as
%   \begin{equation}
%     \label{eq:covariance}
%     \begin{aligned}
%       C_{ij} &= {1 \over T^{2/\alpha}} \sum_{t=1}^T [r_{i,t}-\E(r_i)]
%       [r_{j,t}-\E(r_j)] \\
%       &= {1 \over T^{2/\alpha}} RR'
%     \end{aligned}
%   \end{equation}
%   where $r_{i,t}$ is the return of the i-th asset at time t and is
%   placed at the entry (i, t) of matrix R; $R'$ denotes the
%   transpose of R. In most practical situations, one has abundant data
%   for each and every asset. Thus $T \geq N$ is assumed throughout this
%   thesis.

% \item Auto-regressive processes. A time series $r_t$ is called an
%   auto-regressive process of order $p$ and denoted AR(p), if it can be
%   written in the following form:
%   \begin{eqnarray*}
%     r_t &=& \sum_{i=1}^p \phi_i r_{t-i} + a_t
%   \end{eqnarray*}
%   where, for all $i$, $\phi_i \in (-1, 1)$, and the $a_t$'s are
%   independent and identically distributed (iid.) random variables with
%   zero mean. Obviously this implies the mean of $r_t$ is zero
%   too. Apart from this, their distribution of $a_t$ is not restricted
%   to any particular form.

%   Of particular interest to this thesis is the AR(1) process:
%   \begin{eqnarray*}
%     r_t = \phi r_{t-1} + a_t
%   \end{eqnarray*}
%   Its autocorrelation function $\rho_k = \text{corr}(r_t, r_{t-k})$
%   ($k = 0, 1, 2, \cdots$) can be easily shown to fall off
%   exponentially:
%   \begin{eqnarray*}
%     \rho_k &=& {\E(r_tr_{t-k}) - \E(r_{t-k})\E(r_{t}) \over
%       \sqrt{\var(r_t) \var(r_{t-k})}} \\
%     &=& {\phi \E(r_{t-1}r_{t-k}) + \E(a_t r_{t-k})
%       \over
%       \sqrt{\var(r_t) \var(r_{t-k})}
%     } \\
%     &=& \phi \rho_{k-1}
%   \end{eqnarray*}
%   where $\E(a_t r_{t-k}) = 0$ follows from the fact that any return
%   $r_{t1}$ must not depend on disturbances $a_{t2}$ that occur later
%   in time. The last equation means $\rho_k$ is a geometric series. Since
%   $\rho_0 = 1$, we have
%   \begin{eqnarray*}
%     \rho_k &=& \phi^k \\
%     |\rho_k| &=& e^{k\ln|\phi|} \\
%   \end{eqnarray*}
%   Note that $|\phi| < 1$ and hence $\ln|\phi| < 0$.

%   Although $k$ can only take integer values, it is still useful to
%   define a correlation time $\tau$ such that $\phi^\tau = 1/2$. Such a
%   quantity is more intuitive and constitutes a measure of
%   autocorrelations that is universal and comparable among different
%   time series' models. From the definition of $\tau$ we get
%   \begin{equation}
%     \label{eq:tau_def}
%     \begin{aligned}
%       \phi^\tau &= 1/2 \\
%       \tau &= -{\ln 2 \over \ln\phi} \\
%       \phi &= 2^{-1/\tau}
%     \end{aligned}
%   \end{equation}
%   Figure \ref{fig:AR1-autocorrelation} shows the autocorrelation
%   function of the AR(1) model. As proven above, this function decays
%   exponentially.
%   \begin{figure}[htb!]
%     %\vspace{-15mm}
%     \centering
%     \includegraphics[scale=0.5, clip=true, trim=113 229 115
%     139]{../pics/AR1-autocorrelation.pdf}
%     \caption{\small \it Autocorrelation function of the AR(1) model
%       with $\phi=1/\sqrt{2}$. Red circles: autocorrelations at $k=0, 1,
%       2, \cdots$. Blue line: $e^{t\ln\phi}$.}
%     \label{fig:AR1-autocorrelation}
%   \end{figure}
%   For more details about conventional time series' models, see
%   \cite{BoxJenkins94}.
% \end{itemize}

% % For purposes of later reference, we also list some notations that may
% % cause confusion to the reader:
% % \begin{itemize}
% % \item $\E(x)$ or $\mean{x}$: The expectation value of $x$.
% % \item $\var(x)$: the variance of $x$.
% % \item $\text{cov}(x,y)$: the covariance of $x$ and $y$.
% % \item $\text{corr}(x,y)$: the correlation between $x$ and $y$, i.e.
% %   \begin{equation*}
% %     \text{corr}(x,y) = {\text{cov}(x,y) \over \sqrt{\var(x)\var(y)}}
% %   \end{equation*}
% % \end{itemize}

% \section{Stylized facts}\label{sec:StylizedFacts}
% This section presents and explains a few ``stylized facts'',
% i.e. phenomena that are widely observed and accepted as true.
% \begin{itemize}
% \item Fat tails. It has been consistently reported by various studies
%   - for example \cite{Potters2003} and \cite{Mantegna2000} - that the
%   probability density function (PDF) of stock/index returns are not
%   Gaussian. Unlike the Gaussian PDF, which is symmetric and falls off
%   very quickly as its argument moves from the center to the outskirts
%   (tails), the PDF of stock/index returns are higher on the tails,
%   i.e. the probability of large fluctuations is higher than is dictated
%   by a Gaussian distribution - in fact, even higher than dictated by
%   an exponential function. Empirical studies suggest the distribution
%   function of the returns follows a power-law on the tails --- the
%   exponent of the power depends on the specific stock/index.

%   Figure \ref{fig:FatTail} illustrates this feature.
%   \begin{figure}[htb!]
%     \centering
%     \includegraphics[scale=0.6, clip=true, trim=92 229 110
%     140]{../pics/FatTail.pdf}
%     \caption{\small \it Fat tails of {\it Nordea Bank} 15-minute
%       returns during the period 2013-10-10 and 2014-01-29. The returns
%       are computed using minute-by-minute average prices.}
%     \label{fig:FatTail}
%   \end{figure}

% \item Non-zero skewness. Apart from fat tails, the PDF of stock/index
%   returns are often also skewed. If the skewness is positive
%   (negative), the probability of very large positive (negative)
%   returns is higher than that of very large negative (positive)
%   returns, even though the mean of the returns is 0 or extremely
%   close to 0.

%   Table \ref{tab:EmpiricalSkewness} lists the skewness of a few
%   Swedish stocks traded on the Stockholm OMX market.
%   \begin{table}[htb!]
%     \footnotesize
%     \centering
%     \begin{tabular}{|c|c|c|c|c|c|c|}
%       \hline
%       Nordea Bank & Volvo B & Boliden & ABB Ltd & H\&M & Scania
%       B & Ericsson B \\
%       \hline
%       0.1362 & 0.0471 & 0.0567 & -0.0364 & -0.0168 &
%       -1.1554 & 0.2086 \\
%       \hline
%     \end{tabular}
%     \caption{\small \it Skewness of Stock Returns. All the returns
%       have time-lag of 15 minutes and are computed using paid prices
%       between 2013-10-10 and 2014-01-29. }
%     \label{tab:EmpiricalSkewness}
%   \end{table}

% \item Higher-order autocorrelation. To the lowest order, return series
%   are not auto-correlated --- if they are, the auto-correlations would
%   present an obvious opportunity of making easy profit and be
%   exploited and vanish as soon as they appear. However, the squared
%   returns do have significant auto-correlations, as figure
%   \ref{fig:nordea_15min_acf} illustrates. These auto-correlations
%   suggest the variances of the returns at subsequent time steps are
%   correlated. These are referred to as higher-order auto-correlations
%   and make a subjet of returns' models.

% \end{itemize}
\chapter{Case Study of Some Intraday Series}
\label{chp:appendix2}
\section{Nordea 30-minute Returns}
\label{sec:nordea2_30min}
In this section we study the volatility of Nordea Bank 30-minute
returns in the period 2013/10/10 - 2014/04/04. The realized
volatilities that approximate the volatilities of these 30-minute
returns are computed using 1-minute returns. As in the previous cases,
this choice of 1-minute returns is confirmed by the normality of the
quotient $(r_t - \E(r_t))/\sigma_t$, which is shown in figure
\ref{fig:nordea3_quotient}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=78 255 108
  120]{../pics/nordea3_quotient.pdf}
  \caption{\small \it Normal probability plot of $(r_t -
    \E(r_t))/\sigma_t$}
  \label{fig:nordea3_quotient}
\end{figure}
The auto-correlations of $\ln \sigma_t$ is shown in figure
\ref{fig:nordea3_lv_acf}, where we see an apparent seasonality of
16. By differencing we get $w_t = (1-B)(1-B^s)\ln\sigma_t$ where
$s=16$. Its auto-correlations are shown in figure
\ref{fig:nordea3_w_acf}.
\begin{figure}[htb!]
  \centering
  \subfigure[]{
    \includegraphics[scale=0.4, clip=true, trim=86 256 101
    121]{../pics/nordea3_lv_acf.pdf}
    \label{fig:nordea3_lv_acf}
  }
  \subfigure[]{
    \includegraphics[scale=0.4, clip=true, trim=86 256 103
    122]{../pics/nordea3_w_acf.pdf}
    \label{fig:nordea3_w_acf}
  }
  \caption{\small \it \ref{fig:nordea3_lv_acf}: Auto-correlations
    (ACF) of $\ln \sigma_t$; \ref{fig:nordea3_w_acf}: Auto-correlations
    (ACF) of $w_t = (1-B)(1-B^s)\ln\sigma_t$.}
\end{figure}
Once again, this auto-correlation structure
points to a seasonal moving average model:
\begin{eqnarray*}
  w_t &=& (1-\theta B)(1-\Theta B^s) y_t
\end{eqnarray*}
where $y_t$ is assumed to have Gaussian distribution, neglecting
slight non-normality as before. Maximum likelihood estimation gives
the parameter values listed in table \ref{tab:nordea3_sv_param}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    Parameter & $\theta$ & $\Theta$ & $\var(y_t)$\\
    \hline
    Value & 0.7612 & 0.8036 & 0.0736\\
    \hline
  \end{tabular}
  \caption{\small \it Parameter values of the Seasonal Moving Average
    model}
  \label{tab:nordea3_sv_param}
\end{table}
To fit a \gls{garch} model to the same series, we assume Gaussian
innovations (c.f. equation \ref{eq:garch_def}). The result is a
\gls{garch}(1,1) model. Its parameters are listed in table
\ref{tab:nordea3_garch_param}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$ & $\beta_1$ \\
    \hline
    Value & $2.3 \times 10^{-7}$ & 0.05 & 0.9\\
    \hline
  \end{tabular}
  \caption{\small \it Parameter values of GARCH(1,1) model fitted to
    Nordea 30-minute returns.}
  \label{tab:nordea3_garch_param}
\end{table}

As before, we compare the accuracies of the \gls{garch} and the stochastic
volatility model by comparing their one-ste-ahead forecasts. We
compute the difference between their forecasts and the realized
volatilities and then look at the statistics of these difference
values. Firstly we show the mean and the standard deviation of the
differences in table \ref{tab:nordea3_diff_1}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
     & \gls{sv} & \gls{garch} & Sample mean \\
     \hline
    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & -0.0047 & 0.0130 & 0.0584 \\
    \hline
     $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.2602 &
     0.3194 & 0.3093 \\
    \hline
 \end{tabular}
  \caption{\small \it Mean and standard deviation of the 3 kinds of
    forecasts}
  \label{tab:nordea3_diff_1}
\end{table}
It is seen in this table that the \gls{sv} forecasts have a mean whose
absolute value is just above 1/3 of that of the \gls{garch} forecasts. The
standard deviation of the \gls{sv} forecasts is also smaller than that of
\gls{garch}. In addition, we check the distribution of
$\ln \sigma^F_t - \ln \hat{\sigma}_t$, and the percentage of ``good''
forecasts with respect to different criteria of goodness. These are
shown in figure \ref{fig:nordea3_diff} and table
\ref{tab:nordea3_diff_2}, respectively.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.54, clip=true, trim=21 312 0
    179]{../pics/nordea3_diff.pdf}
  \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
    sample mean forecasts. Left: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t < x)$;
    Right: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t > x)$. Horizontal: $x$.}
  \label{fig:nordea3_diff}
\end{figure}
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    \gls{sv} & \gls{garch} & Sample Mean \\
    \hline
    1\% & 18\% & 18\% & 15\% \\
    \hline
    5\% & 81\% & 71\% & 69\% \\
    \hline
    10\% & 98\% & 93\% & 95\% \\
    \hline
  \end{tabular}
  % \begin{tabular}{|c|c|c|c|}
  %   \hline
  %   ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
  %     \hat{\sigma}_t|}$ &
  %   \gls{sv} & \gls{garch} & Sample Mean \\
  %   \hline
  %   1\% & 17.77\% & 17.77\% & 15.38\% \\
  %   \hline
  %   5\% & 80.90\% & 71.09\% & 69.23\% \\
  %   \hline
  %   10\% & 98.14\% & 93.10\% & 95.23\% \\
  %   \hline
  % \end{tabular}
  \caption{\small \it Percentage of ``good'' forecasts as defined by
    deviating nore more than 1\%, 5\% and 10\% from the measured
    realized volatility.}
  \label{tab:nordea3_diff_2}
\end{table}
It is clear that the \gls{sv} forecasts are considerably more accurate.

\section{Volvo 15-minute returns in 2013/14}
\label{sec:volvo}
In this section we study the volatility of Volvo B 15-minute returns
during the period 2013/10/10 - 2014/04/04.
The subject of modeling is the realized volatility
computed as the square root of the sum of squared 1-minute
returns. The normal probability plot of the quotient $(r_t -
\E(r_t))/\sigma_t$ is shown in figure \ref{fig:volvo_15_quotient},
from which one can see it is normally distributed, confirming the
choice of 1-minute returns for computing the realized volatility.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=83 259 110 
  220]{../pics/volvo_15_quotient.pdf}
  \caption{\small \it Normal probability plot of $(r_t -
    \E(r_t))/\sigma_t$}
  \label{fig:volvo_15_quotient}
\end{figure}
The auto-correlations of $\ln \sigma_t$ and $w_t = (1-B)(1-B^s) \ln
\sigma_t$ are plotted in figure \ref{fig:volvo_15_lv_acf} and
\ref{fig:volvo_15_w_acf}. The former clearly shows the seasonality
$s = 33$ in the auto-correlations of $\ln \sigma_t$.

\begin{figure}[htb!]
  \centering
  \subfigure[]{
  \includegraphics[scale=0.4, clip=true, trim=90 259 103
  220]{../pics/volvo_15_lv_acf.pdf}
  \label{fig:volvo_15_lv_acf}
  }
  \subfigure[]{
  \includegraphics[scale=0.4, clip=true, trim=90 259 103
  220]{../pics/volvo_15_w_acf.pdf}
  \label{fig:volvo_15_w_acf}
  }
  \caption{\small \it Left: auto-correlations of $\ln \sigma_t$;
    Right: auto-correlations of $w_t = (1-B)(1-B^s) \ln \sigma_t$.}
\end{figure}

Based on the auto-correlations of $w_t$, a seasonal moving average
model is estimated:
\begin{eqnarray*}
  (1-B)(1-B^s) \ln\sigma_t &=& (1-\theta_1B - \theta_2B^2 - \theta_3
  B^3)(1 - \Theta B^s)y_t
\end{eqnarray*}
By maximum likelihood estimation, the parameter values listed in table
\ref{tab:volvo_15_sv_param} are obtained.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|}
  \hline
  Parameter & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\Theta$ & 
 $\var(y_t)$ \\
 \hline
 Value & 0.7491 & 0.1170 & 0.0571 & 0.8646 & 0.1245 \\
  \hline
  \end{tabular}
  \caption{\small \it Parameter values of the stochastoc volatility
    (SV) model.}
  \label{tab:volvo_15_sv_param}
\end{table}

A \gls{garch} model (c.f. equation \ref{eq:garch_def}) assuming Gaussian
innovations is also fitted to the same series. The parameter values
are listed in table \ref{tab:volvo_15_garch_param}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
  \hline
  Parameter & $\alpha_0$ & $\alpha_1$ & $\beta_{1}$ \\
  \hline
  Value & $6.54 \times 10^{-7}$ & 0.1565 & 0.6445\\
  \hline
  \end{tabular}
  \caption{\small \it Parameter values of the GARCH model.}
  \label{tab:volvo_15_garch_param}
\end{table}

Performing one-step-ahead forecasts using both models gives the series
$\ln\sigma^F_t - \ln\hat{\sigma}_t$, where $\ln\hat{\sigma}_t$ are the
measured realized volatilities. The mean and standard deviation of
$\ln\sigma^F_t - \ln\hat{\sigma}_t$ are shown in table
\ref{tab:volvo_15_stat}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
     & \gls{sv} & \gls{garch} & Sample mean \\
     \hline
    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.0035 & 0.0313 & -0.1448 \\
    \hline
     $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.3356 &
     0.3894 & 0.3826 \\
    \hline
 \end{tabular}
  \caption{\small \it Mean and standard deviation of the 3 kinds of
    forecasts}
  \label{tab:volvo_15_stat}
\end{table}
We see that both the \gls{sv} and the \gls{garch} model give biased forecasts, but
the \gls{sv} forecasts are only 1/10 as biased as those of \gls{garch} (0.0035
vs. 0.0313). In addition, the standard deviation of the \gls{sv} forecasts
is smaller too. These results are confirmed by the distribution of
$\ln \sigma^F_t - \ln \hat{\sigma}_t$ and the fraction of ``good''
forecasts, which are shown in figure \ref{fig:volvo_15_diff} and table
\ref{tab:volvo_15_diff}, respectively.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.54, clip=true, trim=27 276 0
    236]{../pics/volvo_15_diff.pdf}
  \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
    sample mean forecasts. Left: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t < x)$;
    Right: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t > x)$. Horizontal: $x$.}
  \label{fig:volvo_15_diff}
\end{figure}
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    \gls{sv} & \gls{garch} & Sample Mean \\
    \hline
    1\% & 58\% & 60\% & 39\% \\
    \hline
    5\% & 84\% & 84\% & 66\% \\
    \hline
    10\% & 97\% & 95\% & 91\% \\
    \hline
  \end{tabular}
  \caption{\small \it Percentage of ``good'' forecasts as defined by
    deviating nore more than 1\%, 5\% and 10\% from the measured
    realized volatility.}
  \label{tab:volvo_15_diff}
\end{table}

\section{Ericsson 15-minute Returns}
\label{sec:ericsson_15min}
In this section we model the volatility of {\it Ericsson B}
15-minute returns during the period 2013/10/10 - 2014/04/04.
Using the same method as with other series, we first find the right
higher-frequency returns that best approximate the volatility of the
15-minute returns. These turn out to be 50-second returns, as figure
\ref{fig:ericsson_15_quotient} shows.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=79 259 108
  121]{../pics/ericsson_15_quotient.pdf}
  \caption{\small \it Normal probability plot of $(r_t -\E(r_t))/
    \sigma_t$. $\sigma_t^2$ is computed as the sum of squared
    50-second returns.}
  \label{fig:ericsson_15_quotient}
\end{figure}

Following the procedure described in previous sections, an ARIMA model is
found for this series:
\begin{eqnarray*}
  (1-B)(1-B^s) \ln\sigma_t &=& (1- \theta_1B - \theta_2B^2 -
  \theta_3B^3)(1 - \Theta B^s) y_t
\end{eqnarray*}
where the seasonality $s$ is 33 and the other parameter values are
estimated to be those listed in table \ref{tab:ericsson_15_params}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Parameter & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\Theta$ &
    $\var(y_t)$ \\
    \hline
    Value & 0.8078 & 0.0454 & 0.0943 & 0.8798 & 0.1242 \\
    \hline
  \end{tabular}
  \caption{\small \it Ericsson B log-volatility parameters}
  \label{tab:ericsson_15_params}
\end{table}

Then a \gls{garch} model is also found with parameter values listed in
table \ref{tab:ericsson_15_garch_params}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$ & $\alpha_{s}$ & $\beta_{s}$ \\
    \hline
    Value & $2.4181 \times 10^{-7}$ & 0.1513 & 0.1409 & 0.6027 \\
    \hline
  \end{tabular}
  \caption{\small \it GARCH model of Volvo B 30-minute returns}
  \label{tab:ericsson_15_garch_params}
\end{table}

The forecasts from both models are compared as follows: Table
\ref{tab:ericsson_15_diff1} shows the mean and standard deviation of
the 3 kinds of forecasts:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    & \gls{sv} & \gls{garch} & Sample mean \\
    \hline
    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.0035 &
    0.0600 & 0.1116 \\
    \hline
    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.3278 &
    0.3667 & 0.3590 \\
    \hline
  \end{tabular}
  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
    \ln\hat{\sigma}_t$}
  \label{tab:ericsson_15_diff1}
\end{table}
Consistent with previous results, the bias introduced by the \gls{sv} model
is significantly smaller than that introduced by \gls{garch}. Figure
\ref{fig:ericsson_15_diff2} shows the distribution of the difference
between a forecast and its corresponding measured realized volatility,
i.e. $\ln\sigma^F_t - \ln\hat{\sigma}_t$; table
\ref{tab:ericsson_15_diff3} compares the percentage of ``good''
forecasts using the 3 alternatives.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=45 298 21
    165]{../pics/ericsson_15_diff2.pdf}
  \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
    sample mean forecasts. Left: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t < x)$;
    Right: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t > x)$. Horizontal: $x$.}
  \label{fig:ericsson_15_diff2}
\end{figure}

\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    \gls{sv} & \gls{garch} & Sample Mean \\
    \hline
    1\% & 17\% & 17\% & 16\% \\
    \hline
    5\% & 73\% & 69\% & 69\% \\
    \hline
    10\% & 96\% & 93\% & 94\% \\
    \hline
  \end{tabular}
  \caption{\small \it Fraction of ``good'' forecasts as defined by
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
  \label{tab:ericsson_15_diff3}
\end{table}
It is clear from figure \ref{fig:ericsson_15_diff2} and table
\ref{tab:ericsson_15_diff3} that the \gls{sv} model out-performs \gls{garch}. We
see that the \gls{sv} model yields considerably higher fractions of good
forecasts by all criteria, and in particular, gives much few
under-estimates.

\section{Ericsson 30-minute Returns}\label{sec:ericsson_30min}
In this section we look at the 30-minute returns of Ericsson B during
the period 2013/10/10 - 2014/04/04. Since the methods are the same as
with other series, we shall only present the results here.

First of all, the volatility of this series is found to be well
approximated by realized volatilities computed from 75-second
returns. The normal probability plot is shown in figure
\ref{fig:ericsson_30_quotient}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=79 259 108
  121]{../pics/ericsson_30_quotient.pdf}
  \caption{\small \it Normal probability plot of $(r_t -\E(r_t))/
    \sigma_t$. $\sigma_t^2$ is computed as the sum of squared
    75-second returns.}
  \label{fig:ericsson_30_quotient}
\end{figure}
The following ARIMA model is found for the log-volatility $\ln
\sigma_t$:
\begin{eqnarray*}
  (1-B)(1-B^s) \ln\sigma_t &=& (1- \theta_1B - \theta_2B^2) (1 -
  \Theta B^s) y_t 
\end{eqnarray*}
where the seasonality $s = 16$. The parameter values are estimated to
be those listed in table
\ref{tab:ericsson_30_params}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Parameter & $\theta_1$ & $\theta_2$ & $\Theta$ &
    $\var(y_t)$ \\
    \hline
    Value & 0.6842 & 0.2470 & 0.8391 & 0.0918 \\
    \hline
  \end{tabular}
  \caption{\small \it Ericsson B log-volatility parameters}
  \label{tab:ericsson_30_params}
\end{table}
A \gls{garch} model is also found with parameters listed in table
\ref{tab:ericsson_30_garch_params}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$ & $\alpha_{s}$ & $\beta_{s}$ \\
    \hline
    Value & $7.4856 \times 10^{-7}$ & 0.0532 & 0.1506 & 0.6778 \\
    \hline
  \end{tabular}
  \caption{\small \it GARCH(1, 1) model of Ericsson B 30-minute
    returns}
  \label{tab:ericsson_30_garch_params}
\end{table}

Table \ref{tab:ericsson_30_diff1} shows the mean and standard
deviation of the difference series $\ln\sigma^F_t -
\ln\hat{\sigma}_t$, where $\ln\sigma^F_t$ is the forecast
log-volatility and $\ln\hat{\sigma}_t$ is the measured realized
volatility.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    & \gls{sv} & \gls{garch} & Sample mean \\
    \hline
    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.0004 &
    0.0526 & 0.1039 \\
    \hline
    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.2519 &
    0.3019 & 0.2963 \\
    \hline
  \end{tabular}
  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
    \ln\hat{\sigma}_t$}
  \label{tab:ericsson_30_diff1}
\end{table}
Apparently, the \gls{garch} forecasts have a fairly large bias compared with
those of the \gls{sv} model. Turning to figure \ref{fig:ericsson_30_diff2}
and table \ref{tab:ericsson_30_diff3}, we also see the \gls{sv} model
performs consistently better --- a higher percentage of ``good''
forecasts are delivered and, while it under-estimates to around the
same extent as does \gls{garch}, it certainly over-estimates a lot
less. These are confirmative to what we have observed for the other
series.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=33 281 11
    150]{../pics/ericsson_30_diff2.pdf}
  \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
    sample mean forecasts. Left: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t < x)$;
    Right: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t > x)$. Horizontal: $x$.}
  \label{fig:ericsson_30_diff2}
\end{figure}

\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    \gls{sv} & \gls{garch} & Sample Mean \\
    \hline
    1\% & 22\% & 16\% & 16\% \\
    \hline
    5\% & 78\% & 69\% & 69\% \\
    \hline
    10\% & 98\% & 97\% & 96\% \\
    \hline
  \end{tabular}
  \caption{\small \it Fraction of ``good'' forecasts as defined by
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
  \label{tab:ericsson_30_diff3}
\end{table}

\section{Volatility of Forecast accuracy}
\label{sec:forecast_volatility}
In this section we compare the volatility of the accuracy of
\gls{garch} and \gls{sv} models' forecasts. Table
\ref{tab:5percent_values} lists the the percentage of forecasts
that deviate less than 5\% from the corresponding realized
volatilities, and also gives the standard deviation of the 6 numbers.
It is seen that the forecasts of \gls{sv} models have standard
deviation 0.0432 while those of \gls{garch} have 0.0722. Clearly the
accuracy of \gls{sv} forecasts varies considerably less than does that
of \gls{garch} forecasts. In other words, a \gls{sv} model performs
more consistently around a level of accuracy.
  \begin{table}[htb!]
    \centering
    \begin{tabular}{|c|c|c|c|c|}
      \hline
      & Volvo 15m & Volvo 30m & Ericsson 15m & Ericsson 30m \\
      \hline
      SV & 84\% & 72\% & 73\% & 78\% \\
      \hline
      GARCH & 84\% & 62\% & 69\% & 69\% \\
      \hline
      \hline
      & Nordea 1 15m & Nordea 2 30m & \multicolumn{2}{c|}{standard deviation}\\
      \hline
      SV & 75\% & 81\% & \multicolumn{2}{c|}{0.0432} \\
      \hline
      GARCH & 71\% & 71\% & \multicolumn{2}{c|}{0.0722} \\
      \hline
    \end{tabular}
    % \begin{tabular}{|c|c|c|c|c|}
    %   \hline
    %   & Volvo 15m & Volvo 30m & Ericsson 15m & Ericsson 30m \\
    %   \hline
    %   SV & 0.8351 & 0.7162 & 0.7268 & 0.7844 \\
    %   \hline
    %   GARCH & 0.8363 & 0.6154 & 0.6850 & 0.6900 \\
    %   \hline
    %   \hline
    %   & Nordea 1 15m & Nordea 2 30m & std.\\
    %   \hline
    %   SV & 0.7457 & 0.8090 & 0.0432 \\
    %   \hline
    %   GARCH & 0.7062 & 0.7109 & 0.0722 \\
    %   \hline
    % \end{tabular}
    \caption{\small \it Comparison of the percentage of forecasts that
      lie within 5\% of the corresponding realized
      volatilities. Nordea 1 refers to the Nordea data set of
      2012/01/16 - 2012/04/20 (see section \ref{chp:nordea_15min})
      and Nordea 2 refers to the data set of 2013/10/10 - 2014/04/04
      (see section \ref{sec:nordea2_30min}).}
    \label{tab:5percent_values}
  \end{table}

\chapter{Normalization Constant in the Unconditional PDF of the
  Symmetric SV model}
\label{chp:symmetric_SV_norm_const}
In the following we work out the normalization constant C used in
section \ref{sec:SLV_Symmetric}:
\begin{eqnarray*}
  \int_{-\infty}^{\infty} f_{r'}(x) dx &=& {1 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{-\infty}^{\infty} \text{erfc} \left({1 \over
      \sqrt{2}\sigma} \ln{|x| \over \sqrt{\ln 4}} + {\sigma \over
      \sqrt{2}} \right) dx\\
  &=& {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{0}^{\infty} \text{erfc} \left({1 \over
      \sqrt{2}\sigma} \ln{x \over \sqrt{\ln 4}} + {\sigma \over
      \sqrt{2}} \right) dx\\
\end{eqnarray*}
Let
\begin{eqnarray*}
  a &=& {1 \over \sigma \sqrt 2} \\
  b &=& -{1 \over 2 \sigma \sqrt 2}\ln\ln 4 + {\sigma \over \sqrt 2} \\
  y &=& a\ln|x|+b \\
\end{eqnarray*}
Then
\begin{eqnarray*}
  \int_{-\infty}^{\infty} f_{r'}(x) dx &=& {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{-\infty}^{\infty} dy {e^{(y-b)/a} \over a}
  \text{erfc}(y) \\
  &=& {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \left.e^{(y-b)/a}
    \text{erfc}(y)\right|_{y=-\infty}^{\infty}
  + {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{-\infty}^{\infty} dy {2 \over \sqrt \pi}
  e^{(y-b)/a} e^{-y^2} \\
  &=& {2 \over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}}  2 e^{\ln\ln 4/2
    - \sigma^2/2}\\
  C &=& \sqrt{2 \ln 4\over \pi} \\
  &\approx& 0.9394
\end{eqnarray*}

\chapter{PDF of the covariance matrix of
auto-correlated Gaussian Returns}
\label{app:pdf_gaussian1}
When autocorrelations exist among the columns of the Gaussian returns
matrix $\mtx R$, the distribution of the covariance matrix $\mtx C =
\mtx{R} \mtx{R'}$ is no longer Wishart but is nonethelss closely
related to it. In the following we consider the situation where
$r_{i,t}$ can be represented as a vector auto-regressive
process. Specifically, suppose
\begin{eqnarray*}
  \vec{r}_t &=& \sum_{k=1}^p \phi_k \vec{r}_{t-k} + \vec{a}_t \\
  \vec{a}_t &=& \vec{r}_t - \sum_{k=1}^p \phi_k \vec{r}_{t-k}
\end{eqnarray*}
where $\vec{a}_t = (a_{1,t}, a_{2,t}, \cdots, a_{N,t})' \sim N(0,
\mtx \Sigma)$ and comprise the columes of $\mtx A$; $\vec{r}_t$ are the
columes of $\mtx R$. Here $N(0, \mtx \Sigma)$ denotes the multivarate
normal distribution with covariance matrix $\mtx \Sigma$. The last equation
can be written in matrix form
\[
\mtx{A} = \mtx{R M}
\]
For example, in the case of an AR(1) process
\begin{equation*}
  \begin{pmatrix}
    a_{1,1} & a_{1,2} & \cdots & a_{1,T} \\
    \vdots & \ddots & \vdots \\
    a_{N,1} & a_{N,2} & \cdots & a_{N,T} \\
  \end{pmatrix} =
  \begin{pmatrix}
    r_{1,1} & r_{1,2} & \cdots & r_{1,T} \\
    \vdots & \ddots & \vdots \\
    r_{N,1} & r_{N,2} & \cdots & r_{N,T} \\
  \end{pmatrix}
  \begin{pmatrix}
    1 & -\phi &   &   & \\
      & 1 & -\phi &   & \\
      &   & \ddots  & \ddots &   \\
      &   &   & 1 & -\phi \\
      &   &   &   & 1 \\
  \end{pmatrix}
\end{equation*}
Let $\mtx{QR} = \mtx{RM}$, then $\mtx{Q} = \mtx{RMR^{-1}}$. Since the
set of $\{r_{ij}\}$ for which $\text{det} \mtx{R} = 0$ has probability
zero, a matrix $\mtx Q$ satisfying the above equation almost surely
exists. Thus $\mtx{A} = \mtx{RM} = \mtx{QR}$ and $\mtx{AA'} =
\mtx{QRR'Q'}$ Since the columns of A are not auto-correlated, $\mtx{AA'}
\sim W(\mtx{\Sigma}, T)$. Then $\mtx{RR'} \sim W(\mtx{Q^{-1} \Sigma
  Q'^{-1}}, T)$ follows from \cite{Anderson2003}, section 7.3.3. Now we
observe the Wishart \gls{pdf}
\begin{equation}
  \label{eq:WishartPDF}
  f_W(\mtx{X}) = { |\det X|^{(T-N-1)/2} \exp\left(-{1 \over 2}\tr
      \mtx{\Sigma^{-1}X} \right)
    \over
    2^{NT/2}\pi^{N(N-1)/4}|\det \mtx{\Sigma}|^{T/2}
    \prod_{i=1}^N \Gamma\left[(n+1-i)/2\right]
  }
\end{equation}
where, assuming $\mtx X = \mtx{UU'}$, $\mtx \Sigma$ is the covariance
matrix of the columns of $\mtx U$. we notice in \ref{eq:WishartPDF}
that $\mtx \Sigma$ enters $f_W(\cdot)$ through $\tr (\mtx{\Sigma^{-1}
  RR'})$ and $\det \mtx{\Sigma}$. $\mtx{RR'}$ enters through $\tr
\mtx{\Sigma^{-1} RR'}$ and $\det \mtx{RR'}$. Clearly
\begin{eqnarray*}
  \det \mtx{Q^{-1}\Sigma Q'^{-1}} &=& \det \mtx{\Sigma \over (\det Q)^2} \\
  &=& \det \mtx{\Sigma} \over (\det \mtx{M})^2 \\
  &=& \det \mtx{\Sigma}
\end{eqnarray*}
As to $\tr \mtx{\Sigma^{-1} C}$, we have
\begin{eqnarray*}
  && \tr\left[\mtx{(Q^{-1}\Sigma Q'^{-1})^{-1} RR'}\right] \\
  &=& \tr\left[\mtx{Q'\Sigma^{-1} Q RR'}\right] \\
  &=& \tr\left[\mtx{R'^{-1}M'R'\Sigma^{-1} RMR^{-1} RR'}\right] \\
  &=& \tr\left[\mtx{\Sigma^{-1} RM (RM)'}\right] \\
  &=& \tr\left[\mtx{\Sigma^{-1} AA'}\right] \\
\end{eqnarray*}

Moreover, $\det \mtx{RR'} = \det \mtx{AM^{-1} M'^{-1} A'} = \det
\mtx{AA'}$. Thus if we use $f_R(\cdot)$ to denote the joint
probability density of the entries of $\mtx{RR'}$ and $f_A(\cdot)$ to
denote that of $\mtx{AA'}$, we can write
\begin{equation}\label{eq:cross-corr-matrix-PDF}
  f_R(\mtx{RR'}) = f_A(\mtx{RMM'R'})
\end{equation}
Substituting for $f_A$ the Wishart \gls{pdf} \ref{eq:WishartPDF}, we
get the \gls{pdf} of $\mtx{RR'}$.

\chapter{Asymptotic Distributions of the Elements of a Covariance
  Matrix of Autocorrelated Gaussian Returns}
\label{chp:gaussian_elements_dist}
In this appendix we derive the approximate distributions of the
elements of a Covariance Matrix of autocorrelated Gaussian returns.

In the following we denote the diagonal elements as $C_{ii}$ and the
non-diagonal elements as $C_{ij}$ ($i \neq j$). We assume that
$\var(a_{i,t}) = \sigma^2$ for any $i$ and $t$, and $a_{i,t}$ are not
autocorrelated, i.e. $\text{corr}(a_{i,t}, a_{i, t'}) = 0$ for any
$i$, $t$ and $t'$. Then we can write
\begin{eqnarray*}
  r_{i,t} r_{j, t} &=& \left[
    \phi r_{i, t-1} + a_{i, t}
  \right] \left[
    \phi r_{j, t-1} + a_{j, t}
  \right] \\
  C_{ij} &=& {1 \over T}\sum_{t=1}^T r_{i,t} r_{j, t} \\
  &=& \phi^2 {1 \over T}\sum_{t=1}^Tr_{i,t-1} r_{j, t-1} +
  \phi{1 \over T}\sum_{t=1}^T\left(
    r_{i, t-1} a_{j, t} + r_{j, t-1} a_{i, t}
  \right) + {1 \over T}\sum_{t=1}^T a_{i,t} a_{j,t}
\end{eqnarray*}
We note that the two sums $\sum_{t=1}^T r_{i,t} r_{j, t}$ and
$\sum_{t=1}^T r_{i,t-1} r_{j, t-1}$ only differ by the first and the
last addend, which is negligible for sufficiently large T. Thus we
have
\begin{equation}\label{eq:Offdiag1}
  \begin{aligned}
    (1 - \phi^2)C_{ij} &\approx& \phi {1 \over T} \sum_{t=1}^T
    \left(r_{i, t-1} a_{j, t} + r_{j, t-1} a_{i, t} \right)
    + {1 \over T}\sum_{t=1}^T a_{i,t} a_{j,t}
  \end{aligned}
\end{equation}
Now we write the AR(1) process $r_{i,t}$ as an infinite moving-average
process:
\begin{eqnarray*}
  r_{i, t} &=& \phi r_{i, t-1} + a_{i,t} \\
  (1 - \phi B) r_{i, t} &=& a_{i,t} \\
\end{eqnarray*}
where $B$ is the back-shift operator. Then it follows from the above
equation
\begin{eqnarray*}
  r_{i,t} &=& {1 \over 1 - \phi B} a_{i,t} \\
  &=& \sum_{k=0}^\infty \phi^k B^k a_{i,t} \\
  &=& \sum_{k=0}^\infty \phi^k a_{i,t-k} \\
\end{eqnarray*}
where it is left implicit that $a_{i, t}$ with $t \leq 0$ is zero (in
words, this implies that the $r_{i,t}$ process is not affected at all
by events before $t = 1$).

Substituting this into eq.\ref{eq:Offdiag1} for $r_{i,t-1}$ yields
\begin{equation}
  \label{eq:Offdiag2}
  \begin{aligned}
  (1 - \phi^2)C_{ij} &\approx
  {1 \over T} \sum_{t=1}^T \sum_{k=0}^\infty\phi^{k+1} (a_{i, t-k-1}
  a_{j, t} + a_{j, t-k-1} a_{i, t})
  + {1 \over T} \sum_{t=1}^T a_{i,t} a_{j,t} \\
  &=
  {1 \over T} \sum_{t=1}^{T} \sum_{k=1}^{t-1}\phi^k (a_{i, t-k}
  a_{j, t} + a_{j, t-k} a_{i, t})
  + {1 \over T} \sum_{t=1}^{T} a_{i,t} a_{j,t}\\
  \end{aligned}
\end{equation}

Given two Gaussian random variables $x$ and $y$ with zero mean and
covariance matrix 
\begin{equation*}
  \Sigma = \sigma^2
  \begin{pmatrix}
    1 & \rho \\
    \rho & 1 \\
  \end{pmatrix}
\end{equation*}

The \gls{pdf} of $xy$ can be found by considering $P(xy < z)$:
\begin{eqnarray*}
  P(xy < z) &=& \left(\int_0^\infty dx \int_{-\infty}^{z/x} dy
    +\int_{-\infty}^0 dx \int_{z/x}^\infty dy \right)
    {1 \over 2\pi\sigma\sqrt{1 - \rho^2}} \\
    &&
    \exp\left[
      -{x^2 -2\rho xy + y^2
        \over
        2\sigma^2 (1 - \rho^2)} 
    \right] \\
  f(z; \sigma, \rho) &=& {d \over dz} P(xy < z)\\
  &=& {1 \over \pi \sigma \sqrt{1 - \rho^2}} \exp\left[
    {\rho z \over \sigma^2 (1 - \rho^2)}\right] K_0\left[
    {|z| \over \sigma^2 (1 - \rho^2)}
  \right]
\end{eqnarray*}
where $K_n(z')$ is the modified Bessel function of the second
kind. It is worth taking note that, when $\rho \neq 0$,
$f(z; \sigma, \rho)$ is not symmetric with respect to
$z$. As a result, if $\rho > 0$, the mean of $f(z;
\sigma, \rho)$ is positive, and vice versa.

Secondly, because $|\rho| < 1$ and
\begin{equation*}
  K_0(z) \sim \sqrt{\pi \over 2z} e^{-z}
\end{equation*}
as $z \to \infty$ \cite{Olver:2010:NHMF}, all the moments of
$f(z; \sigma, \rho)$ are finite, implying the
applicability of the Lyapunov central limit theorem provided that $T$
is large, which is very often the case and what we assume here.

With this in mind, we observe that $\phi$ only affects the first sum in
\ref{eq:Offdiag2}. If $a_{i,t-k}$ and $a_{j,t}$ are not correlated for
non-zero $k$, $\phi$ will not affect the mean of
$C_{ij}$. Furthermore, it is also clear from equation
\ref{eq:Offdiag2} that the variance of $C_{ij}$ is
always increased by a non-zero $\phi$, regardless of the sign of
$\phi$. We compute this increment in the following.

In light of the above expression for $f(z; \sigma,
\rho)$, we rewrite equation \ref{eq:Offdiag2} as
\begin{equation}\label{eq:Cij_dist}
  \begin{aligned}
    (1 - \phi^2)C_{ij} \approx &
    {1 \over T}
    \sum_{t=1}^{T} \sum_{k=1}^{t-1}\phi^k (a_{i, t-k} a_{j, t} + a_{j,
      t-k} a_{i, t}) \\
    & + {1 \over T} \sum_{t=1}^{T} \sigma^2 (1 - \rho^2) {a_{i,t} \over
      \sigma \sqrt{1 - \rho^2}} {a_{j,t} \over \sigma \sqrt{1 - \rho^2}} \\
  \end{aligned}
\end{equation}
In addition, we assume
\begin{equation*}
  \text{corr}(a_{i, t-k}, a_{j, t}) = \left\{
    \begin{array}{l l}
      1 & \text{ if $i=j$ and $k = 0$ }\\
      \rho & -1 < \rho < 1. \text{ if $i \neq j$ and $k = 0$ }\\
      0 & \text{otherwise}
    \end{array}
  \right.
\end{equation*}
Then, because $a_{i, t-k}$ and $ a_{j, t}$ with $i \neq j$ and $k
> 0$ are not correlated, the mean of $a_{i, t-k} a_{j, t}$ is 0
($f(z; \sigma, 0)$ is symmetric), and the variance of it
can be found to be $\sigma^6$ using formula (10.43.19) of \cite{NIST:DLMF}:
\begin{equation*}
  \int_0^\infty dt K_\nu(t) t^{\mu-1} = 2^{\mu-2}
  \Gamma\left(
    {\mu + \nu \over 2}
  \right) \Gamma\left(
    {\mu - \nu \over 2}
  \right)
\end{equation*}
On the other hand, $a_{i,t}/\sigma \sqrt{1 - \rho^2}$ and
$a_{j,t}/\sigma \sqrt{1 - \rho^2}$ have variance $1/(1 - \rho^2)$
and are correlated - $\text{corr}(a_{i,t}, a_{j,t}) = \rho$. The mean
of $a_{i,t}a_{j,t}/\sigma^2 (1 - \rho^2)$ can be found using formula
(10.43.22) of \cite{NIST:DLMF}, given that $-1 < \rho < 1$:
\begin{eqnarray*}
  \int_0^\infty t^{\mu - 1} e^{-at} K_\nu(t) dt &=& (\pi/2)^{1/2}
  \Gamma(\mu + \nu) \Gamma(\mu - \nu)(1 - a^2)^{-\mu/2 + 1/4} \times\\
  && P^{-\mu+1/2}_{\nu-1/2} (a)
\end{eqnarray*}
where $P^\mu_\nu(\cdot)$ is Ferrers function of the first
kind\footnote{ Ferrers function of the first kind is defined through
  the hypergeometric functon $F(a, b; c; z)$ \cite{NIST:DLMF}:
  \begin{equation}
    \label{eq:Ferrers_1st}
    \mathop{\mathsf{P}^{\mu}_{\nu}\/}\nolimits\!\left(x\right)=\left(\frac{1+x}{1-
        x}\right)^{\mu/2}\mathop{\mathbf{F}\/}\nolimits\!\left(\nu+1,-\nu;1-\mu;\tfrac
      {1}{2}-\tfrac{1}{2}x\right).
  \end{equation}
}. The result is
\begin{eqnarray*}
  \E\left[{a_{i,t}a_{j,t} \over \sigma^2 (1 - \rho^2)}\right]
  &=&
  {1 \over \sqrt{2\pi} (1 - \rho^2)^{5/4}} \left[
    P^{-3/2}_{-1/2}(-\rho) - P^{-3/2}_{-1/2}(\rho)
  \right]
\end{eqnarray*}
Similarly, the variance of $a_{i,t}a_{j,t}/\sigma^2 (1 - \rho^2)$ is
found to be
\begin{eqnarray*}
  v^2(\rho) &=&
  {4 \over \sqrt{2\pi} (1 - \rho^2)^{7/4}} \left[
    P^{-5/2}_{-1/2}(\rho) + P^{-5/2}_{-1/2}(-\rho)
  \right] \\
  && - \E^2\left[{a_{i,t}a_{j,t} \over
      \sigma^2 (1 - \rho^2)}\right]
\end{eqnarray*}
Now we can apply the Lyapunov central limit theorem
\cite{Billingsley1995} to the sum in equation \ref{eq:Cij_dist}
and write down the asymptotic Gaussian distribution of $C_{ij}$:
\begin{equation*}
C_{ij} \sim N(\mu'_X, \sigma'_X)
\end{equation*}
where
\begin{eqnarray}
  \mu'_X &=& {\sigma^2 \over \sqrt{2\pi} (1 - \phi^2)(1 -
    \rho^2)^{1/4}} \left[ P^{-3/2}_{-1/2}(-\rho) -
    P^{-3/2}_{-1/2}(\rho)
  \right] \label{eq:gaussian_mean}\\
  \sigma'^2_X &=& {1 \over (1 - \phi^2)^2}\left[
    \sum_{t=1}^T \sum_{k=1}^{t-1} 2\left(
      \phi^k \over T
    \right)^2 \sigma^6 + \sum_{t=1}^T
    {\sigma^4 (1 - \rho^2)^2 \over T^2} v^2(\rho)
  \right] \nonumber \\
  &=& {2 \sigma^6 \over T (1 - \phi^2)^2} \left[
    {\phi^2 \over 1 - \phi^2} -
    {\phi^2 (1 - \phi^{2T}) \over
      T(1 - \phi^2)}
  \right] + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
    T (1 - \phi^2)^2} \nonumber \\
  &\approx& {2 \sigma^6 \phi^2 \over T (1 - \phi^2)^3}
  + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
    T (1 - \phi^2)^2} \label{eq:gaussian_variance}
\end{eqnarray}
% Equation \ref{eq:gaussian_mean} tells that, if two return series $i$
% and $j$ are not correlated, auto-correlation in the returns does not
% introduce a bias into the estimation of the covariance; if,
% however, the return series are indeed correlated, auto-correlation
% in the returns rescales the covariance through a multiplicative
% factor $1/(1 - \phi^2)$.

% In addition, equation \ref{eq:gaussian_variance} tells that
% auto-correlation in the returns always makes the covariance
% estimation more noisy. Auto-correlation not only rescales the variance
% of the no-autocorrelation estimation by $1/(1 - \phi^2)$ but even
% adds an extra term ${2 \sigma^6 \phi^2 \over T (1 - \phi^2)^3}$.

We may apply a similar treatment to the diagonal elements of the
covariance matrix, which we denote as $C_{ii}$ here:
\begin{eqnarray*}
  C_{ii} &=& {1 \over T} \sum_{t=1}^T r^2_{it} \\
  &=& {1 \over T} \sum_{t=1}^T \sum_{l=0}^{t-1}\phi^l a_{i, t-l}
  \sum_{k=0}^{t-1}\phi^k a_{i, t-k} \\
  &=& {1 \over T} \sum_{t=1}^T \left[
    \sum_{k=0}^{t-1} \phi^{2k} a^2_{i, t-k} +
    \sum_{k,l = 0}^{t-1} \phi^{k+l} a_{i, t-k} a_{i, t-l}
  \right]
\end{eqnarray*}
By the Lyapunov central limit theorem under the assumption of large T,
the asymptotic distribution of $C_{ii}$ is Gaussian, the mean and
variance being
\begin{eqnarray}
  \E(C_{ii}) &=& {1 \over T}\left[
    \sum_{k=0}^{t-1} \phi^{2k} \sigma^2
  \right] \nonumber \\
  &=& {\sigma^2 \over (1 - \phi^2) T} \left[
    T - {\phi^2 (1 - \phi^{2T}) \over 1 - \phi^2}
  \right] \nonumber \\
  & \approx & {\sigma^2 \over 1 - \phi^2} \left[
    1 - {\phi^2 \over T}
  \right] \label{eq:gaussian_cii_mean}
\end{eqnarray}
and
\begin{eqnarray}
  \var(C_{ii}) &=& \sum_{t=1}^T \left[
    \sum_{k=0}^{t-1} {\phi^{4k} \sigma^4 \over T^2} 2 +
    \sum_{k,l=0}^{t-1} {\phi^{2(k+l)} \over T^2} \sigma^6
  \right] \nonumber \\
  &=& \sum_{t=1}^T \left[
    {2 \sigma^4 \over T^2} {1 - \phi^{4t} \over 1 - \phi^4} +
    {\sigma^6 \over T^2} \left(
      {1 - \phi^{2t} \over 1- \phi^2}
    \right)^2 \right] \nonumber \\
  &=& {2 \sigma^4 \over T (1 - \phi^4)} -
  {2 \sigma^4 \phi^4 (1 - \phi^{4T}) \over T^2(1 - \phi^4)^2} +
  \nonumber \\
  && {\sigma^6 \over T (1 -\phi^2)^2} -
  {2 \sigma^6 \phi^2 (1 - \phi^{2T}) \over T^2 (1 - \phi^2)^3} +
  {\sigma^6 \phi^4 (1 - \phi^{4T}) \over T^2 (1 - \phi^2)^2 (1 -
    \phi^4)} \nonumber \\
  &\approx& {2 \sigma^4 \over T (1 - \phi^4)} + {\sigma^6 \over T (1
    -\phi^2)^2} \label{eq:gaussian_cii_variance}
\end{eqnarray}
% Equation \ref{eq:gaussian_mean} tells that, if two return series $i$
% and $j$ are not correlated, auto-correlation in the returns does not
% introduce a bias into the estimation of the covariance; if,
% however, the return series are indeed correlated, auto-correlation
% in the returns rescales the covariance through a multiplicative
% factor $1/(1 - \phi^2)$.

% In addition, equation \ref{eq:gaussian_variance} tells that
% auto-correlation in the returns always makes the covariance
% estimation more noisy. Auto-correlation not only rescales the variance
% of the no-autocorrelation estimation by $1/(1 - \phi^2)$ but even
% adds an extra term ${2 \sigma^6 \phi^2 \over T (1 - \phi^2)^3}$.

\chapter{Eigenvalues Distribution of Wishart Matrix}
\label{sec:wishart_eigen_dist}
In this section we summarize the analytic results regarding the
eigenvalue distribution of a Wishart matrix $\mtx{RR'}$. In the
simplest case where the elements of $\mtx R$ are all independent of
each other, i.e. neither auto-correlation nor cross-correlation
exists, we have \cite{Chiani2012}
\begin{eqnarray}
  f(x_1, \cdots, x_N) &=& K \prod_{i=1}^N e^{-x_i/2}
  x_i^{(T-N-1)/2} \prod_{i<j}^N (x_i -
  x_j) \label{eq:wishart_eigen_pdf}
\end{eqnarray}
where the eigenvalues have been indexed in descending order: $x_1 \geq
x_2 \geq \cdots \geq x_N$. The normalization constant K is given by
\begin{equation*}
  K = {\pi^{N^2/2} \over 2^{NT/2}\Gamma_N(T/2) \Gamma_N(N/2)}
\end{equation*}
where the function $\Gamma_m(a)$ is defined as
\begin{equation*}
  \Gamma_m(a) = \pi^{m(m-1)/4} \prod_{k=1}^m \Gamma\left(a - {k-1
      \over 2}\right)
\end{equation*}
If the order of the eigenvalues is ignored, the \gls{pdf} of their
distribution is given by the Marcenko-Pastur law \cite{Guhr2007}
\begin{eqnarray}
  f(x) &=& {1 \over 2\pi \sigma^2 q} {
    \sqrt{(x_2 - x)(x - x_1)} \over x
  } \label{eq:MP_pdf}
\end{eqnarray}
where it is assumed $r_{it} \sim N(0, \sigma^2)$ and
\begin{eqnarray*}
  q &=& \lim_{N,T \to \infty} {N \over T} \\
  x_1 &=& \sigma^2 (1 - \sqrt q)^2 \\
  x_2 &=& \sigma^2 (1 + \sqrt q)^2 \\
\end{eqnarray*}
It is imposed that $q = N/T < 1$.

Moreover, K. Johnsson \cite{Johnsson2000} and I. Johnstone
\cite{Johnstone2001} showed that, at the {\it absence} of
autocorrelations and in the asymptotic limit $N, T \to \infty$, $N/T
\to q < \infty$, the maximum eigenvalue  $\lambda_1$ follows the
Tracy-Widom distribution (denote $\mathscr{TW}_1$ here) when properly
relocated and rescaled:
\begin{equation*}
  {\lambda_1 - \mu_{NT} \over \sigma_{NT}} \sim \mathscr{TW}_1
\end{equation*}
where $\mu_{NT}$ and $\sigma_{NT}$ are given by
\begin{eqnarray*}
  \mu_{NT} &=& \left(
    \sqrt{N-1/2} + \sqrt{T - 1/2}
  \right)^2 \\
  \sigma_{NT} &=& \sqrt{\mu_{NT}} \left(
    {1 \over \sqrt{N-1/2}} + {1 \over \sqrt{T-1/2}}
  \right)^{1/3}
\end{eqnarray*}
The $\mathscr{TW}_1$ distribution has the following cummulative
distribution function (CDF) \cite{Chiani2012}
\begin{equation*}
  F_1(x) = \exp\left[
    -{1 \over 2} \int_x^\infty dy \left(q(y) + (y-x)q^2(y)\right)
  \right]
\end{equation*}
where $q(y)$ is defined as the solution to the Painlev\'e II differential
equation
\begin{equation*}
  q''(y) = yq(y) + 2q^3(y)
\end{equation*}
which is unique when imposing the condition
\begin{equation*}
  q(y) \sim \text{Ai}(y) \text{ as } y \to \infty
\end{equation*}

Then Marco Chiani showed recently that the $\mathscr{TW}_1$ distribution
can be well approximated by a gamma distribution based on his proof
that the exact distribution of the maximum eigenvalue is a mixture of
gamma distributions. Specifically,
\begin{equation}\label{eq:TracyWidom-Gamma}
  {\lambda_1 - \mu_{NT} \over \sigma_{NT}} + \alpha \sim
  \mathscr{G}(k ,\theta)
\end{equation}
where $\mathscr{G}(k, \theta)$ denotes the Gamma distribution with
parameters $k$ and $\theta$ \cite{Chiani2012}. The \gls{pdf} of the
gamma distribution is given by
\begin{equation*}
  f_\gamma(x; k, \theta) = {1 \over \Gamma(k) \theta^k} x^{k-1}
  e^{-x/\theta}
\end{equation*}
Moreover, the first 3 moments of the distribution are simple:
\begin{eqnarray*}
  \text{mean} &=& k\theta \\
  \text{variance} &=& k\theta^2 \\
  \text{skewness} &=& 2/\sqrt{k}
\end{eqnarray*}

% \chapter{Collection of Figures and Tables}

% \begin{figure}[htb!]
%   \centering
%   \subfigure[ACF of $z_t$]{
%     \includegraphics[scale=0.4, clip=true, trim=95 236 118
%     200]{../pics/nordea_15min_quotient_acf.pdf}
%     \label{fig:nordea_15min_quotient_acf}
%   }
%   \subfigure[ACF of $z_t^2$]{
%     \includegraphics[scale=0.4, clip=true, trim=95 236 118
%     200]{../pics/nordea_15min_quotient_squared_acf.pdf}
%     \label{fig:nordea_15min_quotient_squared_acf}
%   }
%   \caption{\small \it Nordea 15min $z_t$ and $z_t^2$ auto-correlation
%     function (ACF).}
% \end{figure}

% \begin{figure}[htb!]
%   \centering
%   \subfigure[Normal Probability Plot of $y_t$]{
%     \includegraphics[scale=0.4, clip=true, trim=76 256 109
%     214]{../pics/nordea2_y_normplot.pdf}
%     \label{fig:nordea_15min_y_qq}
%   }
%   \subfigure[ACF of $y_t$]{
%     \includegraphics[scale=0.4, clip=true, trim=92 258 104
%     218]{../pics/nordea_15min_y_acf.pdf}
%     \label{fig:nordea_15min_y_acf}
%   }
%   \caption{\small \it Nordea 15min $y_t$ normal probability plot and
%     autocorrelations}
% \end{figure}