### 2014-12-01 thesis/draft3.tex

\documentclass{article}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{amssymb}
\usepackage{amsfonts, amsmath}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{enumerate}
% \usepackage[bookmarks=true]{hyperref}
% \usepackage[acronym]{glossaries}
%\usepackage{bookmark}
\input{../physics_common.tex}

\DeclareGraphicsExtensions{.eps, .pdf}

\begin{document}
\section{Spectra with same q and different $T$}
The following figure shows the spectra of correlation matrices with
fixed ratio $q = T/N$, and varying $T$ and $N$.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4]{../pics/spectra_q10.eps}
  \caption{\small \it Spectra with q = 10}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.3]{../pics/spectra_q24.eps}
  \caption{\small \it Spectra with q = 24}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.3]{../pics/spectra_q50.eps}
  \caption{\small \it Spectra with q = 50}
\end{figure}

Apparently the spectral density is changed even when the ratio $q$ is left
unchanged by varying $N$ and $T$. For this matter, I notice that for
Burda et al's weighted covariance matrix,
\begin{eqnarray*}
  C &=& {1 \over T} \mtx{RWR'}
\end{eqnarray*}
the spectral density also has $T$ dependence via the parameter $\beta$:
\begin{eqnarray*}
  \beta &=& T(1 - \alpha)
\end{eqnarray*}
where $\alpha$ in their IGARCH(1) model is analogous to $\beta_1$ in a
GARCH(1, 1) model.

\section{Why large deviation from MP at small $\alpha_1$}
Figure \ref{fig:Distance_to_MP} shows that when $\alpha_1$ is small,
deviation of the spectral density from the MP law is large with all
values of the tail exponent $\alpha$. Explanation to this can be found
in the autocorrelations of a GARCH(1,1) process.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.6]{../pics/Distance_to_MP.eps}
    \caption{\small \it Distance of the empirical spectral density to
      the Marchenko-Pastur law measured in terms of the mean square
      error between the two density functions}
    \label{fig:Distance_to_MP}
\end{figure}

A detailed calculation reveals the following autocorrelation
structure for a GARCH(1,1) process:
\begin{eqnarray*}
  \rho_n = \text{corr}(\epsilon_t^2, \epsilon_{t-n}^2) &=& (\alpha_1
  + \beta_1)^{n-1} \rho_1 \\
  \rho_1 = \text{corr}(\epsilon_t^2, \epsilon_{t-1}^2) &=& \alpha_1 +
  \beta_1 - \frac{\beta_1 [1 - (\alpha_1 + \beta_1)^2]}{
    1 - (\alpha_1 + \beta_1)^2 + \alpha_1^2
  }
\end{eqnarray*}
where $n \geq 2$. Figure \ref{fig:autocorr_1} and
\ref{fig:autocorr_1_full} shows $\rho_1$ as a function of $\alpha_1$
and $\beta_1$.
\begin{figure}[htb!]
  \centering
  \subfigure[$\rho_1 = \text{corr}(\epsilon_t^2, \epsilon_{t-1}^2)$ as
  a function of $\alpha_1$ and $\beta_1$. The plot is restricted to a
  region of small $\alpha_1$ and large $\beta_1$ such that $\alpha_1 +
  \beta_1 \approx 1$.]{
    \includegraphics[scale=0.8]{../pics/autocorr_1.pdf}
    \label{fig:autocorr_1}
  }
  \subfigure[$\rho_1 = \text{corr}(\epsilon_t^2, \epsilon_{t-1}^2)$ as
  a function of $\alpha_1$ and $\beta_1$. The plot covers the entire
  region of $\alpha_1 > 0$, $\beta_1 > 0$ and $\alpha_1 + \beta_1 <
  1$. Please ignore the spike, which is due to numerics.]{ 
    \includegraphics[scale=0.5]{../pics/autocorr_1_full.pdf}
    \label{fig:autocorr_1_full}
  }
\end{figure}
It is seen that, as $\alpha + \beta \to 1$, $\rho_1 \to 1$ even if
$\alpha_1$ is rather small. On a curve of constant-$\alpha$, as
$\alpha_1 \to 0$, $\alpha + \beta \to 1$ and hence $\rho_1 \to
1$. Thus for a fixed $\alpha$, very small $\alpha_1$ implies very
strong auto-correlation and hence large deviation of the spectral
density from the MP law.
\end{document}