### 2015-01-04 cv/References.tex
Add CV and cover letters.

\title{References}
% \date{\today}
\author{
Xie Xiaolei
}

\documentclass[12pt]{article}

\begin{document}
\maketitle

My recommenders are:
\begin{itemize}
\item Sven \AA berg, professor and Head of Mathematical Physics at
  Lund University
\item Ferdi Aryasetiawan, professor of Mathematical Physics at
  Lund University
\end{itemize}

\end{document}