### 2015-01-04 cv/ResearchCoverLetter.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Long Lined Cover Letter
% LaTeX Template
% Version 1.0 (1/6/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Matthew J. Miller
% http://www.matthewjmiller.net/howtos/customized-cover-letter-scripts/
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10pt,stdletter,dateno,sigleft]{newlfm} % Extra options: 'sigleft' for a left-aligned signature, 'stdletternofrom' to remove the from address, 'letterpaper' for US letter paper - consult the newlfm class manual for more options

\usepackage{charter} % Use the Charter font for the document text

% \newsavebox{\Luiuc}\sbox{\Luiuc}{\parbox[b]{1.75in}{\vspace{0.5in}
%\includegraphics[width=1.2\linewidth]{logo.png}}} % Company/institution logo at the top left of the page
% \makeletterhead{Uiuc}{\Lheader{\usebox{\Luiuc}}}

\newlfmP{sigsize=10pt} % Slightly decrease the height of the signature field
%\newlfmP{addrfromphone} % Print a phone number under the sender's address
%\newlfmP{addrfromemail} % Print an email address under the sender's address
% \PhrPhone{Phone} % Customize the "Telephone" text
% \PhrEmail{Email} % Customize the "E-mail" text

% \lthUiuc % Print the company/institution logo

%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\namefrom{Xie Xiaolei} % Name

\addrfrom{
%\today\\[12pt] % Date
% Kiviksgatan 7C \\ % Address
% 21440 Malm\"o
% Sweden
}

% \phonefrom{+46 70 489 0472} % Phone number

% \emailfrom{xie.xiaolei@gmail.com} % Email address

%----------------------------------------------------------------------------------------
%	ADDRESSEE AND GREETING/CLOSING
%----------------------------------------------------------------------------------------

\greetto{Dear Sir or Madam,} % Greeting text
\closeline{Sincerely Yours,} % Closing text

\nameto{
% Evaluation Committee
} % Addressee of the letter above the to address

\addrto{
% Recruitment Officer \\ % To address
% The Corporation \\
% 123 Pleasant Lane \\
% City, State 12345
}

%----------------------------------------------------------------------------------------

\begin{document}
\begin{newlfm}

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------

I understand from your department's website that you are offering Ph.D
stipends, and found it a truly exciting opportunity. I have hence
contacted Prof. Thomas Valentin Mikosch and asked for his
supervision. After a kind discussion with me about my master's 
thesis, his research area, and a very interesting Ph.D project in
random matrix theory (RMT) and extreme value theory (EVT), he
encouraged me to apply. The project that he outlined perfectly
continued my master's studies and I really look forward to it. So I
would like to present to you what I did as my master's thesis project
and what my research interests are.

I wrote my thesis under the supervision of Prof. Sven \AA berg at Lund
university in the department of mathematical physics. I first compared
GARCH and stochastic volatility (SV) models by case studying of
intraday returns of a few Swedish stocks, and then I continued to
study the covariance matrix of Gaussian and GARCH(1,1) returns. There
I encountered the problem of eigenvalues distribution of a Wishart
matrix and found its large and charming literature.

By numerical simulation, I studied in the cases of Gaussian and
GARCH(1,1) returns, how the matrix elements and eigenvalues were
distributed, how the eigenvectors were composed, and in particular,
how the situation was changed by autocorrelations in the return
series. For example, I found in the case of GARCH(1,1) returns that
prolonged return series increased the extent of the eigenvectors'
localization while strengthened autocorrelations decreased the extent
of localization.

Among the many fascinating things that I was led to by the thesis
project were the extreme value theory and the random matrix
theory. In my studies of the intraday stock returns, heavy-tailed
distributions, and in particular L\'evy distributions, are very
important, so I read books such as ``Modeling Extremal Events'',
``Handbook of Financial Time Series'', and papers about the tail
behavior of GARCH and SV models, as well as those on the eigenvalues
distribution of random matrices. Quite naturally, these books and
papers led me to Prof. Mikosch. So when I met him and he outlined the
Ph.D project on RMT and EVT, which followed directly from my master's
project, I knew it was truly a great opportunity for me.

If I am so honored to be accepted, I will, first of all, learn more
about EVT in connection with multivariate time 
series, and also about its application in the research of RMT. Then I
would like to explore, for example, if and how the methods developed in the
research of the largest eigenvalue distribution and the spectral
density of random matrices, particularly Wishart matrices, can
be applied, presumably with EVT, to find the eigenvalues distribution
of a covariance matrix of heavy-tailed returns, such as those generated
by GARCH or SV models.

In particular, I understand that the problem of the exact
distribution of the largest eigenvalue of a Wishart matrix has been
attacked by the holonomic gradient method as an alternative to the
more conventional approach of series expansion around the origin. It
would be interesting to see if the same methods are applicable to the
aforementioned problem.

I look forward to an interview with you. If you would like to know more
about me, you are most welcome to call me at +46 70 489 0472 or find
me on LinkedIn as ``Xie Xiaolei''. If you want to contact my thesis
supervisor, Prof. Sven \AA berg, he is reachable via telephone +46 46
222 96 33 or mobile +46 70 557 35 69.

\end{newlfm}
\end{document}