### 2015-01-04 cv/XieXiaolei-Computing.tex

% Jason R. Blevins - Curriculum Vitae
%
% Copyright (C) 2004-2010 Jason R. Blevins
% http://jblevins.org/projects/cv-template/
%
% You may use use this document as a template to create your own CV
% and you may redistribute the source code freely. No attribution is
% required in any resulting documents. I do ask that you please leave
% this notice and the above URL in the source code if you choose to
% redistribute this file.

\documentclass[10pt,letterpaper]{article}

\usepackage{hyperref}
\usepackage{geometry}
\usepackage[T1]{fontenc}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
% Comment the following line to use the default Computer Modern font
% instead of the Palatino font provided by the mathpazo package.
% Remove the 'osf' bit if you don't like the old style figures.
%\usepackage[sc,osf]{mathpazo}

% In practice, I use the following font packages instead of mathpazo.
% beramono provides a nice fixed-width font. xagaramon uses the
% (commercial) Adobe Garamond font.
%\usepackage[scaled=0.75]{beramono}
%\usepackage[osf]{xagaramon}

% Set your name here
\def\name{Xie Xiaolei}

% The following metadata will show up in the PDF properties
\hypersetup{
  colorlinks = true,
  urlcolor = black,
  pdfauthor = {\name},
  pdfkeywords = {computer, mobile, information security},
  pdftitle = {\name: Curriculum Vitae},
  pdfsubject = {Curriculum Vitae},
  pdfpagemode = UseNone
}

\geometry{
  body={6.5in, 9.0in},
  left=1.0in,
  top=1.0in
}

% Customize page headers
\pagestyle{myheadings}
\markright{\name}
\thispagestyle{empty}

% Custom section fonts
\usepackage{sectsty}
\sectionfont{\ttfamily\mdseries\Large}
\subsectionfont{\ttfamily\mdseries\itshape\large}

% Other possible font commands include:
% \ttfamily for teletype,
% \sffamily for sans serif,
% \bfseries for bold,
% \scshape for small caps,
% \normalsize, \large, \Large, \LARGE sizes.

% Don't indent paragraphs.
\setlength\parindent{0em}

% Make lists without bullets and compact spacing
%\renewenvironment{itemize}{
\newenvironment{aperiod}{
  \begin{list}{}{
    \setlength{\leftmargin}{1.5em}
    \setlength{\itemsep}{1em}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0.25em}
  }
}{
  \end{list}
}

\begin{document}

% Place name at left
{\huge \name}

% Alternatively, print name centered and bold:
%\centerline{\huge \bf \name}

\vspace{0.25in}

\begin{minipage}[t]{0.5\textwidth}
  Citizenship: Sweden \\
  Phone: +46 70 489 0472 \\
  Email: \href{mailto:xie.xiaolei@gmail.com}{\tt xie.xiaolei@gmail.com} \\
  Address: Kiviksgatan 7C, 21440 Malm\"{o}, Sweden
\end{minipage}

\section*{Education}
\begin{itemize}
\item Master of Physics, Lund University, January 2011 - June 2014.
  \begin{itemize}
  \item {\it Major Subject:} Econophysics. My thesis project was
    case studying of intraday returns of Nordea, Volvo and
    Ericsson. Covariance matrices of GARCH and stochastic volatility
    returns were also studied. Matlab programming was a crucial part.
  \end{itemize}
\item M.Sc.(Tech) Computer Science and Engineering, Helsinki
  University of Technology, 2005 - 2008.
  \begin{itemize}
  \item {\it Major Subject:} Mobile Computing - Services and Security
  \item {\it Average grade:} 4.17/5.
  \end{itemize}
\item B.Sc. Computer Science \& Technology, Beijing Institute of
  Technology, 2001- 2005.
  \begin{itemize}
  \item {\it Average grade:} 85.5/100.
  \item {\it Honors:} \textit{Distinguished Graduates of 2005}.
  \end{itemize}
\end{itemize}

\section*{Employment}
\begin{aperiod}
\item {\it \href{http://www.stericsson.com}{\tt
      ST-Ericsson}, Audio, Multimedia, Development Engineer, Sep 2008
    - Jan 2011}

  My responsibilities included developing audio low-level software,
  typically device drivers, in the Linux OS. This involved designing and
  implementing data structures and procedures so that the hardware
  devices were used in the optimal way, in terms of parallelism,
  power consumption, robustness, and ease of maintenance.

\item {\it \href{http://www.hiit.fi}{\tt Helsinki
      Institute of Information Technology}, Thesis Worker, Oct 2007 -
    May 2008}

  Funded by HIIT, I wrote my thesis ``A Mutual authentication and Key
  Agreement Protocol for the UMTS Network''. The algorithm of my
  protocol proposal was based on the Diffie-Hellman Exchange, but had
  significant modifications and extensions so that the mutual
  authentication between the mobile user and the network was delegated
  to the serving instead of the home network. The protocol design
  incorporated puzzle schemes that demanded unbalanced computing
  resources on the two sides to mitigate Denial-of-Service attacks. I
  made C programs to test the puzzle schemes.

\item {\it
    \href{http://www.goodmillsystems.com}{\tt Goodmillsystems},
    Software Designer, May 2007 - Sep 2007}

  I developed a daemon program that ran on the company's smart router
  product to collect statistical data about the router's network
  traffic. The emphasis of the program was high performance in the
  context of competing read and write accesses.

\item {\it \href{http://www.cardinal.fi/}{\tt
      Cardinal Information Systems}, Software Developer, Sep 2006 -
    April 2007}

  I implemented the FLUTE protocol for the company's DVB-H
  mobile-end middleware. The implementation was multi-threaded and was
  built on the {\it pthread} library. Good performance and robustness
  were crucial requirements on the implementation.

\item {\it
    \href{http://www.digitalchocolate.com/}{\tt Digital Chocolate},
    Game Engineer, Feb 2006 - Sep 2006}

  At this job, my responsibilities were porting and customizing mobile
  phone games for various handsets, and making game demos.

\item {\it {\tt Beijing Holystone Ltd}, Software
    Engineer, April 2004 - June 2005}

  At this job I mainly developed JSP pages and database routines for
  the company's information system.
\end{aperiod}

\section*{Personal Project}
{\it Trades data collector \& Analyzer}

I have been developing my own stock-trading robot since the summer of
2011. It is a software built on cURL, GLib, MySQL, and Matlab. I find
it a fairly interesting, challenging, and rewarding combination of
programming skills, data analysis techniques, and knowledge of
financial time series. My knowledge in econophysics guides the
development of this software.

\section*{Non-English Language Skills}
\begin{itemize}
\item Swedish: informal communications. ``Swedish B'' passed.
\item Chinese: Native speaker
\end{itemize}

\end{document}