### 2015-01-04 cv/XieXiaolei.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Two Column One Page Curriculum Vitae
% LaTeX Template
% Version 1.1 (24/1/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Alessandro (The CV Inn)
%
% IMPORTANT: THIS TEMPLATE NEEDS TO BE COMPILED WITH XeLaTeX
%
% This template uses several fonts not included with Windows/Linux by
% default. If you get compilation errors saying a font is missing, find the line
% on which the font is used and either change it to a font included with your
% operating system or comment the line out to use the default font.
% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{wasysym}
\usepackage{marvosym} % Required for symbols in the colored box
\usepackage{ifsym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
\setromanfont[Mapping=tex-text]{Gentium} % Main document font
\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Century Schoolbook
  L} % Font for your name at the top
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\color{text1} % Sets the default text color for the whole document to the color defined as 'text1'

%----------------------------------------------------------------------------------------
%	TITLE
%----------------------------------------------------------------------------------------

\par{\centering{\sffamily\Huge Xie Xiaolei}\\ % Your name
{\color{headings}\fontspec{Century Schoolbook L} Curriculum {Vit\fontspec{Century Schoolbook L}\ae}\\[15pt]\par} % Curriculum vitae text in the Century Schoolbook L font

%----------------------------------------------------------------------------------------

\begin{minipage}[t]{0.5\textwidth} % Start the left-hand side of the page
%\vspace{0pt} % Trick for alignment
%----------------------------------------------------------------------------------------
%	EDUCATION
%----------------------------------------------------------------------------------------

\section{Education} 
\begin{itemize} 
\item \textit{Lund University, Sweden}

\begin{tabular}{l p{0.7\textwidth}}
  2011 -- \textsc{Present} & {M.Sc. Econophysics: Mathematical
  Physics in Finance} \\
  Thesis title & Returns Models and Their Covariance Matrix \\
  Description & Compared accuracy of volatility forecasts of {\bf
    GARCH} and {\bf stochastic volatility} (SV) models by case
  studying Nordea, Volvo and Ericsson {\bf intraday} returns using the
  two classes of models. Realized volatilities were computed and
  compared to the forecasts. Also studied eigenvalues' distribution of
  a covariance matrix of GARCH returns. {\bf MATLAB} programming was
  an essential part.
\end{tabular}

\item {\it Helsinki University of Technology, Finland}

\begin{tabular}{l p{0.7\textwidth}}
  2005 -- 2008 & {\bf M.Sc. Computer Science} \\
  Thesis Title & A Mutual Authentication and Key Agreement Protocol
  for the UMTS Network \\
  Description & The algorithm of the proposed protocol was based on
  the Diffie-Hellman Exchange, but had significant modifications and
  extensions so that the mutual authentication between the mobile user
  and the network was delegated to the serving instead of the home
  network. The protocol design incorporated puzzle schemes that
  demanded unbalanced computing resources on the two sides to mitigate
  Denial-of-Service attacks. \\
  Average Grade & 4.17 / 5
\end{tabular}

\item {\it Beijing Institute of Technology, China}

\begin{tabular}{l p{0.7\textwidth}}
  2001 -- 2005 & B.Sc. Computer Science \& Technology \\
  Average Grade & 85.5 / 100 \\
  Honor & Distinguished Graduates of 2005
\end{tabular}
\end{itemize}

\section{Personal Project}
{\it Trades data collector and analyzer, June 2008 -- present}

Developed a Linux daemon program (programmed in C \& Matlab) to
collect and analyze realtime transaction records on the Stockholm
stock exchange. The program maintained and evolved stochastic
volatility models and made volatility forecast in realtime. Collected
data were managed with a MySQL database.

\section{Non-English Language Skills}
\begin{itemize}
\item Swedish: Informal communications
\item Chinese: Native speaker
\end{itemize}

% \section{Non-English Language Skills}
% \begin{tabular}{l p{0.7\textwidth}}
%   Swedish & No problem with informal communications of talking,
%   reading and writing. ``Swedish B'' passed.\\
%   Chinese (Mandarin) & Native speaker \\
% \end{tabular}

%----------------------------------------------------------------------------------------

\end{minipage} % End the left-hand side of the page
\hfill
\begin{minipage}[t]{0.44\textwidth} % Start the right-hand side of the page
\vspace{0pt} % Trick for alignment

%----------------------------------------------------------------------------------------
%	COLORED BOX
%----------------------------------------------------------------------------------------

\colorbox{shade}{\textcolor{text1}{
\begin{tabular}{c|p{7cm}}
\raisebox{-1pt}{\Bowtie} & Swedish citizen \\
\raisebox{-2pt}{\textifsymbol{18}} & Kiviksgatan 7C, 21440 Malm\"o, Sweden \\ % Address
\raisebox{-2pt}{\Mobilefone} & +46 70 489 0472 \\ % Phone number
\raisebox{-1pt}{\Letter} & \href{mailto:xie.xiaolei@gmail.com}{xie.xiaolei@gmail.com} \\ % Email address
\Keyboard & \href{http://www.facebook.com/xie.xiaolei}{http://www.facebook.com/xie.xiaolei} \\ % Website
\end{tabular}
}}\\[10pt]

%----------------------------------------------------------------------------------------
%	WORK EXPERIENCE
%----------------------------------------------------------------------------------------

\section{Work Experience}

%------------------------------------------------
% WORK EXPERIENCE 1
%------------------------------------------------

{\raggedleft\textsc{Sep 2008 -- Jan 2011}\par}

{\raggedright\large Development Engineer at \textsc{ST-Ericsson}, Lund\\
  \textit{Software Development}\\[5pt]}

\normalsize{Developed audio low-level software, typically device
  drivers, in the Linux OS.}\\

%------------------------------------------------
% WORK EXPERIENCE 2
%------------------------------------------------

{\raggedleft\textsc{May 2007 -- Sep 2007}\par}

{\raggedright\large Software Designer at \textsc{Goodmillsystems}, Helsinki\\
  \textit{Software Development}\\[5pt]}

\normalsize{Developed a daemon program that ran on the company's
  smart router product for network traffic monitoring.}\\

%------------------------------------------------
% WORK EXPERIENCE 3
%------------------------------------------------

{\raggedleft\textsc{Sep 2006 -- April 2007}\par}

{\raggedright\large Software Developer at \textsc{Cardinal Information
    Systems}, Helsinki\\
  \textit{Software development}\\[5pt]}

\normalsize{Implemented the FLUTE protocol for the company's DVB-H mobile-end
  middleware.}\\

%------------------------------------------------
% WORK EXPERIENCE 4
%------------------------------------------------
{\raggedleft\textsc{Feb 2006 - Sep 2006}\par}

{\raggedright\large Game Engineer at \textsc{Digital Chocolate}, Helsinki\\
  \textit{Java  mobile game development}\\[5pt]}

\normalsize{Ported and customized mobile phone games for various
  handsets. Also developed some general tools for game customization.}\\

%------------------------------------------------
% WORK EXPERIENCE 5
%------------------------------------------------
{\raggedleft\textsc{April 2004 - June 2005}\par}

{\raggedright\large Software Engineer at \textsc{Holystone Ltd}, Beijing\\
  \textit{Java Software development}\\[5pt]}

\normalsize{Developed JSP pages and database routines for the
  company's information system.}\\

\section{Programming Skills}
\begin{itemize}
\item Very experienced in C, Matlab
\item Experienced in C++, Python, Java, SQL
\item Very familiar with Linux
\end{itemize}

\end{minipage} % End right-hand side of the page

\end{document}