### 2015-01-05 cv/SecondApplicationToKU.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Long Lined Cover Letter
% LaTeX Template
% Version 1.0 (1/6/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Matthew J. Miller
% http://www.matthewjmiller.net/howtos/customized-cover-letter-scripts/
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10pt,stdletter,dateno,sigleft]{newlfm} % Extra options: 'sigleft' for a left-aligned signature, 'stdletternofrom' to remove the from address, 'letterpaper' for US letter paper - consult the newlfm class manual for more options

\usepackage{charter} % Use the Charter font for the document text

% \newsavebox{\Luiuc}\sbox{\Luiuc}{\parbox[b]{1.75in}{\vspace{0.5in}
%\includegraphics[width=1.2\linewidth]{logo.png}}} % Company/institution logo at the top left of the page
% \makeletterhead{Uiuc}{\Lheader{\usebox{\Luiuc}}}

\newlfmP{sigsize=10pt} % Slightly decrease the height of the signature field
%\newlfmP{addrfromphone} % Print a phone number under the sender's address
%\newlfmP{addrfromemail} % Print an email address under the sender's address
% \PhrPhone{Phone} % Customize the "Telephone" text
% \PhrEmail{Email} % Customize the "E-mail" text

% \lthUiuc % Print the company/institution logo

%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\namefrom{Xie Xiaolei} % Name

\addrfrom{
%\today\\[12pt] % Date
% Kiviksgatan 7C \\ % Address
% 21440 Malm\"o
% Sweden
}

% \phonefrom{+46 70 489 0472} % Phone number

% \emailfrom{xie.xiaolei@gmail.com} % Email address

%----------------------------------------------------------------------------------------
%	ADDRESSEE AND GREETING/CLOSING
%----------------------------------------------------------------------------------------

\greetto{Dear Sir or Madam,} % Greeting text
\closeline{Sincerely Yours,} % Closing text

\nameto{
% Evaluation Committee
} % Addressee of the letter above the to address

\addrto{
% Recruitment Officer \\ % To address
% The Corporation \\
% 123 Pleasant Lane \\
% City, State 12345
}

%------------------------------------------------

\begin{document}
\begin{newlfm}
I am currently a self-financing Ph.D student under the supervision of
Prof. Thomas Mikosch. I understand that the department is providing
free stipends to Ph.D students, so I would like to apply for a grant.

I started the Ph.D program here on 1st of October and have been taking
the courses {\it Extreme Value Theory} given by Prof. Mikosch and {\it
  Quantitative Risk Management} given by Prof. Collamore. At the same
time I have done analysis on the eigenvalues of empirical covariance
matrices of the S\&P500 and the DAX indices. A detailed plan of my
Ph.D program is submitted along with this application.

I obtained my master's degree in mathematical physics from Lund
University, with a thesis on econophysics supervised by Prof. Sven \AA
berg. Based on the results of the thesis and with some continued work,
Prof. \AA berg and I have written a paper together under the title
``Effect of Heteroscedasticity on the spectrum of an Empirical
Covariance Matrix'' and plan to submit it to {\it Quantitative
  Finance} for publication. In the paper we argued that, on certain
conditions, autocorrelations in the conditional volatilities could
affect the empirical covariance matrixes' spectrum only through the
unconditional distribution of the volatilities. On these
conditions, we derived in the case of small autocorrelations the
spectrum of an empirical covariance matrix. The derivations were based
on results from free probability theory.

Without a stipend, I am living on my own savings and borrowed money
from my relatives. This admittedly causes hardships and
distractions. So I really hope a stipend can be granted to me. It
would be a great ease not only to me but also to my family. Thank you
for your consideration.

\end{newlfm}
\end{document}