### 2015-01-06 EVT/MidTerm.tex

%% Created by Maple 17.00, Linux
%% Source Worksheet: MidTerm.mw
%% Generated: Mon Jan 05 22:30:46 CET 2015
\documentclass{article}
\usepackage{maplestd2e}
\def\emptyline{\vspace{12pt}}
\begin{document}
\pagestyle{empty}
\DefineParaStyle{Maple Heading 1}
\DefineParaStyle{Maple Text Output}
\DefineParaStyle{Maple Dash Item}
\DefineParaStyle{Maple Bullet Item}
\DefineParaStyle{Maple Normal}
\DefineParaStyle{Maple Heading 4}
\DefineParaStyle{Maple Heading 3}
\DefineParaStyle{Maple Heading 2}
\DefineParaStyle{Maple Warning}
\DefineParaStyle{Maple Title}
\DefineParaStyle{Maple Error}
\DefineCharStyle{Maple Hyperlink}
\DefineCharStyle{Maple 2D Math}
\DefineCharStyle{Maple Maple Input}
\DefineCharStyle{Maple 2D Output}
\DefineCharStyle{Maple 2D Input}
\begin{center}
\begin{Maple Normal}{
\begin{center}
\begin{Maple Normal}{
\textbf{Mid Term Test}}\end{Maple Normal}
\end{center}
}\end{Maple Normal}
\end{center}
\begin{center}
\begin{Maple Normal}{
\begin{center}
\begin{Maple Normal}{
\textbf{Xie Xiaolei}}\end{Maple Normal}
\end{center}
}\end{Maple Normal}
\end{center}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{1.(a) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
For sufficiency, assume\mapleinline{inert}{2d}{t <= F(x)}{$\displaystyle t\leq F \left( x \right) $}
. Because}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&larr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em")), superscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("inf", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("s", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&in;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("R", italic = "true", mathvariant = "italic"), Typesetting:-mo(":", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("s", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&GreaterEqual;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"))}{\[\displaystyle F ^{\eslarr }\left(x \right)=\inf \left(s \mathop{\rm  }\in R :\mathop{\rm  }F \left(s \right)\ge x \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
by definition, it must be so that\mapleinline{inert}{2d}{x >= `#msup(mi("F"),mo("&larr;"))`(t)}{$\displaystyle \mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( t \right) \leq x$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
For necessity, assume\mapleinline{inert}{2d}{`#msup(mi("F"),mo("&larr;"))`(t) <= x}{$\displaystyle \mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( t \right) \leq x$}
. If\mapleinline{inert}{2d}{F(x) < t}{$\displaystyle F \left( x \right) <t$}
, then by the definition of\mapleinline{inert}{2d}{`#msup(mi("F"),mo("&larr;"))`(t)}{$\displaystyle \mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( t \right) $}
,\mapleinline{inert}{2d}{x < `#msup(mi("F"),mo("&larr;"))`(t)}{$\displaystyle x<\mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( t \right) $}
. This contradicts out assumption. Therefore\mapleinline{inert}{2d}{F(x) >= t}{$\displaystyle t\leq F \left( x \right) $}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{1.(b)}Let\mapleinline{inert}{2d}{F(x)}{$\displaystyle F \left( x \right) $}
consists of point mass\mapleinline{inert}{2d}{p}{$\displaystyle p$}
at\mapleinline{inert}{2d}{X = 0}{$\displaystyle X=0$}
and\mapleinline{inert}{2d}{1-p}{$\displaystyle 1-p$}
at\mapleinline{inert}{2d}{X = 1}{$\displaystyle X=1$}
. Then\mapleinline{inert}{2d}{F(x) = p*`I__0&le; x<1__`+`I__x&ge;1`}{$\displaystyle F \left( x \right) =p{\it I\_\_0}\le \mbox {{\tt ` x$<$1\_\_`}}\\
\mbox{}+{\it I\_\_x}\ge \mbox {{\tt `1`}}$}
. Thus\mapleinline{inert}{2d}{`in`(F(X), {0, 1, p})}{$\displaystyle {\it in} \left( F \left( X \right) , \left\{ 0,1,p \right\}  \right) $}
and is not uniformly\mapleinline{inert}{2d}{U(0, 1)}{$\displaystyle U \left( 0,1 \right) $}
distributed.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{1.(c) (i) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
Provided that the distribution function F is continuous,\mapleinline{inert}{2d}{P(X[s] = X[t]) = 0}{$\displaystyle P \left( X_{{s}}=X_{{t}} \right) =0$}
for\mapleinline{inert}{2d}{s <> t}{$\displaystyle s\neq t$}
. Then for some\mapleinline{inert}{2d}{k}{$\displaystyle k$}
,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{P(t = tau[k]) = P(X[t] = M[t]) and P(X[t] = M[t]) = 1/t}{$\displaystyle P \left( t=\tau_{{k}} \right) =P \left( X_{{t}}=M_{{t}} \right) {\rm \: \wedge \:}P \left( X_{{t}}=M_{{t}} \right) ={t}^{-1}$}
\mapleinline{inert}{2d}{}{$\displaystyle $}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Therefore the distribution of the record times are independent of F.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{1.(c).(ii)}Because\mapleinline{inert}{2d}{F(x)}{$\displaystyle F \left( x \right) $}
is continuous,\mapleinline{inert}{2d}{`#msup(mi("F"),mo("&larr;"))`(x)}{$\displaystyle \mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( x \right) $}
is strictly increasing. Thus one can use\mapleinline{inert}{2d}{`#msup(mi("F"),mo("&larr;"))`(x)}{$\displaystyle \mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( x \right) $}
as a transformation to convert the record sequences.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{1.(d)}Because F(x) is continuous on\mapleinline{inert}{2d}{RealRange(1, Open(infinity))}{$\displaystyle {\it RealRange} \left( 1,{\it Open} \left( \infty  \right)  \right) $}
, X can be simulated by a quantile transform from a uniformly U(0,1) distributed variable U: Let\mapleinline{inert}{2d}{U = F(X)}{$\displaystyle U=F \left( X \right) $}
. Then}\end{Maple Normal}

\begin{Maple Normal}{
\mapleinline{inert}{2d}{X = (1-U)^(-1/alpha)}{$\displaystyle X= \left( 1-U \right) ^{-{\alpha}^{-1}}$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{1.(d)}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{P(X[n+1] > max(X[1], () .. (), X[n]))}{\[\displaystyle P \left( \max \left( X_{{1}},{\ldots },X_{{n}} \right) <X_{{n+1}} \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("X", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("max", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("X", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mo("..", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.0em"), Typesetting:-mo(".", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-msub(Typesetting:-mi("X", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), Typesetting:-mi("")), mathvariant = "normal")), mathvariant = "normal"))}{\[\displaystyle =P \left(X _{n +1}=\max \left(X _{1},...,X _{n +1}\right)\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{X[i]}{$\displaystyle X_{{i}}$}
are iid,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{P(X[n+1] = max(X[1], () .. (), X[n+1]))}{\[\displaystyle P \left( X_{{n+1}}=\max \left( X_{{1}},{\ldots },X_{{n+1}} \right)  \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"))}{$\displaystyle =\mathop{\rm  }\frac{1}{n +1}$}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{2.(a) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
Consider}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(L(a*x)/L(x), x = infinity)}{\[\displaystyle \lim _{x\rightarrow \infty }{\frac {L \left( ax \right) }{L \left( x \right) }}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munder(Typesetting:-mi("lim", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accentunder = "false"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("`c__0`"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("`c__0`"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&epsilon;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle =\mathop{\rm  }\Mapleunderset{x \essrarr \infty }{\mathop{\rm lim}}\frac{{\it `c__0` }\left(a \mathop{\rm  }x \right)}{{\it `c__0` }\left(x \right)}\exp \left(\Mapleoverset{a \mathop{\rm  }x }{\Mapleunderset{x }{\int }}\frac{\epsilon \left(t \right)}{t }\mathop{\rm d}t \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&epsilon;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle \frac{\epsilon \left(x \right)}{x }\essrarr 0$}
as\mapleinline{inert}{2d}{proc (x) options operator, arrow; infinity end proc}{$\displaystyle x\mapsto \infty $}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&epsilon;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&approx;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&epsilon;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), mathvariant = "normal"), Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), mathvariant = "normal"), Typesetting:-mi("&epsilon;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{\[\displaystyle \Mapleoverset{a \mathop{\rm  }x }{\Mapleunderset{x }{\int }}\frac{\epsilon \left(t \right)}{t }\mathop{\rm d}t \mathop{\rm  }\esapprox \frac{\epsilon \left(x \right)}{x }\left(a -1\right)x =\left(a -1\right)\epsilon \left(x \right)\essrarr 0\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Therefore}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(L(a*x)/L(x), x = infinity)}{$\displaystyle \lim _{x\rightarrow \infty }{\frac {L \left( ax \right) }{L \left( x \right) }}$}
=\mapleinline{inert}{2d}{limit(c__0(a*x)*exp*{int(`&epsilon;`(t)/t, t = x .. a*x)}/c__0(x), x = infinity) = 1}{$\displaystyle \lim _{x\rightarrow \infty } \left( {\it c\_\_0} \left( ax \right)  \left( \exp \right) \, \left\{ \int _{x}^{ax}\!{\frac {\epsilon  \left( t \right) }{t}}{dt} \right\}  \left( {\it c\_\_0} \left( x \right) \\
\mbox{} \right) ^{-1} \right) =1$}
}\end{Maple Normal}

}\end{Maple Normal}

Thus, by definition,\mapleinline{inert}{2d}{L(x)}{$\displaystyle L \left( x \right) $}
is a slowing varying function.

\textbf{}\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{2.(b) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
For\mapleinline{inert}{2d}{x > a}{$\displaystyle a<x$}
where\mapleinline{inert}{2d}{(log*log)*log*a > 0}{$\displaystyle 0<{{\it log}}^{3}a$}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{[(log*log)*log(x)]^2 = exp[((2*log*log)*log*log)*x]}{$\displaystyle {[{{\it log}}^{2}{\it log} \left( x \right) ]}^{2}=\exp_{{2\,{{\it log}}^{4}x}}$}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{diff([((log*log)*log*log)*x], x) = 1/(x*((ln(10))((log*log)*log*x))((log*log)*x)*log*x)}{\[\displaystyle {\frac {\partial }{\partial x}} \left( [{{\it log}}^{4}x] \right) ={\frac {1}{{x}^{2}\ln  \left( 10 \right)  \left( {{\it log}}^{3}x \right)  \left( {{\it log}}^{2}x \right)  \left( {\it log} \right) \\
\mbox{}}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Therefore we can write}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{L(x) = c*exp[int(1/(x*((ln(10))((log*log)*log*x))((log*log)*x)*log*x), x = 10^4 .. x)]}{$\displaystyle L \left( x \right) =c\exp_{{\int _{10000}^{x}\!{\frac {1}{{x}^{2}\ln  \left( 10 \right)  \left( {{\it log}}^{3}x \right)  \left( {{\it log}}^{2}x \right) \\
\mbox{} \left( {\it log} \right) }}{dx}}}$}
for\mapleinline{inert}{2d}{x >= 10^4}{$\displaystyle 10000\leq x$}
and some constant c. As\mapleinline{inert}{2d}{proc (x) options operator, arrow; infinity end proc}{$\displaystyle x\mapsto \infty $}
,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("10", mathvariant = "normal")), mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{\[\displaystyle \frac{1}{\mathop{\rm ln}\left(10\right)\left(\mathop{\rm log}\mathop{\rm  }\mathop{\rm log}\mathop{\rm  }\mathop{\rm log}\mathop{\rm  }x \right)\left(\mathop{\rm log}\mathop{\rm  }\mathop{\rm log}\mathop{\rm  }x \right)\mathop{\rm  }\mathop{\rm log}\mathop{\rm  }x }\essrarr 0\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Thus\mapleinline{inert}{2d}{L(x)}{$\displaystyle L \left( x \right) $}
is a slowing varying function.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
For\mapleinline{inert}{2d}{x >= a}{$\displaystyle a\leq x$}
, where\mapleinline{inert}{2d}{a > 1}{$\displaystyle 1<a$}
is a constant,\mapleinline{inert}{2d}{L__0(x)}{$\displaystyle {\it L\_\_0} \left( x \right) $}
can be re-written as}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{L__0(x) = exp(sqrt(log*a))*exp*{int(1/(2*t*sqrt(log*t)), t = a .. x)}}{\[\displaystyle {\it L\_\_0} \left( x \right) ={{\rm e}^{ \sqrt{ \left( {\it log} \right) \,a}}} \left( \exp \right) \, \left\{ \int _{a}^{x}\!1/2\,{\frac {1}{t \sqrt{ \left( {\it log} \right) \,t}}}{dt} \right\} \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Clearly}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(1/(2*sqrt(x)), x = infinity) = 0}{\[\displaystyle \lim _{x\rightarrow \infty }1/2\, \left(  \sqrt{x} \right) ^{-1}=0\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Therefore\mapleinline{inert}{2d}{L__0(x)}{$\displaystyle {\it L\_\_0} \left( x \right) $}
is a slowing varying function.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{2(c). Solution:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{log(x) = exp(ln*log*x)}{\[\displaystyle {\it log} \left( x \right) ={{\rm e}^{ \left( {\it ln} \right) \, \left( {\it log} \right) \,x}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("`x__0`")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("10", mathvariant = "normal")), mathvariant = "normal"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0")), mathvariant = "normal")), mathvariant = "normal"))}{\[\displaystyle =\mathop{\rm  }\exp \left(\Mapleoverset{x }{\Mapleunderset{{\it `x__0` }}{\int }}\frac{1}{\mathop{\rm ln}\left(10\right)t \mathop{\rm  }\mathop{\rm  }\mathop{\rm log}\left(t \right)}\mathop{\rm d}t \mathop{\rm  }+\mathop{\rm ln}\mathop{\rm  }\mathop{\rm log}\left(x _{0}\right)\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("`x__0`")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("10", mathvariant = "normal")), mathvariant = "normal"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), mathvariant = "normal")), Typesetting:-mi(""))}{\[\displaystyle =\mathop{\rm  }\mathop{\rm log}\left(x _{0}\right)\mathop{\rm  }\exp \left(\Mapleoverset{x }{\Mapleunderset{{\it `x__0` }}{\int }}\frac{1}{\mathop{\rm ln}\left(10\right)t \mathop{\rm  }\mathop{\rm  }\mathop{\rm log}\left(t \right)}\mathop{\rm d}t \mathop{\rm  }\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
where\mapleinline{inert}{2d}{1 < x[0] and x[0] <= x}{$\displaystyle 1<x_{{0}}{\rm \: \wedge \:}x_{{0}}\leq x$}
. We identify\mapleinline{inert}{2d}{`&epsilon;`(x) = 1/[ln(10)*log(x)]}{$\displaystyle \epsilon  \left( x \right) ={[\ln  \left( 10 \right) {\it log} \left( x \right) ]}^{-1}$}
. Clearly\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("&epsilon;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle \epsilon \left(x \right)\mathop{\rm  }\essrarr 0$}
as\mapleinline{inert}{2d}{proc (x) options operator, arrow; infinity end proc}{$\displaystyle x\mapsto \infty $}
.\mapleinline{inert}{2d}{}{$\displaystyle $}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{2.(d) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("E", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("X", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0")), mathvariant = "normal"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), mathvariant = "normal", open = "[", close = "]"))}{\[\displaystyle E \left(X ^{\delta }\right)=-\Mapleoverset{\infty }{\Mapleunderset{0}{\int }}x ^{\delta }\mathop{\rm d}\left(\Mapleoverset{{\it \_} }{F }\left(x \right)\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munder(Typesetting:-mi("lim", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accentunder = "false"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic"))}{\[\displaystyle =\mathop{\rm  }\Mapleunderset{x \essrarr \infty }{\mathop{\rm lim}}x ^{\delta }\Mapleoverset{{\it \_} }{F }\left(x \right)+\delta \mathop{\rm  }\Mapleoverset{\infty }{\Mapleunderset{0}{\int }}x ^{\delta -\alpha -1}L \left(x \right)\mathop{\rm d}x \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"))}{$\displaystyle \Mapleoverset{{\it \_} }{F }\left(x \right)$}
is regularly varying with index\mapleinline{inert}{2d}{-alpha}{$\displaystyle -\alpha$}
, it has representation}\end{Maple Normal}

}\end{Maple Normal}

\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("c", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "[", close = "]"))}{\[\displaystyle \Mapleoverset{{\it \_} }{F }\left(x \right)=c \left(x \right)\mathop{\rm  }\exp \left(\Mapleoverset{x }{\Mapleunderset{x _{0}}{\int }}\frac{\xi \left(t \right)}{t }\mathop{\rm d}t \right)\]}
for some\mapleinline{inert}{2d}{x[0] > 0}{$\displaystyle 0<x_{{0}}$}
,\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("c", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msub(Typesetting:-mi("c", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"))}{$\displaystyle c \left(x \right)\essrarr c _{0}$}
,\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"))}{$\displaystyle \xi \left(x \right)\essrarr -\alpha $}
as\mapleinline{inert}{2d}{proc (x) options operator, arrow; infinity end proc}{$\displaystyle x\mapsto \infty $}
. Therefore

\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("c", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0")), mathvariant = "normal")), mathvariant = "normal", open = "[", close = "]"))}{\[\displaystyle x ^{\delta }\mathop{\rm  }\Mapleoverset{{\it \_} }{F }\left(x \right)=c \left(x \right)\mathop{\rm  }\exp \left(\Mapleoverset{x }{\Mapleunderset{x _{0}}{\int }}\frac{\xi \left(t \right)+\delta }{t }\mathop{\rm d}t +\delta \mathop{\rm  }\mathop{\rm ln}\left(x _{0}\right)\right)\]}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("c", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0")), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&int;", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("1", executable = "true", font_style_name = "2D Input", mathvariant = "normal")), subscriptshift = "0")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mo("+", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic"), Typesetting:-mo(" ", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("+", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mo(" ", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("ln", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0")), mathvariant = "normal")), mathvariant = "normal", open = "[", close = "]")), Typesetting:-mi(""))}{\[\displaystyle =\mathop{\rm  }c \left(x \right)\mathop{\rm  }\exp \left(\Mapleoverset{x _{1}}{\Mapleunderset{x _{0}}{\int }}\frac{\xi \left(t \right)+\delta }{t }\mathop{\rm d}t +\Mapleoverset{x }{\Mapleunderset{x _{1}}{\int }}\frac{\xi \left(t \right)+\delta }{t }\mathop{\rm d}t \mathop{\rm  }+\delta \mathop{\rm  }\mathop{\rm ln}\left(x _{0}\right)\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mo("+", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle \frac{\xi \left(t \right)+\delta }{t }\essrarr 0$}
as\mapleinline{inert}{2d}{proc (t) options operator, arrow; infinity end proc}{$\displaystyle t\mapsto \infty $}
, we can choose\mapleinline{inert}{2d}{x[1]}{$\displaystyle x_{{1}}$}
so large that}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{`&approx;`(int((xi(t)+delta)/t, t = x[1] .. x), (delta-alpha)*(x-x[1])/x[1])}{\[\displaystyle \mbox {{\tt `\&approx;`}} \left( \int _{x_{{1}}}^{x}\!{\frac {\xi \left( t \right) +\delta}{t}}{dt},{\frac { \left( \delta-\alpha \right) \\
\mbox{} \left( x-x_{{1}} \right) }{x_{{1}}}} \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{If}\mapleinline{inert}{2d}{delta < alpha}{$\displaystyle \delta<\alpha$}
,\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("1", executable = "true", font_style_name = "2D Input", mathvariant = "normal")), subscriptshift = "0")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mo("+", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic"), Typesetting:-mo(" ", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&srarr;", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&uminus0;", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), Typesetting:-mi(""))}{$\displaystyle \Mapleoverset{x }{\Mapleunderset{x _{1}}{\int }}\frac{\xi \left(t \right)+\delta }{t }\mathop{\rm d}t \mathop{\rm  }\essrarr -\infty $}
and\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle x ^{\delta }\mathop{\rm  }\Mapleoverset{{\it \_} }{F }\left(x \right)\essrarr 0$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Moreover, if\mapleinline{inert}{2d}{delta < alpha}{$\displaystyle \delta<\alpha$}
, i.e.\mapleinline{inert}{2d}{delta-alpha-1 < -1}{$\displaystyle \delta-\alpha<0$}
, by Karamata's theorem,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-msup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-msup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle \Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta -\alpha -1}L \left(x \right)\mathop{\rm  }\mathop{\rm d}x \estilde \frac{1}{\alpha -\delta }z ^{\delta -\alpha }L \left(z \right)\estilde \frac{1}{\alpha -\delta }z ^{\delta }\mathop{\rm  }\Mapleoverset{{\it \_} }{F }\left(z \right)\essrarr 0$}
as\mapleinline{inert}{2d}{proc (z) options operator, arrow; infinity end proc}{$\displaystyle z\mapsto \infty $}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
Therefore\mapleinline{inert}{2d}{int(x^(delta-alpha-1)*L(x), x = 0 .. infinity) = 0 and 0 < infinity}{$\displaystyle \int _{0}^{\infty }\!{x}^{\delta-\alpha-1}L \left( x \right) {dx}=0{\rm \: \wedge \:}0<\infty $}
. So\mapleinline{inert}{2d}{E*X^delta < infinity}{$\displaystyle E{X}^{\delta}<\infty $}
when\mapleinline{inert}{2d}{delta < alpha}{$\displaystyle \delta<\alpha$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{if}\mapleinline{inert}{2d}{delta > alpha}{$\displaystyle \alpha<\delta$}
,\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("1", executable = "true", font_style_name = "2D Input", mathvariant = "normal")), subscriptshift = "0")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mo("+", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic"), Typesetting:-mo(" ", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&srarr;", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("&infin;", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic"))}{$\displaystyle \Mapleoverset{x }{\Mapleunderset{x _{1}}{\int }}\frac{\xi \left(t \right)+\delta }{t }\mathop{\rm d}t \mathop{\rm  }\essrarr \infty $}
and\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&infin;", italic = "true", mathvariant = "italic", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo(".", italic = "true", mathvariant = "italic", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"))}{\[\displaystyle x ^{\delta }\mathop{\rm  }\Mapleoverset{{\it \_} }{F }\left(x \right)\essrarr \infty .\]}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
and}\end{Maple Normal}

}\end{Maple Normal}

\textbf{}\mapleinline{inert}{2d}{int(x^(delta-alpha-1)*L(x), x = 0 .. infinity) = limit(int(t^(delta-alpha-1)*L(t), t = 0 .. x), x = infinity)}{\[\displaystyle \int _{0}^{\infty }\!{x}^{\delta-\alpha-1}L \left( x \right) {dx}=\lim _{x\rightarrow \infty } \left( \int _{0}^{x}\!{t}^{\delta-\alpha-1}\\
\mbox{}L \left( t \right) {dt} \right) \]}
\begin{Maple Normal}{
\begin{Maple Normal}{
By Karamata's theorem, as\mapleinline{inert}{2d}{proc (x) options operator, arrow; infinity end proc}{$\displaystyle x\mapsto \infty $}
,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic"))}{$\displaystyle \Mapleoverset{x }{\Mapleunderset{0}{\int }}t ^{\delta -\alpha -1}L \left(t \right)\mathop{\rm d}t \estilde \frac{1}{\delta -\alpha }x ^{\delta -\alpha }L \left(x \right)\essrarr \infty $}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
Thus\mapleinline{inert}{2d}{E*X^delta = infinity}{$\displaystyle E{X}^{\delta}=\infty $}
when\mapleinline{inert}{2d}{delta > alpha}{$\displaystyle \alpha<\delta$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{For an example of}\mapleinline{inert}{2d}{E*X^alpha = infinity}{$\displaystyle E{X}^{\alpha}=\infty $}
, consider\mapleinline{inert}{2d}{F(x) = 1-1/x}{$\displaystyle F \left( x \right) =1-{x}^{-1}$}
for\mapleinline{inert}{2d}{`in`(x, RealRange(1, Open(infinity)))}{$\displaystyle {\it in} \left( x,{\it RealRange} \left( 1,{\it Open} \left( \infty \\
\mbox{} \right)  \right)  \right) $}
, where\mapleinline{inert}{2d}{x[0] > 0}{$\displaystyle 0<x_{{0}}$}
.}\end{Maple Normal}

}\end{Maple Normal}

\mapleinline{inert}{2d}{E*X = int(1/x, x = 1 .. infinity) and int(1/x, x = 1 .. infinity) = infinity}{\[\displaystyle EX=\int _{1}^{\infty }\!{x}^{-1}{dx}{\rm \: \wedge \:}\int _{1}^{\infty }\!{x}^{-1}{dx}=\infty \]}
\textbf{For an example of}\mapleinline{inert}{2d}{E*X^alpha < infinity}{$\displaystyle E{X}^{\alpha}<\infty $}
, consider\mapleinline{inert}{2d}{F(x) = 1-(a*ln^2*a)/(x*ln^2*x)}{$\displaystyle F \left( x \right) =1-{\frac {{a}^{2}}{{x}^{2}}}$}
for\mapleinline{inert}{2d}{`in`(x, RealRange(a, Open(infinity)))}{$\displaystyle {\it in} \left( x,{\it RealRange} \left( a,{\it Open} \left( \infty \\
\mbox{} \right)  \right)  \right) $}
,\mapleinline{inert}{2d}{a > 0}{$\displaystyle 0<a$}
, which is regularly varying with index -1:

\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(t*x)/`#mover(mi("F"),mo("&uminus0;"))`(x) = [x*ln^2*x]/[t*x*(ln*t+ln*x)^2] and [x*ln^2*x]/[t*x*(ln*t+ln*x)^2] = 1/t}{\[\displaystyle {\frac {\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( tx \right) }{\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( x \right) }}={\frac {[{x}^{2}{{\it ln}}^{2}]}{[tx \left(  \left( {\it ln} \right) \,t+ \left( {\it ln} \right) \,x \right) ^{2}]}}{\rm \: \wedge \:}{\frac {[{x}^{2}{{\it ln}}^{2}]}{[tx \left(  \left( {\it ln} \right) \,t+ \left( {\it ln} \right) \,x \right) ^{2}]}}={t}^{-1\\
\mbox{}}\]}
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("E", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("X", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("a", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dx", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("2", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dx", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi(""))}{\[\displaystyle E \mathop{\rm  }X =\Mapleoverset{\infty }{\Mapleunderset{a }{\int }}\frac{a \mathop{\rm  }\mathop{\rm ln}^{2}a }{x \mathop{\rm  }\mathop{\rm ln}^{2}x \mathop{\rm  }}{\it dx} =a \mathop{\rm  }\mathop{\rm ln}^{2}a \mathop{\rm  }\left(-\frac{1}{\mathop{\rm ln}\mathop{\rm  }a }+2\mathop{\rm  }\Mapleoverset{\infty }{\Mapleunderset{a }{\int }}\frac{1}{x \mathop{\rm  }\mathop{\rm ln}^{2}x }{\it dx} \right)=a \mathop{\rm  }\mathop{\rm ln}\left(a \right)<\infty \]}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{2.(e) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(`#mover(mi("\`c__n\`"),mo("&tilde;"))`) = `#mover(mi("F"),mo("&uminus0;"))`(`#mover(mi("\`c__n\`"),mo("&tilde;"))`/c__n*c__n)}{\[\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( \mbox {{\tt `\#mover(mi("`c\_\_n`"),mo("`}}\mbox {{\tt `\~{}`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \right) =\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( \mbox {{\tt `\#mover(mi("`c\_\_n`"),mo("`}}\mbox {{\tt `\~{}`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`}{$\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}$}
is regularly varying, by theorem A3.2,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(`#mover(mi("\`c__n\`"),mo("&tilde;"))`/c__n*c__n) = (`#mover(mi("\`c__n\`"),mo("&tilde;"))`/c__n)^(-alpha)*`#mover(mi("F"),mo("&uminus0;"))`(c__n)}{\[\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( \mbox {{\tt `\#mover(mi("`c\_\_n`"),mo("`}}\mbox {{\tt `\~{}`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \right) = \left( {\frac {\mbox {{\tt `\#mover(mi("`c\_\_n`"),mo("`}}\mbox {{\tt `\~{}`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{}}{{\it c\_\_n}}} \right) ^{-\alpha}\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( {\it c\_\_n} \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
But\mapleinline{inert}{2d}{`~`[`#mover(mi("F"),mo("&uminus0;"))`(c__n)]/n}{$\displaystyle {\frac {\mbox {{\tt `\~{}`}}_{{\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( {\it c\_\_n} \right) }}}{n}}$}
and\mapleinline{inert}{2d}{`~`[`#mover(mi("F"),mo("&uminus0;"))`(`#mover(mi("\`c__n\`"),mo("&tilde;"))`)]/n}{$\displaystyle {\frac {\mbox {{\tt `\~{}`}}_{{\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( \mbox {{\tt `\#mover(mi("`c\_\_n`"),mo("`}}\mbox {{\tt `\~{}`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \right) }}}{n}}$}
, therefore\mapleinline{inert}{2d}{`~`[c__n]*`#mover(mi("\`c__n\`"),mo("&tilde;"))`}{$\displaystyle \mbox {{\tt `\~{}`}}_{{{\it c\_\_n}}}\mbox {{\tt `\#mover(mi("`c\_\_n`"),mo("`}}\mbox {{\tt `\~{}`}}\\
\mbox{}\mbox {{\tt `"))`}}$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{2.(f) (i) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(x__F-1/x) = x^(-alpha)*L(x)}{$\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( {\it x\_\_F}-{x}^{-1} \right) ={x}^{-\alpha}L \left( x \right) $}
implies}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{P(X > x__F-1/x) = x^(-alpha)*L(x)}{\[\displaystyle P \left( {\it x\_\_F}-{x}^{-1}<X \right) ={x}^{-\alpha}L \left( x \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\mapleinline{inert}{2d}{P(1/x > x__F-X) = x^(-alpha)*L(x)}{\[\displaystyle P \left( {\it x\_\_F}-X<{x}^{-1} \right) ={x}^{-\alpha}L \left( x \right) \]}
\mapleinline{inert}{2d}{P(x < 1/(x__F-X)) = x^(-alpha)*L(x)}{\[\displaystyle P \left( x< \left( {\it x\_\_F}-X \right) ^{-1} \right) ={x}^{-\alpha}L \left( x \right) \]}
\mapleinline{inert}{2d}{`#mover(mi("G"),mo("&uminus0;"))`(x) = x^(-alpha)*L(x)}{$\displaystyle \mbox {{\tt `\#mover(mi("G"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( x \right) ={x}^{-\alpha}L \left( x \right) $}
\mapleinline{inert}{2d}{}{$\displaystyle $}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{2.(f) (ii)}\textbf{Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{1/(x__F-X)}{$\displaystyle  \left( {\it x\_\_F}-X \right) ^{-1}$}
has distribution function\mapleinline{inert}{2d}{G}{$\displaystyle G$}
with regularly varying tail as proven in (i),\mapleinline{inert}{2d}{`in`(1/(x__F-X), MDA(`&Phi;__&alpha;`))}{$\displaystyle {\it in} \left(  \left( {\it x\_\_F}-X \right) ^{-1},{\it MDA} \left( \Phi {\it \_\_}\\
\mbox{}\alpha  \right)  \right) $}
, which implies}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("`x__F`"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("`M__n`")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-msup(Typesetting:-mi("G", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&larr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em")), superscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal"), Typesetting:-mi("y", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("e", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msup(Typesetting:-mi("y", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0")), superscriptshift = "0"))}{\[\displaystyle P \left(\frac{1}{{\it `x__F` }-{\it `M__n` }}<G ^{\eslarr }\left(1-\frac{1}{n }\right)y \right)\essrarr e ^{-y ^{-\alpha }}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{`#msup(mi("G"),mo("&larr;"))`(1-1/n) = 1/(x__F-`#msup(mi("F"),mo("&larr;"))`(1-1/n))}{$\displaystyle \mbox {{\tt `\#msup(mi("G"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( 1-{n}^{-1} \right) = \left( {\it x\_\_F}-\mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( 1-{n}^{-1} \right)  \right) ^{-1}$}
, we write}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("`x__F`"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msup(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&larr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em")), superscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal")), Typesetting:-mi("")), Typesetting:-mrow(Typesetting:-mi("`x__F`"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("`M__n`")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("y", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("e", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msup(Typesetting:-mi("y", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0")), superscriptshift = "0")), Typesetting:-mi(""))}{\[\displaystyle P \left(\frac{{\it `x__F` }-F ^{\eslarr }\left(1-\frac{1}{n }\right)}{{\it `x__F` }-{\it `M__n` }}<y \right)\essrarr e ^{-y ^{-\alpha }}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("`M__n`"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("`x__F`")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("`x__F`"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msup(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&larr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em")), superscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal")), Typesetting:-mi("")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("y", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("e", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msup(Typesetting:-mi("y", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0")), superscriptshift = "0"))}{$\displaystyle P \left(\frac{{\it `M__n` }-{\it `x__F` }}{{\it `x__F` }-F ^{\eslarr }\left(1-\frac{1}{n }\right)}<-\frac{1}{y }\right)\essrarr e ^{-y ^{-\alpha }}$}
\mapleinline{inert}{2d}{}{$\displaystyle $}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Let\mapleinline{inert}{2d}{-1/y = x}{$\displaystyle -{y}^{-1}=x$}
. Then}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("`M__n`"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("`x__F`")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("`x__F`"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msup(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&larr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em")), superscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal")), Typesetting:-mi("")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("e", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msup(Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0")), superscriptshift = "0"))}{\[\displaystyle P \left(\frac{{\it `M__n` }-{\it `x__F` }}{{\it `x__F` }-F ^{\eslarr }\left(1-\frac{1}{n }\right)}<x \right)\essrarr e ^{-\left(-x \right)^{\alpha }}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
This means, with constant\mapleinline{inert}{2d}{c__n = x__F-`#msup(mi("F"),mo("&larr;"))`(1-1/n)}{$\displaystyle {\it c\_\_n}={\it x\_\_F}-\mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( 1-{n}^{-1} \right) $}
,\mapleinline{inert}{2d}{`#mover(mo("&rarr;"),mi("d"))`*`~`[Y]*`&Psi;__&alpha;`/c__n(M__n-x__F)}{$\displaystyle {\frac {\mbox {{\tt `\#mover(mo("`}}\rightarrow \mbox {{\tt `"),mi("d"))`}}\\
\mbox{}\\
\mbox{}\,\mbox {{\tt `\~{}`}}_{{Y}}\Psi {\it \_\_}\alpha \\
\mbox{}}{{\it c\_\_n} \left( {\it M\_\_n}-{\it x\_\_F} \right) }}$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{3. (a)}We show the statement is true for von. Vise's function. Since all function in\mapleinline{inert}{2d}{MDA(Lambda)}{$\displaystyle {\it MDA} \left( \Lambda \right) $}
have tails equivalent to that of the von Mise's function, that the statement is true for von Mise's function implies it is also true for the rest of the functions in}\end{Maple Normal}

\begin{Maple Normal}{
have tails equivalent to that of the von Mise's function, that the statement is true for von Mise's function implies it is also true for the rest of the functions in\mapleinline{inert}{2d}{MDA(Lambda)}{$\displaystyle {\it MDA} \left( \Lambda \right) $}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-munder(Typesetting:-mi("lim", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accentunder = "false"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-munder(Typesetting:-mi("lim", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&infin;", italic = "true", mathvariant = "italic", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("y", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dy", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "[", close = "]"))}{\[\displaystyle \Mapleunderset{x \essrarr \infty }{\mathop{\rm lim}}\frac{\Mapleoverset{-}{F }\left(t \mathop{\rm  }x \right)}{\Mapleoverset{-}{F }\left(x \right)}=\Mapleunderset{x \essrarr \infty }{\mathop{\rm lim}}\mathop{\rm  }\exp \left(-\Mapleoverset{t \mathop{\rm  }x }{\Mapleunderset{x }{\int }}\frac{1}{a \left(y \right)}{\it dy} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munder(Typesetting:-mi("lim", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&infin;", italic = "true", mathvariant = "italic", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal", open = "[", close = "]"))}{\[\displaystyle =\mathop{\rm  }\Mapleunderset{x \essrarr \infty }{\mathop{\rm lim}}\mathop{\rm  }\exp \left(-\frac{x \left(t -1\right)}{a \left(x \right)}\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
if\textbf{}\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi(""))}{$\displaystyle a \left(x \right)\essrarr \infty $}
,\mapleinline{inert}{2d}{limit(x(t-1)/a(x), x = infinity) = limit((t-1)/(diff(a(x), x)), x = infinity)}{$\displaystyle \lim _{x\rightarrow \infty }{\frac {x \left( t-1 \right) }{a \left( x \right) }}=\lim _{x\rightarrow \infty }{\frac {t-1}{{\frac {d}{dx}}a \left( x \right) }}$}
. Since\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo("'", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle a \esapos \left(x \right)\essrarr 0$}
,\mapleinline{inert}{2d}{limit((t-1)/(diff(a(x), x)), x = infinity) = piecewise(0 < t and t < 1, -infinity, t > 1, infinity)}{$\displaystyle \lim _{x\rightarrow \infty }{\frac {t-1}{{\frac {d}{dx}}a \left( x \right) }}=\cases{-\infty &$0<t$\  and \ $t<1$\cr \infty &$1<t$\cr}$}
.and\mapleinline{inert}{2d}{limit(exp[-x(t-1)/a(x)], x = infinity) = piecewise(0 < t and t < 1, infinity, t > 1, 0)}{$\displaystyle \lim _{x\rightarrow \infty }\exp_{{-{\frac {x \left( t-1 \right) }{a \left( x \right) }}}}=\cases{\infty &$0<t$\  and \ $t<1$\cr 0&$1<t$\cr}$}

\begin{Maple Normal}{
\begin{Maple Normal}{
if\mapleinline{inert}{2d}{not (a(x) or 0)}{$\displaystyle {\rm \neg \:} a \left( x \right) {\rm \: \vee \:}0$}
or\mapleinline{inert}{2d}{not (a(x) or C)}{$\displaystyle {\rm \neg \:} a \left( x \right) {\rm \: \vee \:}C$}
, where\mapleinline{inert}{2d}{C}{$\displaystyle C$}
is a positive constant, it is clear\mapleinline{inert}{2d}{limit(x(t-1)/a(x), x = infinity) = piecewise(0 < t and t < 1, -infinity, t > 1, infinity)}{$\displaystyle \lim _{x\rightarrow \infty }{\frac {x \left( t-1 \right) }{a \left( x \right) }}=\cases{-\infty &$0<t$\  and \ $t<1$\cr \infty &$1<t$\cr}$}
.and\mapleinline{inert}{2d}{limit(exp[-x(t-1)/a(x)], x = infinity) = piecewise(0 < t and t < 1, infinity, t > 1, 0)}{$\displaystyle \lim _{x\rightarrow \infty }\exp_{{-{\frac {x \left( t-1 \right) }{a \left( x \right) }}}}=\cases{\infty &$0<t$\  and \ $t<1$\cr 0&$1<t$\cr}$}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{3. (b) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("G", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "[", close = "]")), Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"))}{\[\displaystyle \frac{\Mapleoverset{-}{F }\left(x \right)}{\Mapleoverset{-}{G }\left(x \right)}\estilde \frac{\exp \left(-\Mapleoverset{x }{\Mapleunderset{z }{\int }}\frac{1}{a \left(t \right)}{\it dt} \right)}{x ^{-\alpha }}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), mathvariant = "normal", open = "[", close = "]"))}{\[\displaystyle \estilde \mathop{\rm  }\exp \left(-\Mapleoverset{x }{\Mapleunderset{z }{\int }}\frac{1}{a \left(t \right)}{\it dt} +\alpha \mathop{\rm  }\mathop{\rm ln}\left(x \right)\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle \estilde \exp \left(\Mapleoverset{x }{\Mapleunderset{\xi }{\int }}\left(\frac{\alpha }{t }-\frac{1}{a \left(t \right)}\right){\it dt} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
where\mapleinline{inert}{2d}{xi}{$\displaystyle \xi$}
is a very large number of our choice.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle \exp \left(\Mapleoverset{x }{\Mapleunderset{\xi }{\int }}\left(\frac{\alpha }{t }-\frac{1}{a \left(t \right)}\right){\it dt} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("")), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle \estilde \exp \left(\Mapleoverset{x }{\Mapleunderset{\xi }{\int }}\left(\frac{\alpha \mathop{\rm  }a \left(t \right)-t }{t \mathop{\rm  }a \left(t \right)}\right){\it dt} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Here we note that the integrand tends to a constant\mapleinline{inert}{2d}{-1/a__0}{$\displaystyle -{{\it a\_\_0}}^{-1}$}
as\mapleinline{inert}{2d}{proc (xi) options operator, arrow; infinity end proc}{$\displaystyle \xi\mapsto \infty $}
. Therefore}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("")), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle \exp \left(\Mapleoverset{x }{\Mapleunderset{\xi }{\int }}\left(\frac{\alpha \mathop{\rm  }a \left(t \right)-t }{t \mathop{\rm  }a \left(t \right)}\right){\it dt} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{`~exp`*{((x-xi)/(xi*a__0))[a__0*alpha-xi]}}{\[\displaystyle \mbox {{\tt `\~{}exp`}}\, \left\{ {\frac {x-\xi}{\xi\,{\it a\_\_0}}}_{{\alpha\,{\it a\_\_0}-\xi}} \right\} \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{xi}{$\displaystyle \xi$}
can be chosen arbitrarily large, the above expression can be arbitrarily close to 0. Hence}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(`#mover(mi("F"),mo("&uminus0;"))`(x)/`#mover(mi("G"),mo("&uminus0;"))`(x), x = infinity) = 0}{\[\displaystyle \lim _{x\rightarrow \infty }{\frac {\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( x \right) }{\mbox {{\tt `\#mover(mi("G"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( x \right) }}=0\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{Now consider}\mapleinline{inert}{2d}{E*`X__+`^delta}{$\displaystyle E{\mbox {{\tt `X\_\_+`}}}^{\delta}$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("E", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("`X__+`")), mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("`x__0`"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("G", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("")), accent = "true"), Typesetting:-mi(""))}{\[\displaystyle E \mathop{\rm  }\left({\it `X__+` }\right)^{\delta }=-\Mapleoverset{\infty }{\Mapleunderset{{\it `x__0` }\mathop{\rm  }}{\int }}x ^{\delta }\Mapleoverset{}{d \mathop{\rm  }\Mapleoverset{-}{G }\left(x \right)}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
where\mapleinline{inert}{2d}{x__0 = inf*supp(`X__+`)}{$\displaystyle {\it x\_\_0}=\inf\,{\it supp} \left( \mbox {{\tt `X\_\_+`}} \right) $}
. Then}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("E", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("`X__+`")), mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("`x__0`")), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("G", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("G", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mi(""))}{\[\displaystyle E \mathop{\rm  }\left({\it `X__+` }\right)^{\delta }=\mathop{\rm  }-\Mapleoverset{z }{\Mapleunderset{{\it `x__0` }}{\int }}x ^{\delta }d \mathop{\rm  }\Mapleoverset{-}{G }\left(x \right)-\Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta }d \mathop{\rm  }\Mapleoverset{-}{G }\left(x \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
for some large constant z. The 1st integral is clearly finite. As for the 2nd integral,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("G", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("F", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"))}{\[\displaystyle -\Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta }d \mathop{\rm  }\Mapleoverset{-}{G }\left(x \right)<-\Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta }d \mathop{\rm  }\Mapleoverset{-}{F }\left(x \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
where\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(x) = x^(-alpha)*L(x)}{$\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( x \right) ={x}^{-\alpha}L \left( x \right) $}
is any regularly varying function with\mapleinline{inert}{2d}{alpha > 0}{$\displaystyle 0<\alpha$}
. Then}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("F", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munder(Typesetting:-mi("lim", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("C", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("dx", italic = "true", mathvariant = "italic"))}{\[\displaystyle -\Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta }d \mathop{\rm  }\Mapleoverset{-}{F }\left(x \right)=-\Mapleunderset{x \essrarr \infty }{\mathop{\rm lim}}\mathop{\rm  }x ^{\delta -\alpha }L \left(x \right)+C \mathop{\rm  }+\mathop{\rm  }\delta \mathop{\rm  }\Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta -\alpha -1}L \left(x \right)\mathop{\rm  }{\it dx} \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
where C is a constant. For\mapleinline{inert}{2d}{alpha > delta}{$\displaystyle \delta<\alpha$}
,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(x^(delta-alpha)*L(x), x = infinity) = 0}{\[\displaystyle \lim _{x\rightarrow \infty }{x}^{\delta-\alpha}L \left( x \right) =0\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
and by Karamata's theorem}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("dx", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-msup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle \Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta -\alpha -1}L \left(x \right)\mathop{\rm  }{\it dx} \mathop{\rm  }\estilde \frac{1}{\alpha -\delta }z ^{\delta -\alpha }L \left(z \right)\essrarr 0$}
as\mapleinline{inert}{2d}{proc (z) options operator, arrow; infinity end proc}{$\displaystyle z\mapsto \infty $}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Therefore\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("G", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("F", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi(""))}{$\displaystyle -\Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta }d \mathop{\rm  }\Mapleoverset{-}{G }\left(x \right)<-\Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta }d \mathop{\rm  }\Mapleoverset{-}{F }\left(x \right)<\infty $}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{3. (c) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
For n large,\mapleinline{inert}{2d}{1/n = `#mover(mi("F"),mo("&uminus0;"))`(d__n)}{$\displaystyle {n}^{-1}=\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( {\it d\_\_n} \right) $}
implies}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("c", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("`d__n`")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "[", close = "]"))}{\[\displaystyle \frac{1}{n }=c \mathop{\rm  }\exp \left(-\Mapleoverset{{\it `d__n` }}{\Mapleunderset{z }{\int }}\frac{1}{a \left(t \right)}{\it dt} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
where\mapleinline{inert}{2d}{z}{$\displaystyle z$}
is a large constant. We can write}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("`d__n`")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("c", italic = "true", mathvariant = "italic")), mathvariant = "normal"))}{\[\displaystyle -\Mapleoverset{{\it `d__n` }}{\Mapleunderset{z }{\int }}\frac{1}{a \left(t \right)}{\it dt} =-\mathop{\rm ln}\mathop{\rm  }\left(n \mathop{\rm  }c \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Let\mapleinline{inert}{2d}{B(t) = t/a(t)}{$\displaystyle B \left( t \right) ={\frac {t}{a \left( t \right) }}$}
Then\mapleinline{inert}{2d}{(D(B))(t) = 1/a(t) and 1/a(t) > 0}{$\displaystyle \mbox {D} \left( B \right)  \left( t \right) = \left( a \left( t \right)  \right) ^{-1}{\rm \: \wedge \:}0< \left( a \left( t \right)  \right) ^{-1}$}
(notice that\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo("'", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle a \esapos \left(t \right)\essrarr 0$}
as\mapleinline{inert}{2d}{proc (t) options operator, arrow; infinity end proc}{$\displaystyle t\mapsto \infty $}
). Hence\mapleinline{inert}{2d}{B(t)}{$\displaystyle B \left( t \right) $}
is strictly increasing and invertible. So we have}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{B(d__n)-B(z) = ln(n*c)}{\[\displaystyle B \left( {\it d\_\_n} \right) -B \left( z \right) =\ln  \left( nc \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{d__n = [ln(n*c)+B(z)]/B}{\[\displaystyle {\it d\_\_n}={\frac {[\ln  \left( nc \right) +B \left( z \right) ]}{B}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
This is a unique solution.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{3. (d)}We first show that}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(`#mover(mi("F"),mo("&uminus0;"))`(x+a(x)*t)/`#mover(mi("F"),mo("&uminus0;"))`(x), x = x__F) = e^(-t)}{\[\displaystyle \lim _{x\rightarrow {\it x\_\_F}}{\frac {\mbox {{\tt `\#mover(mi("F"),mo("`}}\\
\mbox{}-\mbox {{\tt `"))`}}\\
\mbox{} \left( x+a \left( x \right) t \right) }{\mbox {{\tt `\#mover(mi("F"),mo("`}}\\
\mbox{}-\mbox {{\tt `"))`}}\\
\mbox{} \left( x \right) }}={e}^{-t}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
The LHS is}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("u", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("du", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle \frac{\Mapleoverset{-}{F }\left(x \mathop{\rm  }+\mathop{\rm  }a \left(x \right)\mathop{\rm  }t \right)}{\Mapleoverset{-}{F }\left(x \right)}=\exp \left(-\Mapleoverset{x +t \mathop{\rm  }a \left(x \right)}{\Mapleunderset{x }{\int }}\frac{1}{a \left(u \right)}\mathop{\rm  }{\it du} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Let\mapleinline{inert}{2d}{v = (u-x)/a(x)}{$\displaystyle v={\frac {u-x}{a \left( x \right) }}$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("v", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "[", close = "]")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dv", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle \frac{\Mapleoverset{-}{F }\left(x \mathop{\rm  }+\mathop{\rm  }a \left(x \right)\mathop{\rm  }t \right)}{\Mapleoverset{-}{F }\left(x \right)}=\exp \left(-\Mapleoverset{t }{\Mapleunderset{0}{\int }}\frac{a \left(x \right)}{a \left(v \mathop{\rm  }a \left(x \right)+x \right)}{\it dv} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo("'", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle a \esapos \left(x \right)\essrarr 0$}
as\mapleinline{inert}{2d}{proc (x) options operator, arrow; x__F end proc}{$\displaystyle x\mapsto {\it x\_\_F}$}
we conclude\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("v", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"))}{$\displaystyle a \left(v \mathop{\rm  }a \left(x \right)+x \right)\essrarr a \left(x \right)$}
and thus}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("v", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "[", close = "]")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("1", mathvariant = "normal"))}{\[\displaystyle \frac{a \left(x \right)}{a \left(v \mathop{\rm  }a \left(x \right)+x \right)}\essrarr 1\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
It immediately follows}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(`#mover(mi("F"),mo("&uminus0;"))`(x+a(x)*t)/`#mover(mi("F"),mo("&uminus0;"))`(x), x = x__F) = e^(-t)}{\[\displaystyle \lim _{x\rightarrow {\it x\_\_F}}{\frac {\mbox {{\tt `\#mover(mi("F"),mo("`}}\\
\mbox{}-\mbox {{\tt `"))`}}\\
\mbox{} \left( x+a \left( x \right) t \right) }{\mbox {{\tt `\#mover(mi("F"),mo("`}}\\
\mbox{}-\mbox {{\tt `"))`}}\\
\mbox{} \left( x \right) }}={e}^{-t}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Now we choose\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(d__n) = 1/n}{$\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( {\it d\_\_n} \right) ={n}^{-1}$}
, i.e.\mapleinline{inert}{2d}{d__n = `#msup(mi("F"),mo("&larr;"))`(1-1/n)}{$\displaystyle {\it d\_\_n}=\mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( 1-{n}^{-1} \right) $}
we have}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(n*`#mover(mi("F"),mo("&uminus0;"))`(d__n+a(d__n)*t), x = x__F) = e^(-t)}{\[\displaystyle \lim _{x\rightarrow {\it x\_\_F}}n\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( {\it d\_\_n}+a \left( {\it d\_\_n} \right) t \right) ={e}^{-t}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
By Poisson's approximation}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{P(M__n <= a(d__n)*x+d__n)*`#mover(mo("&rarr;"),mi("d"))`*exp(-e^(-t))}{\[\displaystyle P \left( {\it M\_\_n}\leq a \left( {\it d\_\_n} \right) x+{\it d\_\_n} \right) \mbox {{\tt `\#mover(mo("`}}\rightarrow \mbox {{\tt `"),mi("d"))`}}\\
\mbox{}\\
\mbox{}\,{{\rm e}^{-{e}^{-t}}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
And thus\mapleinline{inert}{2d}{`in`(F, MDA(Lambda))}{$\displaystyle {\it in} \left( F,{\it MDA} \left( \Lambda \right)  \right) $}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{4. (i) Solution:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(z) = C*(int((1+x^2/k)^(-(k+1)*(1/2)), x = z .. infinity))}{\[\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( z \right) =C\int _{z}^{\infty }\! \left( 1+{\frac {{x}^{2}}{k}} \right) ^{-1/2\,k-1/2}{dx}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("C", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("k", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mn("2", mathvariant = "normal"), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"))), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mn("2", mathvariant = "normal"), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), superscriptshift = "0"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("k", italic = "true", mathvariant = "italic")), superscriptshift = "0"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("hypergeom", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mn("2", mathvariant = "normal"), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"))), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mn("2", mathvariant = "normal"), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"))), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mn("2", mathvariant = "normal"), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"))), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mn("2", mathvariant = "normal"), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"))))), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mn("2", mathvariant = "normal"), superscriptshift = "0")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"))), mathvariant = "normal"))}{\[\displaystyle =\mathop{\rm  }C \mathop{\rm  }k ^{\frac{1}{2}~k -\frac{1}{2}}~z ^{-k }~\mathop{\rm hypergeom}\left(\left(\frac{1}{2}~k ,\frac{1}{2}~k +\frac{1}{2}\right),\left(1+\frac{1}{2}~k \right),-\frac{k }{z ^{2}}\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
So\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(z)}{$\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( z \right) $}
is regularly varying with index\mapleinline{inert}{2d}{-k}{$\displaystyle -k$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{4. (ii)}Since\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(x)}{$\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( x \right) $}
is regularly varying,\mapleinline{inert}{2d}{`in`(F, MDA(`&Phi;__k`))}{$\displaystyle {\it in} \left( F,{\it MDA} \left( \Phi {\it \_\_k}\\
\mbox{} \right)  \right) $}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{4. (iii)}we choose\mapleinline{inert}{2d}{c__n = `#msup(mi("F"),mo("&larr;"))`(1-1/n) and `#msup(mi("F"),mo("&larr;"))`(1-1/n) = {C*hypergeom([(1/2)*k, (1/2)*k+1/2], [1+(1/2)*k], -k/z^2)*k^((1/2)*k-1/2)*n}^(1/k)}{$\displaystyle {\it c\_\_n}=\mbox {{\tt `\#msup(mi("F"),mo("`}}\\
\mbox{}\mbox {{\tt `\&larr;`}}\mbox {{\tt `"))`}}\\
\mbox{} \left( 1-{n}^{-1} \right) {\rm \: \wedge \:}\mbox {{\tt `\#msup(mi("F"),mo("`}}\\
\mbox{}\mbox {{\tt `\&larr;`}}\mbox {{\tt `"))`}}\\
\mbox{} \left( 1-{n}^{-1} \right) ={ \left\{ C{\mbox{$_2$F$_1$}(1/2\,k,1/2\,k+1/2;\,1+1/2\,k;\,-{\frac {k}{{z}^{2}}})}\\
\mbox{}{k}^{1/2\,k-1/2}n \right\} }^{{k}^{-1}}$}
and\mapleinline{inert}{2d}{d__n = 0}{$\displaystyle {\it d\_\_n}=0$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{5. (i) Solution:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{P(X^p > x)}{\[\displaystyle P \left( x<{X}^{p} \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("X", italic = "true", mathvariant = "italic"), Typesetting:-mo(">", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), superscriptshift = "0")), mathvariant = "normal"))}{\[\displaystyle =\mathop{\rm  }P \left(X >x ^{\frac{1}{p }}\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), superscriptshift = "0"), Typesetting:-mi(""))}{\[\displaystyle =\mathop{\rm  }x ^{-\frac{\alpha }{p }}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
So\mapleinline{inert}{2d}{X^p}{$\displaystyle {X}^{p}$}
has regularly varying tail with index\mapleinline{inert}{2d}{-alpha/p}{$\displaystyle -{\frac {\alpha}{p}}$}
, and thus belong to\mapleinline{inert}{2d}{MDA(`&Phi;__&beta;`)}{$\displaystyle {\it MDA} \left( \Phi {\it \_\_}\beta  \right) $}
where\mapleinline{inert}{2d}{beta = alpha/p}{$\displaystyle \beta={\frac {\alpha}{p}}$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{P(ln*X > x)}{\[\displaystyle P \left( x< \left( {\it ln} \right) \,X \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("X", italic = "true", mathvariant = "italic"), Typesetting:-mo(">", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-msup(Typesetting:-mi("e", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), superscriptshift = "0")), mathvariant = "normal"))}{\[\displaystyle =\mathop{\rm  }P \left(X >e ^{x }\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("e", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), superscriptshift = "0"))}{\[\displaystyle =\mathop{\rm  }e ^{-\alpha \mathop{\rm  }x }\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle =\mathop{\rm  }\exp \left(-\Mapleoverset{x }{\Mapleunderset{0}{\int }}\alpha \mathop{\rm  }{\it dt} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle =\mathop{\rm  }\exp \left(-\Mapleoverset{x }{\Mapleunderset{0}{\int }}\frac{1}{\alpha ^{-1}}{\it dt} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
This is a von. Mise's function with auxilliary function\mapleinline{inert}{2d}{a(x) = 1/alpha}{$\displaystyle a \left( x \right) ={\alpha}^{-1}$}
. Obviously it satisfies\mapleinline{inert}{2d}{diff(a(x), x) = 0}{$\displaystyle {\frac {d}{dx}}a \left( x \right) =0$}
. Therefore it is in the MDA of Gumbel distribution.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{5. (ii)}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
Following the formula for functions in MDA of Fréchet distribution, we choose, for\mapleinline{inert}{2d}{X^p}{$\displaystyle {X}^{p}$}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{c__n = n^(p/alpha)}{\[\displaystyle {\it c\_\_n}={n}^{{\frac {p}{\alpha}}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
and\mapleinline{inert}{2d}{d__n = 0}{$\displaystyle {\it d\_\_n}=0$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{5. (iii)}Using the formula for von Mise's functions, we choose for\mapleinline{inert}{2d}{ln(X)}{$\displaystyle \ln  \left( X \right) $}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{d__n = ln(n)/alpha}{\[\displaystyle {\it d\_\_n}={\frac {\ln  \left( n \right) }{\alpha}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{c__n = 1/alpha}{\[\displaystyle {\it c\_\_n}={\alpha}^{-1}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{}{\[\displaystyle \]}
}\end{Maple Normal}
}\end{Maple Normal}
\end{document}