### 2015-01-26 thesis/Comparison_to_TW.tex
Add the file that compares the largest eigenvalue's distribution of the Tracy-Widom.

\documentclass{article}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{amssymb}
\usepackage{amsfonts, amsmath}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\input{../physics_common.tex}

\DeclareGraphicsExtensions{.pdf}
\begin{document}
\begin{figure}[htb!]
  \centering
  \subfigure[q = 0.1]{
    \includegraphics[scale=0.5]{../r/eigmax_TW_q0dot1.pdf}
  }
  \subfigure[q = 0.2]{
    \includegraphics[scale=0.5]{../r/eigmax_TW_q0dot2.pdf}
  }
  \caption{\small \it Comparison of the largest eigenvalue's
    distribution with Tracy-Widom.}
  \label{fig:eig1_TW}
\end{figure}

\end{document}