### 2015-01-28 thesis/Comparison_to_TW.tex
Estimated the spectral density of S&P-500 stocks.

\documentclass{article}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{amssymb}
\usepackage{amsfonts, amsmath}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\input{../physics_common.tex}

\DeclareGraphicsExtensions{.pdf}
\begin{document}
How the maximum eigenvalue's distribution compares to Tracy-Widom:

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{../r/eigmax_TW_q0dot1.pdf}
  \caption{\small \it Comparison of the largest eigenvalue's
    distribution with Tracy-Widom. $q=N/T=0.1$}
  \label{fig:eig1_TW_q0.1}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{../r/eigmax_TW_q0dot2.pdf}
  \caption{\small \it Comparison of the largest eigenvalue's
    distribution with Tracy-Widom. $q=N/T=0.2$}
  \label{fig:eig1_TW_q0.2}
\end{figure}

Comparison between the empirical spectral density and the theoretical
density. The value of $v$ used to obtain the theoretical density is
the result of {\it maximum likelihood estimation}.
\begin{figure}[htb!]
  \centering
  \subfigure[q = 0.5, KL = 0.19]{
    \includegraphics[scale=0.4, clip=true, trim=115 271 109
    204]{../pics/spectral_density_q0dot5.pdf}
  }
  \subfigure[q = 0.65, KL = 0.15]{
    \includegraphics[scale=0.4, clip=true, trim=115 271 109
    204]{../pics/spectral_density_q0dot65.pdf}
  }
  \subfigure[q = 0.8, KL=0.14]{
    \includegraphics[scale=0.4, clip=true, trim=115 271 109
    204]{../pics/spectral_density_q0dot8.pdf}
  }
  \subfigure[q = 0.95, KL=0.15]{
    \includegraphics[scale=0.4, clip=true, trim=115 271 109
    204]{../pics/spectral_density_q0dot95.pdf}
  }
  \caption{\small \it Comparison of the empirical and theoretical
    spectral densities. The largest empirical eigenvalue ($\lambda_1$)
    is excluded. The rest is rescaled by $1/\lambda_1$. The
    theoretical density functions are ajusted accordingly. KL is the
    Kullback-Leibler divergence of the theoretical density function from
    the empirical density function. Only intervals on which the
    empirical density is non-zero are included in the calculation of
    KL.}
  \label{fig:spectral_density}
\end{figure}

\end{document}