### 2015-01-29 thesis/Results.tex
Files belonging to the paper with Sven.

\documentclass{article}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{amssymb}
\usepackage{amsfonts, amsmath}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\input{../physics_common.tex}

\DeclareGraphicsExtensions{.pdf}
\begin{document}
Figure \ref{fig:eig1_TW_q0.5} and \ref{fig:eig1_TW_q1.0} are
quantile-to-quantile plots of the maximum eigenvalue's distribution
compared to Tracy-Widom. A straight line is expected if the two
distributions differ only by translation and/or scaling.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{../r/eigmax_TW_q0dot5.pdf}
  \caption{\small \it Comparison of the largest eigenvalue's
    distribution (quantiles along the horizontal axis) with
    Tracy-Widom (quantiles along the vertical axis). $q=N/T=0.5$}
  \label{fig:eig1_TW_q0.5}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{../r/eigmax_TW_q1dot0.pdf}
  \caption{\small \it Comparison of the largest eigenvalue's
    distribution (quantiles along the horizontal axis) with
    Tracy-Widom (quantiles along the vertical axis). $q=N/T=1.0$}
  \label{fig:eig1_TW_q1.0}
\end{figure}

\pagebreak
Figure \ref{fig:eigmax_HillPlot} shows the estimated tail-index of the
maximum eigenvalue distribution against the number of upper-order
statistics used in the estimation (Hill-plot). It is seen that the
region approximately at $(210, 250)$ is stable, suggesting a tail
index of 2.34.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{../r/TailIndex_eignmax_q1dot0_v0dot748.pdf}
  \caption{\small \it Hill plot of the largest eigenvalue's
    distribution. $q = 1$, $v = 0.748$.}
  \label{fig:eigmax_HillPlot}
\end{figure}

\pagebreak
Figure \ref{fig:spectral_density} compares the empirical spectral
density with the theoretical density. The value of $v$ used to obtain
the theoretical density is the result of {\it maximum likelihood
  estimation}.
\begin{figure}[htb!]
  \centering
  \subfigure[q = 0.5, v = 0.9415, KL = 0.19]{
    \includegraphics[scale=0.33, clip=true, trim=115 271 109
    204]{../pics/spectral_density_q0dot5.pdf}
  }
  \subfigure[q = 0.65, v=0.8409, KL = 0.15]{
    \includegraphics[scale=0.33, clip=true, trim=115 271 109
    204]{../pics/spectral_density_q0dot65.pdf}
  }
  \subfigure[q = 0.8, v=0.8808, KL=0.14]{
    \includegraphics[scale=0.33, clip=true, trim=115 271 109
    204]{../pics/spectral_density_q0dot8.pdf}
  }
  \subfigure[q = 0.95, v=0.8859, KL=0.15]{
    \includegraphics[scale=0.33, clip=true, trim=115 271 109
    204]{../pics/spectral_density_q0dot95.pdf}
  }
  \caption{\small \it Comparison of the empirical and theoretical
    spectral densities. The largest empirical eigenvalue ($\lambda_1$)
    is excluded. The rest is rescaled by $1/\lambda_1$. The
    theoretical density functions are rescaled accordingly. KL
    is the Kullback-Leibler divergence of the theoretical density
    function from the empirical density function. Only intervals on
    which the empirical density is non-zero are included in the
    calculation of KL.}
  \label{fig:spectral_density}
\end{figure}

\pagebreak
Figure \ref{fig:SP500-TailIndices} shows the upper- and lower-tail
indices of 441 stocks in the S\&P-500 index.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=94 229 115
  121]{../pics/SP500-TailIndices.pdf}
  \caption{\small \it Upper- and lower-tail indices of 441 stocks in
    the S\&P-500 index. The horizontal coordinate of a point in the
    figure is the upper-tail index and the vertical coordinate is the
    lower-tail index.}
  \label{fig:SP500-TailIndices}
\end{figure}

\end{document}