### 2015-02-12 thesis/Figures.tex
Worked on the paper with Sven.

\documentclass{article}
\usepackage{amssymb,amsmath,amsthm,bm}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage[usenames,dvipsnames]{color}
\usepackage{mathrsfs}
\graphicspath{{/home/xxie/econophysics/pics/InCaseNeeded/}}

\input{/home/xxie/econophysics/physics_common.tex}

\begin{document}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.6]{eigmax_TW.pdf}
  \caption{Distribution of simulated $\lambda_\M$ compared to
    Tracy-Widom. From the 1st to the 4th row, q = 0.1, 0.2, 0.5, 1.}
  \label{fig:eigmax_TW}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.6, clip=true, trim=98 233 120
  127]{AppleLogvol-QQ.pdf}
  \caption{QQ-plot of $\ln{ S_\M - S_\m \over S_\m}$. The data are daily trading statistics of Apple from
    2005-01-02 to 2015-01-23. The lower tail of sample log-vol is a
    bit lighter and the upper tail a bit heavier than their
    counterparts of standard normal. }
  \label{fig:qqplot_logvol}
\end{figure}

\end{document}