### 2016-06-10 mw/garch_covariance/FX.tex
Worked on the minorization.

%% Created by Maple 2015.1, Linux
%% Source Worksheet: FX.mw
%% Generated: Thu Jun 09 15:57:19 CEST 2016
\documentclass{article}
\usepackage{maplestd2e}
\def\emptyline{\vspace{12pt}}
\begin{document}
\pagestyle{empty}
\DefineParaStyle{Maple Heading 1}
\DefineParaStyle{Maple Text Output}
\DefineParaStyle{Maple Dash Item}
\DefineParaStyle{Maple Bullet Item}
\DefineParaStyle{Maple Normal}
\DefineParaStyle{Maple Heading 4}
\DefineParaStyle{Maple Heading 3}
\DefineParaStyle{Maple Heading 2}
\DefineParaStyle{Maple Warning}
\DefineParaStyle{Maple Title}
\DefineParaStyle{Maple Error}
\DefineCharStyle{Maple Hyperlink}
\DefineCharStyle{Maple 2D Math}
\DefineCharStyle{Maple Maple Input}
\DefineCharStyle{Maple 2D Output}
\DefineCharStyle{Maple 2D Input}
\begin{Maple Normal}{
\begin{Maple Normal}{
From\mapleinline{inert}{2d}{x[i, t] := sigma[i, t]*z[i, t]}{$\displaystyle x_{{i,t}}\, := \,\sigma_{{i,t}}z_{{i,t}}$}
and\mapleinline{inert}{2d}{x[j, t] := sigma[j, t]*z[j, t]}{$\displaystyle x_{{j,t}}\, := \,\sigma_{{j,t}}z_{{j,t}}$}
we have}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{x[i, t]*x[j, t] = sigma[i, t]*sigma[j, t]*z[i, t]*z[j, t]}{\[\displaystyle x_{{i,t}}x_{{j,t}}=\sigma_{{i,t}}\sigma_{{j,t}}z_{{i,t}}z_{{j,t}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Since\mapleinline{inert}{2d}{sigma[i, t] in R[-2*alpha[i]]}{$\displaystyle \sigma_{{i,t}} \in R_{{-2\,\alpha_{{i}}}}$}
,\mapleinline{inert}{2d}{sigma[j, t] in R[-2*alpha[j]]}{$\displaystyle \sigma_{{j,t}} \in R_{{-2\,\alpha_{{j}}}}$}
,\mapleinline{inert}{2d}{x[i, t]*x[j, t] in R[-2*min*{alpha[i], alpha[j]}]}{$\displaystyle x_{{i,t}}x_{{j,t}} \in R_{{-2\, \left( \min \right) \, \left\{ \alpha_{{i}},\alpha_{{j}} \right\} }}$}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{(sum(x[i, t]*x[j, t]*`#mover(mo("&rarr;"),mi("v"))`*xi[i, j], t = 1 .. n))/a[n]}{\[\displaystyle {\frac {\sum _{t=1}^{n}x_{{i,t}}x_{{j,t}}\mbox {{\tt `\#mover(mo("`}}\rightarrow \mbox {{\tt `"),mi("v"))`}}\\
\mbox{}\\
\mbox{}\,\xi_{{i,j}}}{a_{{n}}}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
where\mapleinline{inert}{2d}{a[n]}{$\displaystyle a_{{n}}$}
satisfies\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), subscriptshift = "0"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), subscriptshift = "0"), Typesetting:-mo(">", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), subscriptshift = "0")), mathvariant = "normal"), Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"))}{$\displaystyle P \left(x _{i ,t }\cdot x _{j ,t }>a _{n }\right)\estilde \frac{1}{n }$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{(sum(x[i, t]^2, t = 1 .. n))/n}{\[\displaystyle {\frac {\sum _{t=1}^{n}{x_{{i,t}}}^{2}}{n}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-msubsup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal")), Typesetting:-mi(""))}{\[\displaystyle =\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}z _{i ,t }^{2}\mathop{\rm  }\left(\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}a _{i ,1}b _{i ,1}^{l -1}x _{i ,t -l }^{2}+\frac{a _{i ,0}\left(2-b _{i ,1}\right)}{1-b _{i ,1}}\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"))}{\[\displaystyle =\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}a _{i ,1}b _{i ,1}^{l -1}\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}z _{i ,t }^{2}\cdot x _{i ,t -l }^{2}+\frac{a _{i ,0}\left(2-b _{i ,1}\right)}{1-b _{i ,1}}\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}z _{i ,t }^{2}\mathop{\rm  }\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi(""), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), mathvariant = "normal")), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mi(""), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"))}{$\displaystyle =\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}a _{i ,1}b _{i ,1}^{l -1}\left(\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}z _{i ,t }^{2}\cdot x _{i ,t -l }^{2}-\left(\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}z _{i ,t }^{2}\right)\mathop{\rm  }\left(\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}x _{i ,t -l }^{2}\right)\right)+\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}a _{i ,1}b _{i ,1}^{l -1}\left(\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}z _{i ,t }^{2}\right)\mathop{\rm  }\left(\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}x _{i ,t -l }^{2}\right)+\frac{a _{i ,0}\left(2-b _{i ,1}\right)}{1-b _{i ,1}}\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}z _{i ,t }^{2}$}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{sigma[i, t]^2 = sum((b[i, 1]*sigma[i, t-l]^2+a[i, 0])*a[i, 1]^(l-1)*(product(z[i, t-k]^2, k = 1 .. l-1)), l = 1 .. r)+a[i, 1]^r*x[i, t-r]^2*(product(z[i, t-k]^2, k = 1 .. l-1))}{$\displaystyle {\sigma_{{i,t}}}^{2}=\sum _{l=1}^{r} \left(  \left( b_{{i,1}}{\sigma_{{i,t-l}}}^{2}+a_{{i,0}} \right) {a_{{i,1}}}^{l-1}\\
\mbox{}\prod _{k=1}^{l-1}{z_{{i,t-k}}}^{2} \right) +{a_{{i,1}}}^{r}{x_{{i,t-r}}}^{2}\prod _{k=1}^{l-1}{z_{{i,t-k}}}^{2}$}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&prod;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("k", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi(""))}{\[\displaystyle =\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}\left(a _{i ,0}+b _{i ,1}\sigma _{i ,t -l }^{2}\right)a _{i ,1}^{l -1}\mathop{\rm  }\cdot \Mapleoverset{l -1}{\Mapleunderset{k =1}{\prod }}z _{i ,t -k }^{2}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&prod;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("k", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-munderover(Typesetting:-mo("&prod;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("k", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi(""))}{$\displaystyle =\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}a _{i ,0}a _{i ,1}^{l -1}\mathop{\rm  }\cdot \Mapleoverset{l -1}{\Mapleunderset{k =1}{\prod }}z _{i ,t -k }^{2}+\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}b _{i ,1}\cdot a _{i ,1}^{l -1}\cdot \sigma _{i ,t -l }^{2}\cdot \Mapleoverset{l -1}{\Mapleunderset{k =1}{\prod }}z _{i ,t -k }^{2}$}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{sigma[i, t]^2*sigma[j, t]^2 = (sum(a[i, 0]*a[i, 1]^(l-1)*(product(z[i, t-k]^2, k = 1 .. l-1)), l = 1 .. infinity)+sum(b[i, 1]*a[i, 1]^(l-1)*sigma[i, t-l]^2*(product(z[i, t-k]^2, k = 1 .. l-1)), l = 1 .. infinity))*(sum(a[j, 0]*a[j, 1]^(p-1)*(product(z[j, t-q]^2, q = 1 .. p-1)), p = 1 .. infinity)+sum(b[j, 1]*a[j, 1]^(p-1)*sigma[j, t-p]^2*(product(z[j, t-q]^2, q = 1 .. p-1)), p = 1 .. infinity))}{\[\displaystyle {\sigma_{{i,t}}}^{2}{\sigma_{{j,t}}}^{2}= \left( \sum _{l=1}^{\infty } \left( a_{{i,0}}{a_{{i,1}}}^{l-1}\prod _{k=1}^{l-1}{z_{{i,t-k}}}^{2} \right) +\sum _{l=1}^{\infty } \left( b_{{i,1}}{a_{{i,1}}}^{l-1}{\sigma_{{i,t-l}}}^{2}\prod _{k=1}^{l-1}{z_{{i,t-k}}}^{2} \right) \\
\mbox{} \right)  \left( \sum _{p=1}^{\infty } \left( a_{{j,0}}{a_{{j,1}}}^{p-1}\prod _{q=1}^{p-1}{z_{{j,t-q}}}^{2} \right) +\sum _{p=1}^{\infty } \left( b_{{j,1}}{a_{{j,1}}}^{p-1}{\sigma_{{j,t-p}}}^{2}\prod _{q=1}^{p-1}{z_{{j,t-q}}}^{2} \right) \\
\mbox{} \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&Sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mi("&infin;", italic = "false", mathvariant = "normal"), accent = "false", accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&prod;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("k", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Product;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("q", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("q", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), Typesetting:-mn("2", mathvariant = "normal"), superscriptshift = "0", subscriptshift = "0")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&Sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), accent = "false", accentunder = "false"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("p", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), Typesetting:-mn("2", mathvariant = "normal"), superscriptshift = "0", subscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&prod;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("k", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Product;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("q", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("q", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), Typesetting:-mn("2", mathvariant = "normal"), superscriptshift = "0", subscriptshift = "0")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mi(""))}{$\displaystyle =\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}\mathop{\rm  }\Mapleoverset{\infty }{\Mapleunderset{p =1}{\sum }}\mathop{\rm  }a _{i ,0}\mathop{\rm  }a _{j ,0}\mathop{\rm  }a _{i ,1}^{l -1}~a _{j ,1}^{p -1}\mathop{\rm  }\left(\Mapleoverset{l -1}{\Mapleunderset{k =1}{\prod }}z _{i ,t -k }^{2}\right)\mathop{\rm  }\left(\Mapleoverset{p -1}{\Mapleunderset{q =1}{\prod }}\mathop{\rm  }z _{j ,t -q }^{2}\right)+\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}\mathop{\rm  }\Mapleoverset{\infty }{\Mapleunderset{p =1}{\sum }}b _{i ,1}b _{j ,1}\cdot a _{i ,1}^{l -1}a _{j ,1}^{p -1}\cdot \sigma _{i ,t -l }^{2}\mathop{\rm  }\sigma _{j ,t -p }^{2}\mathop{\rm  }\left(\Mapleoverset{l -1}{\Mapleunderset{k =1}{\prod }}z _{i ,t -k }^{2}\right)\mathop{\rm  }\left(\Mapleoverset{p -1}{\Mapleunderset{q =1}{\prod }}\mathop{\rm  }z _{j ,t -q }^{2}\right)$}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{sum(sum(a[i, 0]*a[i, 1]^(l-1)*b[j, 1]*a[j, 1]^(p-1)*sigma[j, t-p]^2*(product(z[i, t-k]^2, k = 1 .. l-1))*(product(z[j, t-q]^2, q = 1 .. p-1)), p = 1 .. infinity), l = 1 .. infinity)+sum(sum(b[i, 1]*a[i, 1]^(l-1)*sigma[i, t-l]^2*a[j, 0]*a[j, 1]^(p-1)*(product(z[i, t-k]^2, k = 1 .. l-1))*(product(z[j, t-q]^2, q = 1 .. p-1)), p = 1 .. infinity), l = 1 .. infinity)}{$\displaystyle \sum _{l=1}^{\infty } \left( \sum _{p=1}^{\infty } \left( a_{{i,0}}{a_{{i,1}}}^{l-1}b_{{j,1}}{a_{{j,1}}}^{p-1}\\
\mbox{}{\sigma_{{j,t-p}}}^{2}\prod _{k=1}^{l-1}{z_{{i,t-k}}}^{2}\prod _{q=1}^{p-1}{z_{{j,t-q}}}^{2} \right)  \right) +\sum _{l=1}^{\infty } \left( \sum _{p=1}^{\infty } \left( b_{{i,1}}{a_{{i,1}}}^{l-1}{\sigma_{{i,t-l}}}^{2}a_{{j,0}}\\
\mbox{}{a_{{j,1}}}^{p-1}\prod _{k=1}^{l-1}{z_{{i,t-k}}}^{2}\prod _{q=1}^{p-1}{z_{{j,t-q}}}^{2} \right)  \right) $}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&Sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), accent = "false", accentunder = "false"), Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0")), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&prod;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("k", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Product;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("q", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("q", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), Typesetting:-mn("2", mathvariant = "normal"), superscriptshift = "0", subscriptshift = "0")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mi(""), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("p", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), Typesetting:-mn("2", mathvariant = "normal"), superscriptshift = "0", subscriptshift = "0"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("p", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), Typesetting:-mn("2", mathvariant = "normal"), superscriptshift = "0", subscriptshift = "0")), Typesetting:-mi("")), mathvariant = "normal"))}{\[\displaystyle =\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}\mathop{\rm  }\Mapleoverset{\infty }{\Mapleunderset{p =1}{\sum }}a _{i ,1}^{l -1}~a _{j ,1}^{p -1}\left(\Mapleoverset{l -1}{\Mapleunderset{k =1}{\prod }}z _{i ,t -k }^{2}\right)\mathop{\rm  }\left(\Mapleoverset{p -1}{\Mapleunderset{q =1}{\prod }}\mathop{\rm  }z _{j ,t -q }^{2}\right)\mathop{\rm  }\left(a _{i ,0}\mathop{\rm  }a _{j ,0}+a _{i ,0}b _{j ,1}\sigma _{j ,t -p }^{2}+a _{j ,0}b _{i ,1}\sigma _{i ,t -l }^{2}+b _{i ,1}b _{j ,1}\sigma _{i ,t -l }^{2}\mathop{\rm  }\sigma _{j ,t -p }^{2}\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\end{document}