### 2016-09-19 papers/FX/fx.tex
Summarized the results about the multivariate GARCH model with Thomas and Olivier.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\usepackage{bookmark}
\usepackage{graphicx}

\newif\ifALL

\ALLtrue

\input{../../physics_common}
\title{On FX Rates and a Multivariate GARCH model}
\author{Xie Xiaolei}
\date{\today}
\begin{document}
\maketitle

\section{Introduction}
We study the covariance matrix of a $p$-dimensional sequence with each
component modeled by a GARCH(1, 1) process:

\begin{eqnarray*}
  X_{i,t} &=& \sigma_{i,t} Z_{i,t} \\
  (Z_{1, t}, Z_{2,t}, \dots, Z_{p,t}) &\sim& N(0, \Sigma) \\
  \sigma_{i, t}^2 &=& \alpha_{i,0} + \alpha_{i,1} X_{i, t-1}^2 + \beta_{i,1} \sigma_{i, t-1}^2
\end{eqnarray*}

\section{IGARCH(1,1) Processes}
We first consider the cases where $\alpha_{i, 1} + \beta_{i, 1} = 1$, i.e. IGARCH models.
\subsection{IARCH(1)}
In this case $\alpha_{i, 1} = 1$, $\beta_{i, 1} = 0$. Thus, from the equation
\begin{equation*}
  \E (Z_i^2)^{\xi_i} = 1
\end{equation*}
we obtain $\xi_i = 1$. Suppose
\begin{eqnarray*}
  \E (Z_i^2 Z_j^2)^\xi &=& 1 \\
  (Z_i, Z_j) &\sim& N(0, \Sigma) \\
  \Sigma &=&
  \begin{pmatrix}
    1 & \rho \\
    \rho & 1
  \end{pmatrix}
\end{eqnarray*}
Re-writing the above equation gives
\begin{equation}
  \label{eq:iarch1_1st}
  {1 \over 2\pi}
  \int \int
  \left(
    \rho^2 x^4 + (1 - \rho^2) x^2 y^2 + 2 \rho \sqrt{1 - \rho^2} x^3 y
  \right)^\xi
  \exp\left(
    -{x^2 + y^2 \over 2}
  \right)
  dx dy = 1
\end{equation}
When $\rho = \pm 1$, it is clear $\xi = 1/2$. When $-1 < \rho < 1$,
we may re-write \eqref{eq:iarch1_1st} in polar coordinates to obtain
\begin{equation*}
  {1 \over 2\pi} \int_0^{\infty} r^{4\xi + 1} e^{-r^2 / 2} dr
  \int_{0}^{2\pi} \cos(\theta)^{2 \xi}
  \left(
    2 \rho^2 \cos(\theta)^2
    + 2 \rho \sqrt{1 - \rho^2} \cos(\theta) \sin(\theta) - \cos(\theta)^2 + 1 - \rho^2
  \right)^\xi d\theta = 1
\end{equation*}
Integrating the radial part of the integral results in
\begin{equation*}
  \int_{0}^{2\pi}
  \cos(\theta)^{2\xi}
  \left(
    2 \rho^2 \cos(\theta)^2 + \rho \sqrt{1 - \rho^2} \sin(2\theta) - \cos(\theta)^2 - \rho^2 + 1
  \right)^\xi d\theta
  = {2 \pi
    \over
    2^{2 \xi} \Gamma(2\xi + 1)
  }
\end{equation*}
After some manipulations we arrive at
\begin{equation}
  \label{eq:iarch_2nd}
  \int_{-1}^1 (\sin(\pi z) + \rho)^{2\xi} dz = {2 \over \Gamma(2\xi + 1)}
\end{equation}
Using this result, we immediately obtain for ARCH(1) models
\begin{equation*}
  \label{eq:arch_1st}
  \int_{-1}^1 (\sin(\pi z) + \rho)^{2\xi} dz =
  {2 \over \Gamma(2\xi + 1)}
  {1 \over (\alpha_i \alpha_j)^\xi}
\end{equation*}

\section{General IARCH(1)}
For general IARCH(1) models, the equation
\begin{equation*}
  \E \left[
    (
    \alpha_{11} Z_1^2 + 1 - \alpha_{11}
    )^\xi
    (
    \alpha_{11} Z_1^2 + 1 - \alpha_{11}
    )^\xi\right] = 1
\end{equation*}
has to be solved numerically.
\section{Spectra of S\&P sectors}
\subsection{Utilities}
Fitting the daily returns' data of 25 companies that have been
included in the S\&P 500 index since 2010-01-01 and categorized in the
sector ``Utilities'', we obtained 23 GARCH(1, 1) models that satisfy
the condition
\begin{equation}
  \label{eq:garch_cond1}
  \alpha_1 + \beta_1 < 1  
\end{equation}
Then we simulate the GARCH(1, 1) processes with innovations having the
same covariance structure as those of the real data. Figure
\ref{fig:Utilities_eigenvalues} and \ref{fig:Utilities_eigenvectors1},
\ref{fig:Utilities_eigenvectors2}
compare the eigenvalues and eigenvectors, respectively, of the 
covariance matrices of the real and the simulated data.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.6]{Utilities_eigenvalues.pdf}
  \caption{Eigenvalues of Utilities: real \& simulated}
  \label{fig:Utilities_eigenvalues}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{Utilities_eigenvectors1.pdf}
  \caption{Eigenvectors of Utilities: real \& simulated}
  \label{fig:Utilities_eigenvectors1}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{Utilities_eigenvectors2.pdf}
  \caption{Eigenvectors of Utilities: real \& simulated continued}
  \label{fig:Utilities_eigenvectors2}
\end{figure}

\subsection{Materials}
A similar analysis on the ``Materials'' sector produces the following
spectra. In this sector, 21 out of 25 return sequencies are
successfully fitted to GARCH(1, 1) models that satisfy condition
\eqref{eq:garch_cond1}.

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.6]{Materials_eigenvalues.pdf}
  \caption{Eigenvalues of Materials: real \& simulated}
  \label{fig:Materials_eigenvalues}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{Materials_eigenvectors1.pdf}
  \caption{Eigenvectors of Materials: real \& simulated}
  \label{fig:Materials_eigenvectors1}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{Materials_eigenvectors2.pdf}
  \caption{Eigenvectors of Materials: real \& simulated continued}
  \label{fig:Materials_eigenvectors2}
\end{figure}

\bibliographystyle{unsrt}
\bibliography{../thesis/econophysics}
\end{document}