### 2017-04-25 cv/CoverLetter/Eurex.tex

%% start of file `template.tex'.
%% Copyright 2006-2013 Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.

\documentclass[10pt,a4paper, gentium]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')

% moderncv themes
\moderncvstyle{banking}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
\moderncvcolor{red}                                % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
\renewcommand{\familydefault}{\sfdefault}         % to set the default font; use '\sfdefault' for the default sans serif font, '\rmdefault' for the default roman one, or any tex font name
%\nopagenumbers{}                                  % uncomment to suppress automatic page numbering for CVs longer than one page

% character encoding
\usepackage[utf8]{inputenc}                       % if you are not using xelatex ou lualatex, replace by the encoding you are using
%\usepackage{CJKutf8}                              % if you need to use CJK to typeset your resume in Chinese, Japanese or Korean

% adjust the page margins
\usepackage[scale=0.75]{geometry}
%\setlength{\hintscolumnwidth}{3cm}                % if you want to change the width of the column with the dates
%\setlength{\makecvtitlenamewidth}{10cm}           % for the 'classic' style, if you want to force the width allocated to your name and avoid line breaks. be careful though, the length is normally calculated to avoid any overlap with your personal info; use this at your own typographical risks...

% personal data
\name{Xie}{Xiaolei}
%\title{Resumé title}                               % optional, remove / comment the line if not wanted
\address{Kiviksgatan 7C}{21440 Malm\"o}{Sweden}% optional, remove / comment the line if not wanted; the "postcode city" and and "country" arguments can be omitted or provided empty
\phone[mobile]{+46~76~060~6638}                   % optional, remove / comment the line if not wanted
%\phone[fixed]{+2~(345)~678~901}                    % optional, remove / comment the line if not wanted
%\phone[fax]{+3~(456)~789~012}                      % optional, remove / comment the line if not wanted
\email{xie.xiaolei@gmail.com}                               % optional, remove / comment the line if not wanted
%\homepage{www.johndoe.com}                         % optional, remove / comment the line if not wanted
%\extrainfo{additional information}                 % optional, remove / comment the line if not wanted
%\photo[64pt][0.4pt]{picture}                       % optional, remove / comment the line if not wanted; '64pt' is the height the picture must be resized to, 0.4pt is the thickness of the frame around it (put it to 0pt for no frame) and 'picture' is the name of the picture file
%\quote{Some quote}                                 % optional, remove / comment the line if not wanted

% to show numerical labels in the bibliography (default is to show no labels); only useful if you make citations in your resume
%\makeatletter
%\renewcommand*{\bibliographyitemlabel}{\@biblabel{\arabic{enumiv}}}
%\makeatother
%\renewcommand*{\bibliographyitemlabel}{[\arabic{enumiv}]}% CONSIDER REPLACING THE ABOVE BY THIS

% bibliography with mutiple entries
%\usepackage{multibib}
%\newcites{book,misc}{{Books},{Others}}
%----------------------------------------------------------------------------------
%            content
%----------------------------------------------------------------------------------
\begin{document}
%-----       letter       ---------------------------------------------------------
% recipient data
% \recipient{Cross Asset Quantitative Analysis team}{Citi}{London}
\recipient{Deutsche B\"orse Group}{Frankfurt am Main}{Germany}
\date{\today}
\opening{Dear Sir/Madam,}
\closing{Best Regards,}
%\enclosure[Attached]{curriculum vit\ae{}}          % use an optional argument to use a string other than "Enclosure", or redefine \enclname
% \enclosure[Attached]{curriculum vit\ae{}, Master's thesis in
%   Econophysics}          % use an optional argument to use a string other than "Enclosure", or redefine \enclname
\makelettertitle

I see on {\em Monster} that you are recruiting a quantitative risk
analyst to join your team of {\em Risk methodology and New products}.
From your description of the responsibilities and requirements, I
understand you are looking for someone who is strong both in
statistical analysis and in software development using Matlab, C++ and
SQL.

As is detailed in my CV, I am anticipating to get my PhD in {\em
  Mathematical Statistics} from Copenhagen University this September
and have co-authored 2 papers on the eigen systems of sample
covariance matrices of heavy tailed time series. At the time of
writing, I am also finalizing two other papers, one about tail
parameters of equity returns and the other about estimation of
the probabilities of large losses by importance sampling.
These results are applicable to principle component analysis,
portfolio optimization as well as risk management; and indeed, I am
keen to apply them and see them making a real impact on business.

Speaking about the application of mathematical models, I think my past
experience as a software developer is worth mentioning. As is detailed
in my CV, I worked as a software developer for 6.5 years in several
companies and accumulated abundant experience in C, C++, R, Matlab
programming as well as in data analysis and algorithm implementation.
Thus I am confident to take on the challenges of a quantitative risk
analyst at {\em Eurex}.

I look forward to discussing the job with you in more
details. You can always call me at +46 76 060 6638 or email me at
xie.xiaolei@gmail.com. It would be my pleasure to hear from you.

\makeletterclosing

\end{document}