### 2017-04-25 cv/References.tex

\title{References}
% \date{\today}
\author{
Xie Xiaolei
}

\documentclass[12pt]{article}

\begin{document}
\maketitle

I would like to refer you to the following contacts for references of
my education and employment.
\begin{itemize}
\item Thomas Mikosch, {\em professor of Mathematics at Copenhagen
    University}, my PhD supervisor
  \begin{itemize}
  \item Email: mikosch@math.ku.dk
  \item Office phone: +45 353 30414
  \end{itemize}
\item Sven \AA berg, {\em professor and Head of Mathematical Physics
    at Lund University}, my Master thesis supervisor
  \begin{itemize}
  \item Email: sven.aberg@matfys.lth.se
  \item Office phone: +46 46 222 96 33
  \item Mobile phone: +46 70 557 35 69
  \end{itemize}
\item Ferdi Aryasetiawan, {\em professor of Mathematical Physics at
    Lund University}, my teacher of multiple courses
  \begin{itemize}
  \item Email: ferdi.aryasetiawan@teorfys.lu.se
  \item Office phone: +46 46 222 90 89
  \end{itemize}
\item Fredrik l\"of, {\em Technical Project Manager and Team Leader at
    Bombardier Transportation}, my supervisor at ST-Ericsson
  \begin{itemize}
  \item Email: fredrik.n.lof@gmail.com
  \item LinkedIn: https://www.linkedin.com/in/fredrik-lof/
  \end{itemize}
\end{itemize}

\end{document}