### 2017-07-21 papers/FX/draft.tex
Added my own thesis.

\documentclass[11pt,a4]{amsart}
\input{../../ThomasDefs}
\begin{document}
\title{Do return series have power-law tails with the same index?}
\author{Thomas Mikosch, Casper de Vries, Xie Xiaolei}
\date{\today}

\maketitle

\input{PaperBody}

\bibliographystyle{unsrt}
\bibliography{../../thesis/econophysics}
\end{document}