### 2017-07-21 phd/Chapterextremes.tex

%extremes Chapter
\chapter[Extreme value analysis for the sample autocovariance matrices of heavy-tailed multivariate time series]{{\huge Extreme value analysis for the sample autocovariance matrices of heavy-tailed multivariate time series}}\label{ch:extremes}
\chaptermark{Sample autocovariance matrices}

\begin{center}
\textsc{Richard Davis, Johannes Heiny, \\Thomas Mikosch \& Xiaolei Xie\\
{\em Extremes 19}, 3 (2016), 517--547.}
\end{center}

\input{../papers/DavisMikoschHeinyXie/PaperBody}