### 2017-08-01 phd/MyThesis.tex
Debugged the program for finding M.

\input{Preamble}
\input{../physics_common}

\usepackage[numbers,sort&compress]{natbib}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{appendix}
\graphicspath{
        {../papers/FX/}{../papers/Jeffrey1/}
        {../papers/Anja/}
        {../papers/DavisMikoschHeinyXie/}
}

\begin{document}

\frontmatter
\include{Titlepage}

\newpage
\leavevmode\thispagestyle{empty}

%\vspace*{1cm}
\hspace*{0.5cm}  {\bf Xiaolei Xie}\\
\hspace*{1.5cm} Department of Mathematical Sciences\\
\hspace*{1.5cm} University of Copenhagen\\
\hspace*{1.5cm} Universitetsparken 5\\
\hspace*{1.5cm} 2100 København Ø\\
\hspace*{1.5cm} Denmark\\
\hspace*{1.5cm} xie@math.ku.dk\\
\hspace*{1.5cm} xie.xiaolei@gmail.com

\vspace*{3cm}

\begin{tabular}{ll}
Supervisor: & Prof.~Thomas Mikosch\\
& University of Copenhagen\\
& \\
Co-supervisor: & Associate Prof.~Jeffrey Collamore\\
& University of Copenhagen\\
& \\
Assessment committee: &  Prof.~Catalin St{\u{a}}ric{\u{a}}\\
& University of Neuchatel, CH\\
& \\

& Prof.~Olivier Wintenberger\\
& University of Marie \& Pierre Curie, Paris\\
& \\

& Lecturer.~Jesper Lund Pedersen (chairman)\\
& University of Copenhagen\\
\end{tabular}

\vspace*{5cm}
\noindent \hspace*{0.5cm} Copyright:\\
\hspace*{0.5cm} Chapters 1-4: \copyright Johannes Heiny (according to Danish legislation)\\
\hspace*{0.5cm} Chapter 2: \copyright Elsevier\\

\noindent \hspace*{0.5cm} ISBN: $978$-$87$-$7078$-$921$-$9$

\begin{comment}
{	\pagestyle{empty}
	\setlength{\topmargin}{0pt}
	\setlength{\headheight}{0pt}
	\setlength{\headsep}{0pt}
	\setlength{\footskip}{0pt}

	\setlength{\parskip}{12pt}

	%\begin{center}
		%\vspace*{2.5cm}
 \large \textbf{\textsc{Johannes Heiny}}

		\vspace*{1cm}

	%\end{center}

	\vspace*{\fill}}
\end{comment}

\chapter*{}
\vspace*{-2cm}
\section*{Abstract}
\addcontentsline{toc}{section}{Abstract}

This thesis is about heavy-tailed time series.

\section*{Resum{\'e}}

\newpage
\section*{Acknowledgments}
\addcontentsline{toc}{section}{Acknowledgments}

I owe my gratitude to many people and especially to my supervisors Prof. Mikosch and Prof. Collamore.
\bigskip

{\em Xie Xiaolei\qquad \qquad \qquad \qquad \qquad Copenhagen, September $2017$.}

\chapter*{Summary}
\addcontentsline{toc}{section}{Summary}
This PhD thesis provides results from analysis of heavy-tailed time
series. Its main part consists of the following research papers
written between October $2014$ and Augest $2017$.

\begin{enumerate}
\bibitem[P2]{davis:mikosch:heiny:xie:20151}
{\sc Davis, R.~A., Heiny, J., Mikosch, T., and Xie, X.}
\newblock Extreme value analysis for the sample autocovariance matrices of
  heavy-tailed multivariate time series.
\newblock {\em Extremes 19}, 3 (2016), 517--547. \href{http://link.springer.com/article/10.1007/s10687-016-0251-7}{\blue{[pdf]}}
\end{enumerate}

Now we explain the structure of this thesis.

{\em Xie Xiaolei \qquad \qquad \qquad \qquad \qquad Copenhagen, January $2017$.}

\newpage
\tableofcontents

\mainmatter

\def\phdthesis{1}
\input{Chapterintroduction}
%------------------------------------------------------------------------------------
\input{ChapterTailParameters}
\input{ChapterImportanceSampling}
\input{Chapterextremes}
\input{ChapterBernoulli}

% \bibliography{libraryjohannes}
\bibliography{../thesis/econophysics}

%for print version only
\newpage
\leavevmode\thispagestyle{empty}
\phantom{bla}

\end{document}